### 2011-07-12 papers/paper1/paper.tex
Added paper.

\documentclass[letterpaper]{article}

\usepackage{outline}

\begin{document}

\title{An I/O Automata Framework for Distributed System Programming}
\author{Justin R. Wilson \and Christopher D. Gill}
\date{}

\maketitle

\begin{abstract}
In an effort to be portable, system developers often favor existing architectures, languages, operating systems, and standard libraries.
The dominant abstraction in established computing technologies for concurrency and therefore distributed system development is threads.
Consequently, distributed system developers are locked into using threads and experience the difficulties of thread-based development including obscure race conditions, poor support for the semantics of distributed systems, and difficulties integrating software developed using different thread models.
Furthermore, the demand for portability and generality dismisses many of the techniques proposed in research that involve new programming languages, language extensions, or domain-specific solutions.
This paper presents a development framework based on the I/O automata formal model implemented in C++.
The I/O automata model is a proven technique for distributed system design because its semantics match those found in distributed systems.
With the framework, developers can reason about correctness directly from the program text using the I/O automata formalism.
The I/O  automata model also provides a common abstraction for concurrency which facilitates software integration and promotes reuse.
TODO:  What are we doing for evaluation?
%% Our ability to realize sophisticated distributed systems such as those in cyber-physical systems and pervasive computing is predicated on a solid model and implementation of asynchrony and concurrency.
%% Threads, the de facto standard in practice, are not a suitable foundation because 1) ad hoc communication among threads prohibits software reuse, 2) threads are difficult to treat formally and therefore \emph{very} difficult to treat informally, and 3) threads are ill-suited for the asynchronous semantics present in the systems we desire to build.
%% In this paper, we outline the semantics necessary for building sophisticated distributed systems and connect these to the I/O automata formal model.
%% We then describe an implementation of the I/O automata formal model and cover extensions to the model necessary for operating in a dynamic environment.
%% Finally, we conclude with an evaluation of our I/O automata framework.
\end{abstract}

\section{Introduction}

Advances in embedded systems and networking are paving the way for distributed systems of unprecedented sophistication.
Cyber-physical systems go beyond the traditional embedded system paradigm by explicitly modeling the dynamics of the physical system and communication network in the computational task.
The inclusion of wireless network interfaces in modern electronic devices is transforming our homes, offices, and hospitals into pervasive computing environments.
The interactions these systems have with the physical environment and its community of users require a degree of interoperability, configurability, adaptability, scalability, robustness, and security not found in existing systems.

One of the forces that prevents the development of any kind of sophisticated software system is accidental complexity~\cite{brooks_nsb}.
In the context of distributed systems, a major source of accidental complexity is a mismatch between the conceptual model of the system and the technology used to implement the system.
The mismatch between the conceptual model and implementation method introduces defects due to translation errors, slows rapid prototyping, and retards iterative refinement.
While there is no ``silver bullet''~\cite{brooks_nsb}, reducing the accidental complexity derived from the mismatch between concept and implementation will increase productivity and make new levels of sophistication possible with existing levels of effort.

Threads are the de facto implementation vehicle for most of computing including distributed systems.
Threads dominate because modern processors, programming languages, and operating systems were all developed to support the thread model.
Furthermore, the widespread availability of thread libraries makes them a natural choice for system development.
While threads dominate in practice, using threads to develop systems has a number of significant drawbacks.

First, ad hoc communication among threads prohibits software reuse.
To illustrate this problem consider the challenge of developing an application with a graphical user interface that also uses a multi-threaded middleware library.
Graphical user interfaces often use a single-threaded event-based design, e.g., X, Swing.
The system integrator must write glue code that bridges the single-threaded event-based semantics to multi-threaded semantics.
Bridging between two thread models is achievable, however, the systems we desire to build will rely on many libraries each with their own thread model.
The importance of a unified approach to concurrency becomes apparent when one considers bridging between different threading models, e.g., event-based, reactor, proactor, active objects, and their attributes, e.g., priority, thread-creation, blocking vs. non-blocking I/O.

Second, threads are difficult to treat formally and therefore \emph{very} difficult to treat informally.
The correctness of programs is reasoned about either formally as part of formal software design and verification and/or informally during the debugging process.
Since most programs will never be formally verified, the correctness of most programs hinges solely on the developers ability to reason about the program.
Based on the historical example of structured programming~\cite{goto_considered_harmful}, there is a strong correlation between what is ``easy'' in the formal arena and what is ``easy'' in the informal arena, especially when the implementation mimics the formal model, e.g., structured programming languages.
It is well known that a formal treatment of threads is difficult due to the arbitrary interleaving of critical sections and ability to define critical sections of arbitrary scope~\cite{lee_threads}.
Thus, the correctness of most multi-threaded programs hinges on the developer's ability to explore arbitrary interleavings of critical sections, i.e., model checking, in their head.

Third, threads are ill-suited for the asynchronous semantics present in the systems we desire to build.
The distributed systems we desire to build are asynchronous at both the device and network levels.
Embedded devices must respond to asynchronous events in the environment.
The interrupt-driven TinyOS operating system for wireless sensor network motes is an exemplar in this area.
The fundamental communication primitive in real-world distributed systems is asynchronous message passing.
Realizing asynchrony in a threaded program requires the use of signals (interrupts) and/or (a)synchronous I/O multiplexing.
Signals are primitive, rigid, and difficult to use correctly.
Synchronous I/O multiplexing, e.g., a select loop, is limited to operating system abstractions, e.g., file descriptors, and induces a reactive state machine structure that is difficult to understand and maintain.
Asynchronous I/O improves efficiency by avoiding the dispatching necessary to determine the I/O operation but are subject to the same limitations of shared-state thread-based programming.

Difficulties in developing asynchronous and concurrent software using threads has prompted a number of concurrent programming languages, e.g., Esterel, Erlang, Ptolemy, Autopipe, and language extensions, e.g., OpenMP, Cilk, Split-C.
Solutions that involve creating a new language or extending an existing language suffer from two main problems.
First, such solutions are often domain-specific and do not generalize to different applications, a necessity for heterogeneous distributed systems.
Furthermore, solutions not connected to a general-purpose formal model lack the important property that the complete system including the environment can be modeled as another component.
Second, such solutions often fail to become mainstream due to practical considerations.
For example, the combination of C and pthreads is much more popular than Erlang due to the popularity of C-like languages and C's connection with UNIX operating system.
Most language extensions are targeted at niche markets and therefore never standardized and implemented in mainstream compilers.

Thus, in order to build sophisticated distributed systems, we require a common semantics and model for building concurrent and asynchronous systems with straight-forward abstractions.
%% The model must be general to support the wide variety of devices and protocols necessary for cyber-physical and pervasive computing systems.
%% The model must facilitate the development of concurrent modules that can be reused.
%% The model must have a strong connection with a formal model to simplify the task of reasoning about individual components and complete systems.
%% The model must be implemented in such a way that it is easily approachable.
In section~\ref{system_model}, we expound upon the requirements of the model and connect the requirements with the I/O automata formal model.
Section~\ref{representation} describes how I/O automata are represented in our I/O automata framework.
In section~\ref{design}, we describe the design and implementation of our I/O automata framework.
Section~\ref{evaluation} provides an evaluation and discussion of our framework.
Section~\ref{related_work} contains an overview of related work and we offer our conclusions and thoughts for future work in section~\ref{conclusion}.

%% We provide a common semantics and model for building concurrent and asynchronous systems with straight-forward abstractions.

%%formal models, e.g., UNITY, I/O Automata,

%% Devices are asynchronous and concurrent.
%% Networks are asynchronous and concurrent.
%% Why not start with a model that is asynchronous and concurrent?

%% What inspires us:
%% \begin{itemize}
%%   \item CORBA - Remote Method Invocation (Remote Procedure Call with a ``this'' pointer).
%%     ``Let's make synchronous function calls over a network.''
%%     Assumes a thread of control.
%%     Asynchronous calls were added later (Did these result in obfuscation?)

%%   \item Patterns for asynchronous concurrency in a synchronous, thread-based world: Active objects, Reactor, Proactor, etc.
%%     Examples: X server, OS2 Presentation Manager, Java Swing

%%   \item AJAX - The modern Web is built on JavaScript + \emph{asynchronous} web page requests = AJAX.

%%   \item Michi Henning - ICE

%%   \item Steve Vinoski - Toolbox of programming languages

%% \end{itemize}

%% We provide a common semantics and model for building concurrent and asynchronous systems with straight-forward abstractions.

\pagebreak

\section{System Model and Formalization\label{system_model}}

Our system model is influenced by the semantics of distributed systems, the semantics of devices, and the desire for conceptual integrity.

\paragraph{Local State.}
First, we require that all modules manipulate local state only including modules that might be collocated on the same distributed node. 
Said another way, there is no global or shared state.
Computation designed around the manipulation of local state only is easily modularized and consequently, easier to distribute and execute concurrently.

\paragraph{Message Passing.}
Nodes in distributed systems interact by passing messages over communication links.
Message passing can either be synchronous or asynchronous.
In synchronous message passing, the events of message production and delivery are the same, i.e., atomic.
In asynchronous message passing, the events of message production and delivery are logically distinct events subject to being interposed by other events.
Synchronous message passing is most often used between modules on the same node while asynchronous message passing is most often used between nodes.
Both synchronous and asynchronous message passing are required.
Synchronous message passing is necessary for building composable systems because of it association with atomicity.
Asynchronous message passing is necessary for building real distributed systems as this matches the semantics of the networks found in practice.

\paragraph{Asynchronous Events.}
Events are natural way to deal with the asynchrony involved in receiving messages from the network or interacting with the physical environment.
A programming system for distributed systems, therefore, should provide semantics that allow modules to both generate and receive events in a general and principled way.
Compare this to UNIX signals that are neither general (fixed-number with no associated data) nor principled

%% Communicating with external resources such as the operating system or physical environment is foundational to writing useful programs.
%% Consider the history of I/O in the C run-time library.
%% The synchronous function call semantics of the C programming language made it possible to write system calls, e.g., read and write, that transfer control to the operating system which then completes the operation and return control to the process after the system call.
%% This is called \emph{blocking I/O} because the process blocks while it waits for the operating system to complete the I/O.
%% Blocking I/O has the severe restriction that only one I/O source can be used at a time.
%% To remedy this, system calls such as select and poll were introduced that allow a process to monitor I/O sources for events and then respond to these events.
%% Asynchronous I/O was a further advancement that allows for the system (instead of the process) to dispatch a function upon the completion of I/O.

%% Interrupts are difficult to use correctly.

\begin{outline}
\item State
  \begin{outline}
  \item There is no shared state in a 
  \item Local state only
  \item Shared state
    \begin{outline}
      \item Impossible in distributed systems
      \item Dangerous in local systems
    \end{outline}
  \end{outline}
\item Communication
  \begin{outline}
  \item Atomic asynchronous message passing
  \item Network sets size of atom (UDP)
  \item Can build reliable streams (TCP)
  \item Local equivalent is passing a value
  \item Model should lend itself to writing protocols
  \end{outline}
\item Asynchrony
  \begin{outline}
    \item Model must have natural support for asynchrony, i.e., event-based
    \item Leads to a more efficient implementation because changed state and enabled actions become obvious
  \end{outline}
\item Concurrency
  \begin{outline}
    \item Reason about systems using non-deterministic interleaving of atomic actions
    \item Model should admit implementations that execute concurrently
  \end{outline}
\item Dynamics
  \begin{outline}
    \item Configuration - Edges in graph of communicating components can change at run-time.
      \begin{outline}
      \item Already required in distributed settings
      \item Not addressed in formal models
      \end{outline}
    \item Extension - Nodes in graph of communicating components can change at run-time.
  \end{outline}
  \item Reflection
\end{outline}

I/O Automata
\begin{itemize}
  \item Compare with UNITY
  \item Compare with esterel
  \item Compare with pi calculus
  \item Compare with Ptolemy
\end{itemize}

\section{I/O Automata Representation and Programming Model\label{representation}}

Conclude with our recommendations for programming, highlight need for exploration.

\section{Design and Implementation\label{design}}

\begin{itemize}
  \item Problem
  \item Design forces
  \item Solution
  \item Consequences
\end{itemize}

\section{Evaluation and Discussion\label{evaluation}}

\begin{itemize}
\item Translation of automaton in \emph{Distributed Algorithms} to C++.
\item Simulate a protocol then replace with network components
\item Compare a protocol using: single-threaded event-based (select loop), multi-threaded, I/O automata
\item Show a buggy program and then apply an invariant to find the bug.
\end{itemize}

Bring up style mentioned in section \ref{representation}.

\section{Related Work\label{related_work}}

pthreads (?)
Edward Lee - The Problem with Threads
Herb Sutter - The Free Lunch is Over
Early work on concurrency

\section{Conclusion and Future Work\label{conclusion}}

\begin{itemize}
  \item We are going to use it to build the substrate.
  \item Speculate on moving down into operating system (device drivers would be easy, IPC including filesystem replaced by automata)
  \item Speculate on moving down into the hardware level (Local talent, Ivan Sutherland)
  \item We can take advantage of multi-core in a very straight-forward way
  \item New problems in scheduling (Pinning automata to processors to minimize actions that span two processors.  Maximum independent set.)
  \item We are non-blocking all the way.  Combine this with a deterministic implementation of the scheduler and model and there are serious opportunities for real-time.
\end{itemize}

\end{document}

### 2011-07-13 papers/paper1/paper.tex
Made a little headway in the Model section of the paper

@@
 \documentclass[letterpaper]{article}
-
-\usepackage{outline}
 
@@
 
-\title{An I/O Automata Framework for Distributed System Programming}
-\author{Justin R. Wilson \and Christopher D. Gill}
+\title{Multicast File Transfer Protocol}
+\author{Adam Drescher \and Magdalena Cassel}
 \date{}
@@
 
-\section{System Model and Formalization\label{system_model}}
+\section{Model\label{model}}
 
-Our system model is influenced by the semantics of distributed systems, the semantics of devices, and the desire for conceptual integrity.
+Moving data across a network can take a variety of different forms.
+These forms correspond to different user needs, and consequently different protocols.
+Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network.
+%% This is useful for 1) device discovery 2) file transfer in the absence of many technologies (including internet).
+the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
+\paragraph{Files.}
+Files are a sequence of bytes with a length and a user-defined type. They are broken down into equally sized pieces called fragments.
 
-\paragraph{Local State.}
-First, we require that all modules manipulate local state only including modules that might be collocated on the same distributed node. 
-Said another way, there is no global or shared state.
-Computation designed around the manipulation of local state only is easily modularized and consequently, easier to distribute and execute concurrently.
+\paragraph{File ID's.}
+A structure that contains a hash, length, and type. Each file ID is unique to its file.
 
-\paragraph{Message Passing.}
-Nodes in distributed systems interact by passing messages over communication links.
-Message passing can either be synchronous or asynchronous.
-In synchronous message passing, the events of message production and delivery are the same, i.e., atomic.
-In asynchronous message passing, the events of message production and delivery are logically distinct events subject to being interposed by other events.
-Synchronous message passing is most often used between modules on the same node while asynchronous message passing is most often used between nodes.
-Both synchronous and asynchronous message passing are required.
-Synchronous message passing is necessary for building composable systems because of it association with atomicity.
-Asynchronous message passing is necessary for building real distributed systems as this matches the semantics of the networks found in practice.
+\paragraph{Matching.}
+Given a file $f_1$ and $f_2$, one can define a matching predicate $\mu$.
 
-\paragraph{Asynchronous Events.}
-Events are natural way to deal with the asynchrony involved in receiving messages from the network or interacting with the physical environment.
-A programming system for distributed systems, therefore, should provide semantics that allow modules to both generate and receive events in a general and principled way.
-Compare this to UNIX signals that are neither general (fixed-number with no associated data) nor principled
-
+\paragraph{Message types.}
+Request: a file ID and a set of fragment indices. We implemented requests as the file ID and a set of contiguous regions.  However, one can envision the request to contain anything that defines a set.
+Fragment: a file ID, fragment index, and the data itself.
+Match: a file ID and a set of matching files.
+\paragraph{How it's put together.} %% change
+In a simple scenario, if you get a request for a file you have fragments for, you can send those fragments out.
+Fragments are periodically broadcasted to indicate what files exist in the network.
+A client can use requests to pull for file fragments.
+There are two different types of matching which are not commutative (??). 
+The first using $\mu_0$ to take two file ID's and determine whether there could be a match between the two files.  
+The second using $\mu_1$ to take two files and determine whether or not there is a match.
+If $mu_1$($f_1$,$f_2$), $mu_0$($f_1$,$f_2$) must be true.
+However, if $mu_0$($f_1$,$f_2$) is true, $mu_1$($f_1$,$f_2$) does \emph{not} have to be true. 
 %% Communicating with external resources such as the operating system or physical environment is foundational to writing useful programs.
@@
 %% Interrupts are difficult to use correctly.
-
-\begin{outline}
-\item State
-  \begin{outline}
-  \item There is no shared state in a 
-  \item Local state only
-  \item Shared state
-    \begin{outline}
-      \item Impossible in distributed systems
-      \item Dangerous in local systems
-    \end{outline}
-  \end{outline}
-\item Communication
-  \begin{outline}
-  \item Atomic asynchronous message passing
-  \item Network sets size of atom (UDP)
-  \item Can build reliable streams (TCP)
-  \item Local equivalent is passing a value
-  \item Model should lend itself to writing protocols
-  \end{outline}
-\item Asynchrony
-  \begin{outline}
-    \item Model must have natural support for asynchrony, i.e., event-based
-    \item Leads to a more efficient implementation because changed state and enabled actions become obvious
-  \end{outline}
-\item Concurrency
-  \begin{outline}
-    \item Reason about systems using non-deterministic interleaving of atomic actions
-    \item Model should admit implementations that execute concurrently
-  \end{outline}
-\item Dynamics
-  \begin{outline}
-    \item Configuration - Edges in graph of communicating components can change at run-time.
-      \begin{outline}
-      \item Already required in distributed settings
-      \item Not addressed in formal models
-      \end{outline}
-    \item Extension - Nodes in graph of communicating components can change at run-time.
-  \end{outline}
-  \item Reflection
-\end{outline}
 

### 2011-07-13 papers/paper1/paper.tex
Revised structure of model paper and added some transitions.

@@
 the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
-\paragraph{Files.}
-Files are a sequence of bytes with a length and a user-defined type. They are broken down into equally sized pieces called fragments.
-
-\paragraph{File ID's.}
-A structure that contains a hash, length, and type. Each file ID is unique to its file.
-
-\paragraph{Matching.}
+The protocol employs a number of conceptually simple structures--a reflection of its overall simplicity, both in design and in purpose.
+One of these structures is aptly called a file. 
+Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called fragments.  
+Each file has a respective file ID, which is a structure that contains a hash, length, and type. 
+Each file ID is unique to its file.
+From these different file and file ID's a matching concept is established.
 Given a file $f_1$ and $f_2$, one can define a matching predicate $\mu$.
-
-\paragraph{Message types.}
-Request: a file ID and a set of fragment indices. We implemented requests as the file ID and a set of contiguous regions.  However, one can envision the request to contain anything that defines a set.
-Fragment: a file ID, fragment index, and the data itself.
-Match: a file ID and a set of matching files.
-\paragraph{How it's put together.} %% change
+Finally, there are three types of files which are conceptually important to the protocol, the first of which being a request.
+A request contains a file ID and a set of fragment indices. 
+We implemented requests as the file ID and a set of contiguous regions.  
+However, one can envision the request to contain anything that defines a set.
+The second type is a fragment, which consists of a file ID, fragment index, and the data itself.
+Last of all is a match, which consists of a file ID and a set of matching files.
+\paragraph {}
+With all of the conceptual structures in place, it is time to define their interactions.
+%% TODO: change above transition, it is terrible but I couldn't think of anything else at the time.
 In a simple scenario, if you get a request for a file you have fragments for, you can send those fragments out.
@@
 A client can use requests to pull for file fragments.
-There are two different types of matching which are not commutative (??). 
+Two different types of matching can be established.
+%% There are two different types of matching which are not commutative (??). 
 The first using $\mu_0$ to take two file ID's and determine whether there could be a match between the two files.  
 The second using $\mu_1$ to take two files and determine whether or not there is a match.
+However, the two matching predicates are not commutative.
 If $mu_1$($f_1$,$f_2$), $mu_0$($f_1$,$f_2$) must be true.
@@
 
-I/O Automata
-\begin{itemize}
-  \item Compare with UNITY
-  \item Compare with esterel
-  \item Compare with pi calculus
-  \item Compare with Ptolemy
-\end{itemize}
+%%I/O Automata
+%%\begin{itemize}
+%%  \item Compare with UNITY
+%%  \item Compare with esterel
+%%  \item Compare with pi calculus
+%%  \item Compare with Ptolemy
+%%\end{itemize}
 

### 2011-07-13 papers/paper1/paper.tex
Thoroughly revised Model.

--- 2011-07-13 papers/paper1/paper.tex
@@
 
-Moving data across a network can take a variety of different forms.
-These forms correspond to different user needs, and consequently different protocols.
+Moving data across a network can take a variety of forms.
+These forms correspond to different user needs and consequently different protocols.
 Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network.
-%% This is useful for 1) device discovery 2) file transfer in the absence of many technologies (including internet).
+%% This is useful for 1) device discovery 2) File transfer in the absence of many technologies (including internet).
 the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
-\paragraph{Files.}
-Files are a sequence of bytes with a length and a user-defined type. They are broken down into equally sized pieces called fragments.
-
-\paragraph{File ID's.}
-A structure that contains a hash, length, and type. Each file ID is unique to its file.
-
-\paragraph{Matching.}
-Given a file $f_1$ and $f_2$, one can define a matching predicate $\mu$.
-
-\paragraph{Message types.}
-Request: a file ID and a set of fragment indices. We implemented requests as the file ID and a set of contiguous regions.  However, one can envision the request to contain anything that defines a set.
-Fragment: a file ID, fragment index, and the data itself.
-Match: a file ID and a set of matching files.
-\paragraph{How it's put together.} %% change
-In a simple scenario, if you get a request for a file you have fragments for, you can send those fragments out.
-Fragments are periodically broadcasted to indicate what files exist in the network.
-A client can use requests to pull for file fragments.
-There are two different types of matching which are not commutative (??). 
-The first using $\mu_0$ to take two file ID's and determine whether there could be a match between the two files.  
-The second using $\mu_1$ to take two files and determine whether or not there is a match.
-If $mu_1$($f_1$,$f_2$), $mu_0$($f_1$,$f_2$) must be true.
-However, if $mu_0$($f_1$,$f_2$) is true, $mu_1$($f_1$,$f_2$) does \emph{not} have to be true. 
-%% Communicating with external resources such as the operating system or physical environment is foundational to writing useful programs.
-%% Consider the history of I/O in the C run-time library.
-%% The synchronous function call semantics of the C programming language made it possible to write system calls, e.g., read and write, that transfer control to the operating system which then completes the operation and return control to the process after the system call.
-%% This is called \emph{blocking I/O} because the process blocks while it waits for the operating system to complete the I/O.
-%% Blocking I/O has the severe restriction that only one I/O source can be used at a time.
-%% To remedy this, system calls such as select and poll were introduced that allow a process to monitor I/O sources for events and then respond to these events.
-%% Asynchronous I/O was a further advancement that allows for the system (instead of the process) to dispatch a function upon the completion of I/O.
-
-%% Interrupts are difficult to use correctly.
-
-I/O Automata
-\begin{itemize}
-  \item Compare with UNITY
-  \item Compare with esterel
-  \item Compare with pi calculus
-  \item Compare with Ptolemy
-\end{itemize}
+\paragraph {Components.}
+The protocol employs a number of conceptually simple structures--a reflection of its overall simplicity, both in design and in purpose.
+One of these structures is called a File. 
+Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called fragments.  
+Each File has a respective File ID, which is a structure that contains a hash of the file content, length, and type. 
+Each File ID is unique to its File.
+From these different File and File IDs a matching concept is established: given a File $f_1$ and $f_2$, one can define a matching predicate $\mu$.
+Finally, there are three types of Files which are conceptually important to the protocol, the first of which being a request.
+A request contains a File ID and a set of contiguous regions.
+In our implementation, the set of contiguous regions was a set of fragment indices, but one can envision the request to contain anything that defines a set.
+%%A request contains a File ID and a set of fragment indices. 
+%%We implemented requests as the File ID and a set of contiguous regions.  
+%%However, one can envision the request to contain anything that defines a set.
+The second type is a fragment, which consists of a File ID, fragment index, and the data itself.
+Last of all is a match, which consists of a File ID and a set of matching File IDs.
+\paragraph {Component Interactions.}
+In a simple scenario, if one gets a request for a File one has fragments for, one can send those fragments out.
+Fragments are periodically broadcasted to indicate what Files exist in the network.
+A client can use requests to pull for File fragments, which can span from individual fragments to the whole file.
+Two different matching steps are established.
+The first uses $\mu_0$ to take two File ID's and determine whether there could be a match between the two Files.  
+If this is true, the second uses $\mu_1$ to take two Files and determine whether or not there is a match.
+If $\mu_1$($f_1$,$f_2$) is true, $\mu_0$($fid_1$,$fid_2$) must be true.
+However, if $\mu_0$($fid_1$,$fid_2$) is true, $\mu_1$($f_1$,$f_2$) does \emph{not} have to be true.
 

### 2011-07-13 papers/paper1/paper.tex
Revised Model section.

@@
 One of these structures is called a File. 
-Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called fragments.  
-Each File has a respective File ID, which is a structure that contains a hash of the file content, length, and type. 
+Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called \emph{fragments}.  
+Each File has a File ID, which is a structure that contains a hash of the File content, the File's length, and the File's type. 
 Each File ID is unique to its File.
-From these different File and File IDs a matching concept is established: given a File $f_1$ and $f_2$, one can define a matching predicate $\mu$.
-Finally, there are three types of Files which are conceptually important to the protocol, the first of which being a request.
-A request contains a File ID and a set of contiguous regions.
+From Files and File IDs a matching concept is established: given two Files $f_1$ and $f_2$, one can define a matching predicate $\mu$.
+Finally, there are three types of Files which are conceptually important to the protocol, the first of which being a \emph{request}.
+A \emph{request} contains a File ID and a set of contiguous regions of the File.
 In our implementation, the set of contiguous regions was a set of fragment indices, but one can envision the request to contain anything that defines a set.
@@
 %%However, one can envision the request to contain anything that defines a set.
-The second type is a fragment, which consists of a File ID, fragment index, and the data itself.
-Last of all is a match, which consists of a File ID and a set of matching File IDs.
+The second type is a \emph{fragment}, which consists of a File ID, a fragment index, and the data comprising that fragment of the file.
+Last is a \emph{match}, which consists of a File ID and a set of matching File IDs.
 \paragraph {Component Interactions.}
-In a simple scenario, if one gets a request for a File one has fragments for, one can send those fragments out.
-Fragments are periodically broadcasted to indicate what Files exist in the network.
-A client can use requests to pull for File fragments, which can span from individual fragments to the whole file.
-Two different matching steps are established.
-The first uses $\mu_0$ to take two File ID's and determine whether there could be a match between the two Files.  
-If this is true, the second uses $\mu_1$ to take two Files and determine whether or not there is a match.
+In a simple scenario, if a server gets a request for a File it has fragments for, it can send those fragments out.
+Fragments are periodically broadcast even in the absence of requests to indicate what Files exist in the network.
+A client can use requests to pull for regions of a File which can span from individual fragments to the whole file.
+A client which has successfully downloaded any number of fragments of a File can then serve those fragments.
+%%Two different matching steps are established.
+Matching is a two-step process.
+The first step applies $\mu_0$ to two File IDs to determine whether a match is possible between the two Files.  
+The second step applies $\mu_1$ to the two Files to determine whether there is a match.
 If $\mu_1$($f_1$,$f_2$) is true, $\mu_0$($fid_1$,$fid_2$) must be true.

### 2011-07-13 papers/paper1/paper.tex
Comments from Justin on model.

@@
 
+% jrw:  Move this to the introduction.
 Moving data across a network can take a variety of forms.
@@
 the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
-\paragraph {Components.}
+
+\paragraph {Files.}
+% jrw:  This rhetoric is unnecessary, i.e., there is no need to claim that you are simpler unless you are comparing yourself to someone else.
+% jrw:  I would start with ``The main concept in MFTP is that of a file. ...''
 The protocol employs a number of conceptually simple structures--a reflection of its overall simplicity, both in design and in purpose.
+% jrw:  Lowercase File here and following.  Emphasize first time.
 One of these structures is called a File.
+% jrw:  Fragments should be a separate sentence.  Since you doing definitions, some of the sentences will be short and choppy.  This is okay.
 Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called \emph{fragments}.  
 Each File has a File ID, which is a structure that contains a hash of the File content, the File's length, and the File's type. 
+% jrw:  Maybe ``Each file is uniquely identified by its fileid.''
 Each File ID is unique to its File.
+
+\paragraph{Matching.}
+% jrw:  This is awkward.  Maybe ``Given two file f1 and f2, one can define a matching predicate mu.''
 From Files and File IDs a matching concept is established: given two Files $f_1$ and $f_2$, one can define a matching predicate $\mu$.
+% jrw:  These are messages not files.  Define the purpose of the message before describing its contents.
 Finally, there are three types of Files which are conceptually important to the protocol, the first of which being a \emph{request}.
+% jrw:  Backwards.  A request contains a set of fragment indices.  (Might need to define fragment index.)  Our implementation has requests that contain a set of spans.
 A \emph{request} contains a File ID and a set of contiguous regions of the File.
@@
 Last is a \emph{match}, which consists of a File ID and a set of matching File IDs.
-\paragraph {Component Interactions.}
+
+\paragraph {Tasks.}
+% jrw:  MFTP is a peer-to-peer (?) or peer-based (?) protocol since everyone can be both a client and a server.
 In a simple scenario, if a server gets a request for a File it has fragments for, it can send those fragments out.
 Fragments are periodically broadcast even in the absence of requests to indicate what Files exist in the network.
-A client can use requests to pull for regions of a File which can span from individual fragments to the whole file.
+A client can use requests to poll for fragments of a File which can span from individual fragments to the whole file.
 A client which has successfully downloaded any number of fragments of a File can then serve those fragments.
@@
 Matching is a two-step process.
+% jrw:  mu_0 came out of left field.
 The first step applies $\mu_0$ to two File IDs to determine whether a match is possible between the two Files.  

### 2011-07-13 papers/paper1/paper.tex
another revision of Model

--- 2011-07-13 papers/paper1/paper.tex
@@
 the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
-\paragraph {Components.}
-The protocol employs a number of conceptually simple structures--a reflection of its overall simplicity, both in design and in purpose.
-One of these structures is called a File. 
-Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called fragments.  
-Each File has a respective File ID, which is a structure that contains a hash of the file content, length, and type. 
-Each File ID is unique to its File.
-From these different File and File IDs a matching concept is established: given a File $f_1$ and $f_2$, one can define a matching predicate $\mu$.
-Finally, there are three types of Files which are conceptually important to the protocol, the first of which being a request.
-A request contains a File ID and a set of contiguous regions.
-In our implementation, the set of contiguous regions was a set of fragment indices, but one can envision the request to contain anything that defines a set.
-%%A request contains a File ID and a set of fragment indices. 
-%%We implemented requests as the File ID and a set of contiguous regions.  
-%%However, one can envision the request to contain anything that defines a set.
-The second type is a fragment, which consists of a File ID, fragment index, and the data itself.
-Last of all is a match, which consists of a File ID and a set of matching File IDs.
-\paragraph {Component Interactions.}
-In a simple scenario, if one gets a request for a File one has fragments for, one can send those fragments out.
-Fragments are periodically broadcasted to indicate what Files exist in the network.
-A client can use requests to pull for File fragments, which can span from individual fragments to the whole file.
-Two different matching steps are established.
-The first uses $\mu_0$ to take two File ID's and determine whether there could be a match between the two Files.  
-If this is true, the second uses $\mu_1$ to take two Files and determine whether or not there is a match.
+
+\paragraph {Files.}
+The main concept in MFTP is that of a \emph{file}.
+Files are a sequence of bytes with a length and a user-defined type.
+They are broken down into equally sized pieces called \emph{fragments}.  
+Each file is uniquely identified by its \emph{file ID}, a structure that contains a hash of the file content, the file's length, and the file's type.
+Given two files $f_1$ and $f_2$, one can define a matching predicate $\mu$.
+
+\paragraph{Messages.}
+File transfer takes place via \emph{messages}.
+There are three types of messages in the protocol, the first of which being a \emph{request}.
+A \emph{request} contains a file ID and a set of fragment indices.
+We implemented requests as the file ID and a set of contiguous regions.  
+However, one can envision the request to contain anything that defines a set.
+The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
+Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.
+
+\paragraph {Tasks.}
+MFTP is a peer-to-peer protocol since all participants can act as both a client and a server.
+In a simple scenario, if a peer gets a request for a file it has fragments for, it can send those fragments out.
+Fragments are periodically broadcast even in the absence of requests to indicate what files exist in the network.
+A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
+A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.
+Matching is a two-step process.
+The first step applies a matching function $\mu_0$ to two file IDs to determine whether a match is possible between the two files.  
+The second step applies a matching function $\mu_1$ to the two files to determine whether there is a match.
 If $\mu_1$($f_1$,$f_2$) is true, $\mu_0$($fid_1$,$fid_2$) must be true.

### 2011-07-13 papers/paper1/paper.tex
yet another revision of Model

@@
 File transfer takes place via \emph{messages}.
-There are three types of messages in the protocol, the first of which being a \emph{request}.
-A \emph{request} contains a file ID and a set of fragment indices.
+There are three types of messages in the protocol.
+The first message type is a \emph{request}.
+A request contains a file ID and a set of fragment indices.
 We implemented requests as the file ID and a set of contiguous regions.  
@@
 \paragraph {Tasks.}
-MFTP is a peer-to-peer protocol since all participants can act as both a client and a server.
-In a simple scenario, if a peer gets a request for a file it has fragments for, it can send those fragments out.
-Fragments are periodically broadcast even in the absence of requests to indicate what files exist in the network.
+MFTP is a peer-to-peer protocol because each participant can act as both a client and a server.
+A peer can send out any fragments for which it gets a request.
+Fragments are periodically broadcast, even in the absence of requests, to indicate what files exist in the network.
 A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
 A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.
+
 Matching is a two-step process.

### 2011-07-14 papers/paper1/paper.tex
Small changes

@@
 \section{Introduction}
+% jrw:  Move this to the introduction.
+Moving data across a network can take a variety of forms.
+These forms correspond to different user needs and consequently different protocols.
+Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network.
+%% This is useful for 1) device discovery 2) File transfer in the absence of many technologies (including internet).
+the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
 
@@
 
-% jrw:  Move this to the introduction.
-Moving data across a network can take a variety of forms.
-These forms correspond to different user needs and consequently different protocols.
-Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network.
-%% This is useful for 1) device discovery 2) File transfer in the absence of many technologies (including internet).
-the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
-
 \paragraph {Files.}
@@
 
-\section{I/O Automata Representation and Programming Model\label{representation}}
+%%TODO: Talk about validation
 
-Conclude with our recommendations for programming, highlight need for exploration.
+\section{Implementation}
 
-\section{Design and Implementation\label{design}}
+%%TODO: Talk about validation - possible coinage 'incremental digest'
 
-\begin{itemize}
-  \item Problem
-  \item Design forces
-  \item Solution
-  \item Consequences
-\end{itemize}
+\paragraph{I/O Automata}
+We used an I/O Automata framework to develop an implementation of the MFTP protocol.
+General overview of the framework and formal model.
+Describe our implementation at the 50k ft view and work our way done until things are no longer interesting.
+Justin's picture:
+
+The MFTP protocol is built off of an I/O Automata framework.  I/O Automata is *blank* (cite Nancy Lynch). The framework we used for our purposes is a C++ framework developed by Justin R. Wilson.
+
+%%\begin{itemize}
+%%  \item Problem
+%%  \item Design forces
+%%  \item Solution
+%%  \item Consequences
+%%\end{itemize}
 

### 2011-07-15 papers/paper1/paper.tex
Implementation/puesto en practica

@@
 \documentclass[letterpaper]{article}
-
+\usepackage{graphicx}
 \begin{document}
@@
 Matching is a two-step process.
-The first step applies a matching function $\mu_0$ to two file IDs to determine whether a match is possible between the two files.  
-The second step applies a matching function $\mu_1$ to the two files to determine whether there is a match.
-If $\mu_1$($f_1$,$f_2$) is true, $\mu_0$($fid_1$,$fid_2$) must be true.
-However, if $\mu_0$($fid_1$,$fid_2$) is true, $\mu_1$($f_1$,$f_2$) does \emph{not} have to be true.
+The first step applies a matching function $\iota$ to two file IDs to determine whether a match is possible between the two files.  
+The second step applies a matching function $\mu$ to the two files to determine whether there is a match.
+If $\mu$($f_1$,$f_2$) is true, $\iota$($fid_1$,$fid_2$) must be true.
+However, if $\iota$($fid_1$,$fid_2$) is true, $\mu$($f_1$,$f_2$) does \emph{not} have to be true.
 
@@
 \paragraph{I/O Automata}
-We used an I/O Automata framework to develop an implementation of the MFTP protocol.
-General overview of the framework and formal model.
-Describe our implementation at the 50k ft view and work our way done until things are no longer interesting.
-Justin's picture:
+We used the ioa framework to develop an implementation of the MFTP protocol.  I/O Automata model stateful distributed system components that can interact and be composed with each other.
+They change state via transitions classified as input, output or internal actions.
+Input and output actions in different automata are bound; output actions in the one serve as interrupts to trigger input actions in the other.
 
-The MFTP protocol is built off of an I/O Automata framework.  I/O Automata is *blank* (cite Nancy Lynch). The framework we used for our purposes is a C++ framework developed by Justin R. Wilson.
+%%General overview of the framework and formal model.
+\includegraphics[scale=0.55]{diagramOne}
 
+%%Describe our implementation at the 50k ft view and work our way down until things are no longer interesting.
+In our implementation of MFTP, we defined a basic MFTP automaton, a sender and a receiver.
+The responsibilities of the MFTP automaton are to acquire[?] data and to help others acquire it.
+Each MFTP automaton begins with a file, encapsulated in a File object, which is either incomplete or complete.
+If the file is incomplete, the automaton generates requests for the missing fragments.
+If the file is complete, the automaton may only serve the file, or it may serve the file while performing matching on it.
+
+%%Picture of mFTP automaton?
+
+MFTP automata receive messages via the receive input action.  If the input message is a valid fragment of its own file, it saves the fragment.
+If the input message is a request for fragments of its file, it flags those fragments as having been requested.
+If the input message is a match containing its file ID, it records the matching IDs for download.
+
+The automaton holds a queue of one message to send. 
+The internal actions which fill this queue are triggered periodically.
+The automaton generates messages containing one requested fragment if some fragments in the file have been requested.
+The fragment sent is selected at random to help avoid collision with other MFTP automata serving the same file.
+The automaton generates requests if some fragments in the automaton's own file are missing.
+The automaton generates match messages if it knows about any matches to its file.
+
+The automaton may report that its file is complete using the download complete output action.
+
+\paragraph{Matching.}
+Each MFTP automaton contains one flag that determines whether it will perform matching and another that determines whether it will self-destruct when it has finished downloading its file.
+If the matching flag is set, the automaton applies a function $\iota$ to the file IDs of incoming fragments to determine whether it is a fragment of a file that could be a match.
+If $\iota$ is true, the automaton recursively constructs another MFTP automaton to download the potential match.
+This child automaton does not perform matching.  Its self-destruct flag is set so that it will self-destruct once the potential match has been downloaded.
+When the download is complete, the parent automaton applies $\mu$ to the file to determine whether it is a match.
+
+The sending automaton uses the UDP sender built into ioa to multicast messages one at a time.
+The receiving automaton uses the ioa UDP receiver.
+The sender and receiver automata may be connected to more than one MFTP automaton.
+
+\paragraph{Jam.} %%needs different header
+Our implementation defines three file types, \emph{file}, \emph{meta} and \emph{query}.  A file of file type contains data from any file in the network.
+A meta file contains as data the file ID of a file in the network and the name which is associated with it.
+A query file contains only the name of the file which is being queried.
+
+We wrote two programs, \emph{Share} and \emph{Get}, which use MFTP, sender and receiver automata to disseminate a file or files. %%Can one 'disseminate' a file?
+Share creates two MFTP automata, \emph{file server} and \emph{meta server}.
+File server is constructed with a complete file \emph{f}, and meta server is constructed with a meta file.
+The meta file is generated from the fileID and name of f.
+At the start, Get creates one MFTP automaton, \emph{query server}, constructed with a query file.
+The query file is generated from the name of the file to download, supplied by the command line argument.
+
+We defined a functor interface so that Share and Get could use different $\iota$ and $\mu$ functions.
+In Share, meta server performs matching on the meta file and broadcasts matches to f.
+Its $\iota$ returns true if the incoming file is a query file.
+Its $\mu$ returns true if the contents of the downloaded query file match the name of f.
+In Get, query server performs matching on the query file.
+Its $\iota$ returns true if the incoming file is a meta file.
+Its $\mu$ returns true if the contents of the downloaded meta file match the name of the file being queried.
+
+Query server passes the meta files downloaded by its child automata to Get.
+Using the file IDs contained in the meta files, Get constructs further MFTP automata which download the files pertaining to those file IDs.
+
+\vspace{10 mm}
+
+\includegraphics[scale=0.55]{share_diagram}
+
+\vspace{10 mm}
+
+\includegraphics[scale=0.65]{get_diagram}
+
+\vspace{10 mm}
 %%\begin{itemize}

### 2011-07-15 papers/paper1/paper.tex
diagram one
Finished up (for now) Implementation

@@
 
-\paragraph{I/O Automata}
-We used the ioa framework to develop an implementation of the MFTP protocol.  I/O Automata model stateful distributed system components that can interact and be composed with each other.
-They change state via transitions classified as input, output or internal actions.
-Input and output actions in different automata are bound; output actions in the one serve as interrupts to trigger input actions in the other.
-
-%%General overview of the framework and formal model.
-\includegraphics[scale=0.55]{diagramOne}
+\paragraph{MFTP Automaton.}
+Our implementation of MFTP uses the ioa framework.
+I/O Automata model stateful distributed system components that can interact and be composed with each other.
+They change state through transitions classified as input, output or internal actions.
+Input and output actions in different automata are bound; output actions in the one trigger input actions in the other.
 
 %%Describe our implementation at the 50k ft view and work our way down until things are no longer interesting.
-In our implementation of MFTP, we defined a basic MFTP automaton, a sender and a receiver.
-The responsibilities of the MFTP automaton are to acquire[?] data and to help others acquire it.
+We define a basic MFTP automaton, a sender, and a receiver.
+The responsibilities of the MFTP automaton are to acquire data and to help others acquire it.
 Each MFTP automaton begins with a file, encapsulated in a File object, which is either incomplete or complete.
 If the file is incomplete, the automaton generates requests for the missing fragments.
-If the file is complete, the automaton may only serve the file, or it may serve the file while performing matching on it.
+If the file is complete, the automaton may serve the file, or it may serve the file while performing matching on it.
 
@@
 
-MFTP automata receive messages via the receive input action.  If the input message is a valid fragment of its own file, it saves the fragment.
+MFTP automata receive messages via the receive input action.
+If the input message is a valid fragment of its own file, the automaton saves the fragment.
 If the input message is a request for fragments of its file, it flags those fragments as having been requested.
@@
 The internal actions which fill this queue are triggered periodically.
-The automaton generates messages containing one requested fragment if some fragments in the file have been requested.
-The fragment sent is selected at random to help avoid collision with other MFTP automata serving the same file.
+The automaton generates messages containing one requested fragment if any fragments in the file have been requested.
+The fragment sent is selected at random to help avoid collisions with other MFTP automata serving the same file.
 The automaton generates requests if some fragments in the automaton's own file are missing.
-The automaton generates match messages if it knows about any matches to its file.
+The automaton generates match messages if it has recorded any matches to its file.
 
-The automaton may report that its file is complete using the download complete output action.
+The automaton reports that its file is complete using the download complete output action.
 
 \paragraph{Matching.}
-Each MFTP automaton contains one flag that determines whether it will perform matching and another that determines whether it will self-destruct when it has finished downloading its file.
-If the matching flag is set, the automaton applies a function $\iota$ to the file IDs of incoming fragments to determine whether it is a fragment of a file that could be a match.
+Each MFTP automaton contains a flag that determines whether it will perform matching. 
+It has another flag that determines whether it will self-destruct when it has finished downloading its file.
+If the matching flag is set, the automaton applies the matching function $\iota$ to the file IDs of incoming fragments.
 If $\iota$ is true, the automaton recursively constructs another MFTP automaton to download the potential match.
-This child automaton does not perform matching.  Its self-destruct flag is set so that it will self-destruct once the potential match has been downloaded.
-When the download is complete, the parent automaton applies $\mu$ to the file to determine whether it is a match.
+This child automaton does not perform matching.  
+Its self-destruct flag is set so that it will self-destruct once the potential match has been downloaded.
+When the download is complete, the parent automaton applies the matching function $\mu$ to the file to determine whether it is a match.
 
+\paragraph{Some useful stuff.} %% We know it's rough, but what else can we do?
 The sending automaton uses the UDP sender built into ioa to multicast messages one at a time.
@@
 
+\vspace{5mm}
+\includegraphics[scale=0.65]{diagramOne}
+
 \paragraph{Jam.} %%needs different header
-Our implementation defines three file types, \emph{file}, \emph{meta} and \emph{query}.  A file of file type contains data from any file in the network.
-A meta file contains as data the file ID of a file in the network and the name which is associated with it.
+Our implementation defines three file types, \emph{data}, \emph{meta} and \emph{query}.  
+A data file contains data from any file in the network.
+A meta file contains as data the file ID of a file in the network and its name. 
 A query file contains only the name of the file which is being queried.
@@
 At the start, Get creates one MFTP automaton, \emph{query server}, constructed with a query file.
-The query file is generated from the name of the file to download, supplied by the command line argument.
+The query file is generated from the name of the file to download, supplied by a command line argument.
 
@@
 Its $\mu$ returns true if the contents of the downloaded query file match the name of f.
+
+\vspace{2 mm}
+
+\includegraphics[scale=0.55]{share_diagram}
+
+\vspace{2 mm}
+
 In Get, query server performs matching on the query file.
@@
 Its $\mu$ returns true if the contents of the downloaded meta file match the name of the file being queried.
-
 Query server passes the meta files downloaded by its child automata to Get.
@@
 
-\vspace{10 mm}
-
-\includegraphics[scale=0.55]{share_diagram}
-
-\vspace{10 mm}
+\vspace{2.5 mm}
 
@@
 
-\vspace{10 mm}
+\vspace{2 mm}
+
 %%\begin{itemize}

### 2011-07-18 papers/paper1/paper.tex
Justin's edits.

@@
 \usepackage{graphicx}
+\usepackage{outline}
 \begin{document}
 
-\title{Multicast File Transfer Protocol}
+\title{A protocol for implementing read-only databases in dynamic distributed environments}
 \author{Adam Drescher \and Magdalena Cassel}
@@
 \begin{abstract}
-In an effort to be portable, system developers often favor existing architectures, languages, operating systems, and standard libraries.
-The dominant abstraction in established computing technologies for concurrency and therefore distributed system development is threads.
-Consequently, distributed system developers are locked into using threads and experience the difficulties of thread-based development including obscure race conditions, poor support for the semantics of distributed systems, and difficulties integrating software developed using different thread models.
-Furthermore, the demand for portability and generality dismisses many of the techniques proposed in research that involve new programming languages, language extensions, or domain-specific solutions.
-This paper presents a development framework based on the I/O automata formal model implemented in C++.
-The I/O automata model is a proven technique for distributed system design because its semantics match those found in distributed systems.
-With the framework, developers can reason about correctness directly from the program text using the I/O automata formalism.
-The I/O  automata model also provides a common abstraction for concurrency which facilitates software integration and promotes reuse.
-TODO:  What are we doing for evaluation?
-%% Our ability to realize sophisticated distributed systems such as those in cyber-physical systems and pervasive computing is predicated on a solid model and implementation of asynchrony and concurrency.
-%% Threads, the de facto standard in practice, are not a suitable foundation because 1) ad hoc communication among threads prohibits software reuse, 2) threads are difficult to treat formally and therefore \emph{very} difficult to treat informally, and 3) threads are ill-suited for the asynchronous semantics present in the systems we desire to build.
-%% In this paper, we outline the semantics necessary for building sophisticated distributed systems and connect these to the I/O automata formal model.
-%% We then describe an implementation of the I/O automata formal model and cover extensions to the model necessary for operating in a dynamic environment.
-%% Finally, we conclude with an evaluation of our I/O automata framework.
+The abstract abstract.
 \end{abstract}
@@
 \section{Introduction}
-% jrw:  Move this to the introduction.
+
+Dynamic distributed systems are often governed by immutable or slowly changing configuration information contributed by each node in the system.
+A common problem is listing all of the participants
+
+For example, systems that use discovery and advertisement are governed by service specifications and descriptions contributed by service consumers and providers.
+%  A recurring problem in dynamic distributed systems is the collection and processing of configuration information.
+
+Both centralized and distributed algorithms have been proposed to collect and process configuration information.
+Centralized algorithms elect a leader which then facilitates the collection of configuration information and query processing.
+The leader might perform the processing itself or delegate processing to a subordinate.
+Distributed algorithms for configuration information collection and processing often use one of two techniques depending on the size of the records.
+If the records are bigger than the size of a message in the network, i.e., a UDP packet, then nodes discover other nodes using broadcast messages and then use point-to-point connections to collect and process the configuration information.
+If the records are smaller than the size of a message in the network, then the nodes exchange data using a multicast network, e.g., the Multicast Domain Name System.
+
+An unexplored area of the design space is techniques that support
+
+The problem of processing the configuration information can be cast using database techniques.
+Continuing the example, a service specification is a record in a table of service specifications and the process of discovery involves projecting the specification record on to the table of service descriptions.
+
+A recurring problem in dynamic distributed environments Read-only datab
+
+Dynamic distributed systems often require a protocol that allows a node to share public immutable data with all of the other nodes in the system.
+A typical example is discovery where service specifications are compared against service descriptions to match service providers with service consumers.
+Service descriptions and specifications are immutable even though they might exist at different times in the system since the set of nodes and therefore service providers and consumers may change with time.
+The immutability and public nature of the data implies that it can be freely replicated and therefore learned by gossiping.
+
+Often, the number of times the data is read is proportional to the number of nodes.
+
+There's a number of ways to do this.
+1.  Centralized approach where a leader becomes a data broker
+2.  Discover other nodes and then poll them.
+
+The main problem with the first approach is that its centralized.
+A problem with the second approach is that its inefficient (if you are n then you will be contacted n - 1 times).
+
+By dynamic, we mean that components come in and out.
+
+We assume a moderately sized ad hoc wireless network.
+
+We would like to take advantage of gossip.
+We would like to be fully distributed.
+
+Thus we developed a dissemination protocol for remote data exchange in wireless networks.
+
 Moving data across a network can take a variety of forms. 
@@
 
-Advances in embedded systems and networking are paving the way for distributed systems of unprecedented sophistication.
-Cyber-physical systems go beyond the traditional embedded system paradigm by explicitly modeling the dynamics of the physical system and communication network in the computational task.
-The inclusion of wireless network interfaces in modern electronic devices is transforming our homes, offices, and hospitals into pervasive computing environments.
-The interactions these systems have with the physical environment and its community of users require a degree of interoperability, configurability, adaptability, scalability, robustness, and security not found in existing systems.
-
-One of the forces that prevents the development of any kind of sophisticated software system is accidental complexity~\cite{brooks_nsb}.
-In the context of distributed systems, a major source of accidental complexity is a mismatch between the conceptual model of the system and the technology used to implement the system.
-The mismatch between the conceptual model and implementation method introduces defects due to translation errors, slows rapid prototyping, and retards iterative refinement.
-While there is no ``silver bullet''~\cite{brooks_nsb}, reducing the accidental complexity derived from the mismatch between concept and implementation will increase productivity and make new levels of sophistication possible with existing levels of effort.
-
-Threads are the de facto implementation vehicle for most of computing including distributed systems.
-Threads dominate because modern processors, programming languages, and operating systems were all developed to support the thread model.
-Furthermore, the widespread availability of thread libraries makes them a natural choice for system development.
-While threads dominate in practice, using threads to develop systems has a number of significant drawbacks.
-
-First, ad hoc communication among threads prohibits software reuse.
-To illustrate this problem consider the challenge of developing an application with a graphical user interface that also uses a multi-threaded middleware library.
-Graphical user interfaces often use a single-threaded event-based design, e.g., X, Swing.
-The system integrator must write glue code that bridges the single-threaded event-based semantics to multi-threaded semantics.
-Bridging between two thread models is achievable, however, the systems we desire to build will rely on many libraries each with their own thread model.
-The importance of a unified approach to concurrency becomes apparent when one considers bridging between different threading models, e.g., event-based, reactor, proactor, active objects, and their attributes, e.g., priority, thread-creation, blocking vs.
-non-blocking I/O.
-
-Second, threads are difficult to treat formally and therefore \emph{very} difficult to treat informally.
-The correctness of programs is reasoned about either formally as part of formal software design and verification and/or informally during the debugging process.
-Since most programs will never be formally verified, the correctness of most programs hinges solely on the developers ability to reason about the program.
-Based on the historical example of structured programming~\cite{goto_considered_harmful}, there is a strong correlation between what is ``easy'' in the formal arena and what is ``easy'' in the informal arena, especially when the implementation mimics the formal model, e.g., structured programming languages.
-It is well known that a formal treatment of threads is difficult due to the arbitrary interleaving of critical sections and ability to define critical sections of arbitrary scope~\cite{lee_threads}.
-Thus, the correctness of most multi-threaded programs hinges on the developer's ability to explore arbitrary interleavings of critical sections, i.e., model checking, in their head.
-
-Third, threads are ill-suited for the asynchronous semantics present in the systems we desire to build.
-The distributed systems we desire to build are asynchronous at both the device and network levels.
-Embedded devices must respond to asynchronous events in the environment.
-The interrupt-driven TinyOS operating system for wireless sensor network motes is an exemplar in this area.
-The fundamental communication primitive in real-world distributed systems is asynchronous message passing.
-Realizing asynchrony in a threaded program requires the use of signals (interrupts) and/or (a)synchronous I/O multiplexing.
-Signals are primitive, rigid, and difficult to use correctly.
-Synchronous I/O multiplexing, e.g., a select loop, is limited to operating system abstractions, e.g., file descriptors, and induces a reactive state machine structure that is difficult to understand and maintain.
-Asynchronous I/O improves efficiency by avoiding the dispatching necessary to determine the I/O operation but are subject to the same limitations of shared-state thread-based programming.
-
-Difficulties in developing asynchronous and concurrent software using threads has prompted a number of concurrent programming languages, e.g., Esterel, Erlang, Ptolemy, Autopipe, and language extensions, e.g., OpenMP, Cilk, Split-C.
-Solutions that involve creating a new language or extending an existing language suffer from two main problems.
-First, such solutions are often domain-specific and do not generalize to different applications, a necessity for heterogeneous distributed systems.
-Furthermore, solutions not connected to a general-purpose formal model lack the important property that the complete system including the environment can be modeled as another component.
-Second, such solutions often fail to become mainstream due to practical considerations.
-For example, the combination of C and pthreads is much more popular than Erlang due to the popularity of C-like languages and C's connection with UNIX operating system.
-Most language extensions are targeted at niche markets and therefore never standardized and implemented in mainstream compilers.
-
-Thus, in order to build sophisticated distributed systems, we require a common semantics and model for building concurrent and asynchronous systems with straight-forward abstractions.
-%% The model must be general to support the wide variety of devices and protocols necessary for cyber-physical and pervasive computing systems.
-%% The model must facilitate the development of concurrent modules that can be reused.
-%% The model must have a strong connection with a formal model to simplify the task of reasoning about individual components and complete systems.
-%% The model must be implemented in such a way that it is easily approachable.
-In section~\ref{system_model}, we expound upon the requirements of the model and connect the requirements with the I/O automata formal model.
-Section~\ref{representation} describes how I/O automata are represented in our I/O automata framework.
-In section~\ref{design}, we describe the design and implementation of our I/O automata framework.
-Section~\ref{evaluation} provides an evaluation and discussion of our framework.
-Section~\ref{related_work} contains an overview of related work and we offer our conclusions and thoughts for future work in section~\ref{conclusion}.
-
-%% We provide a common semantics and model for building concurrent and asynchronous systems with straight-forward abstractions.
-
-%%formal models, e.g., UNITY, I/O Automata,
-
-%% Devices are asynchronous and concurrent.
-%% Networks are asynchronous and concurrent.
-%% Why not start with a model that is asynchronous and concurrent?
-
-%% What inspires us:
-%% \begin{itemize}
-%%   \item CORBA - Remote Method Invocation (Remote Procedure Call with a ``this'' pointer).
-%%     ``Let's make synchronous function calls over a network.''
-%%     Assumes a thread of control.
-%%     Asynchronous calls were added later (Did these result in obfuscation?)
-
-%%   \item Patterns for asynchronous concurrency in a synchronous, thread-based world: Active objects, Reactor, Proactor, etc.
-%%     Examples: X server, OS2 Presentation Manager, Java Swing
-
-%%   \item AJAX - The modern Web is built on JavaScript + \emph{asynchronous} web page requests = AJAX.
-
-%%   \item Michi Henning - ICE
-
-%%   \item Steve Vinoski - Toolbox of programming languages
-
-%% \end{itemize}
-
-%% We provide a common semantics and model for building concurrent and asynchronous systems with straight-forward abstractions.
-
-\pagebreak
-
 \section{Model\label{model}}
@@
 The main concept in MFTP is that of a \emph{file}.
-Files are a sequence of bytes with a length and a user-defined type.
-They are broken down into equally sized pieces called \emph{fragments}.  
-Each file is uniquely identified by its \emph{file ID}, a structure that contains a hash of the file content, the file's length, and the file's type.
-Given two files $f_1$ and $f_2$, one can define a matching predicate $\mu$.
+A file is a sequence of bytes with a length and a user-defined type.
+Files are broken down into pieces of equal size called \emph{fragments}.
+Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's content.
+% TODO:  We need to move this.
+% Given two files $f_1$ and $f_2$, one can define a matching predicate $\mu$.
 

### 2011-07-18 papers/paper1/paper.tex
introduction and revisions to Model

@@
 \usepackage{graphicx}
-\usepackage{outline}
+
 \begin{document}
@@
 \section{Introduction}
+The proliferation of embedded networking is transforming homes into distributed computing environments.
+A variety of devices, including desktops, laptops, tablets, printers, digital picture frames, televisions, digital cameras, and optical disc players, are now equipped with wireless network interfaces.
+The ability of these devices to communicate with one another is creating opportunities for applications that engage multiple devices.
+Typically, these networks will consist of a small to moderate number of devices, many of which are resource-constrained.
+Home networks are built up as users contribute new devices to the network.
+When a new device is added, i.e., turned on, the user expects the device to be available in a timely fashion.
 
-Dynamic distributed systems are often governed by immutable or slowly changing configuration information contributed by each node in the system.
-A common problem is listing all of the participants
+Applications in home networks, and other dynamic distributed systems, are often governed by immutable configuration records contributed by each node in the system.
+A simple directory application might gather and display a list of all of the devices in the system.
+Similarly, applications using discovery and advertisement are governed by the service specifications and descriptions contributed by service consumers and providers.
+One of the tasks in the system, then, is to match specifications to descriptions for the purpose of service binding.
 
-For example, systems that use discovery and advertisement are governed by service specifications and descriptions contributed by service consumers and providers.
-%  A recurring problem in dynamic distributed systems is the collection and processing of configuration information.
+Both centralized and distributed algorithms have been proposed to collect and process configuration records.
+Centralized algorithms elect a leader which then facilitates query processing.
+The leader may perform the processing itself or delegate processing to a subordinate node.
+Distributed algorithms for configuration record processing often use one of two techniques, depending on the size of the records.
+If the records are bigger than the maximum size of a message, i.e., a UDP packet, nodes discover other nodes using broadcast messages and then use point-to-point connections to collect remote configuration records.
+Using this approach, the $n$th node to join the network will be queried $n-1$ times.
+If the records are smaller than the maximum size of a message, nodes exchange configuration records using a multicast network, e.g., the Multicast Domain Name System.
 
-Both centralized and distributed algorithms have been proposed to collect and process configuration information.
-Centralized algorithms elect a leader which then facilitates the collection of configuration information and query processing.
-The leader might perform the processing itself or delegate processing to a subordinate.
-Distributed algorithms for configuration information collection and processing often use one of two techniques depending on the size of the records.
-If the records are bigger than the size of a message in the network, i.e., a UDP packet, then nodes discover other nodes using broadcast messages and then use point-to-point connections to collect and process the configuration information.
-If the records are smaller than the size of a message in the network, then the nodes exchange data using a multicast network, e.g., the Multicast Domain Name System.
+To collect and process configuration records in home networks, we require a protocol that is both fully distributed and supports records of arbitrary size.
+The resource-poor nature of the devices in the network, combined with the dynamic nature of the network, implies that a fully distributed algorithm is more appropriate.
+In order to do capabilities-based matching, we require the ability to support configuration records of arbitrary size.
+Finally, we would like to use gossip to avoid sending duplicate records when a new node joins the network.
 
-An unexplored area of the design space is techniques that support
+This paper describes the Multicast File Transfer Protocol (MFTP), a distributed dissemination protocol for home networks that supports records of arbitrary size.
 
-The problem of processing the configuration information can be cast using database techniques.
-Continuing the example, a service specification is a record in a table of service specifications and the process of discovery involves projecting the specification record on to the table of service descriptions.
-
-A recurring problem in dynamic distributed environments Read-only datab
-
-Dynamic distributed systems often require a protocol that allows a node to share public immutable data with all of the other nodes in the system.
-A typical example is discovery where service specifications are compared against service descriptions to match service providers with service consumers.
-Service descriptions and specifications are immutable even though they might exist at different times in the system since the set of nodes and therefore service providers and consumers may change with time.
-The immutability and public nature of the data implies that it can be freely replicated and therefore learned by gossiping.
-
-Often, the number of times the data is read is proportional to the number of nodes.
-
-There's a number of ways to do this.
-1.  Centralized approach where a leader becomes a data broker
-2.  Discover other nodes and then poll them.
-
-The main problem with the first approach is that its centralized.
-A problem with the second approach is that its inefficient (if you are n then you will be contacted n - 1 times).
-
-By dynamic, we mean that components come in and out.
-
-We assume a moderately sized ad hoc wireless network.
-
-We would like to take advantage of gossip.
-We would like to be fully distributed.
-
-Thus we developed a dissemination protocol for remote data exchange in wireless networks.
-
-Moving data across a network can take a variety of forms. 
-These forms correspond to different user needs and consequently different protocols. 
-Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network. 
+%Moving data across a network can take a variety of forms. 
+%These forms correspond to different user needs and consequently different protocols. 
+%Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network. 
 %% This is useful for 1) device discovery 2) File transfer in the absence of many technologies (including internet).
@@
 
+We want to establish relationships between configuration records, which means we need to be able to download configuration records.
+
+%% The problem of processing configuration information can be cast using relational database techniques.
+%% A database consists of one or more tables, where each table consists of a number of identically structured records.
+%% Database tables are manipulated by the insert, delete and update methods, which insert, delete and modify records.
+%% Database tables are queried using the select operator.
+%% For processing configuration information, each configuration record is mapped into a table of the appropriate type.
+%% Listing all of the components in the system corresponds to selecting all of the records in the database.
+%% Discovery can be accomplished by projecting a service specification record onto the table of service descriptions.
+
+%% A record is defined by its content.
+%% Consequently, each record is immutable and the update method is not allowed.
+%% Configuration records are immutable and under the control of the node that stores them.
+
+%% Each record is immutable. Update doesn't exist. The records stored on each node are under the control of that node. Insert and delete only exist locally.
+
+%% Each table and each record has an ID
+%% Each record ID consists of the ID of the table it belongs to and the length and hash of its binary representation.
+%% The records are uniquely identifiable by their IDs.
+%% Each node then contributes zero or more immutable records to one or more of the tables.
+%% We can consider each configuration item as a record in a read-only database.
+
+%% Each record is locally controlled
+%% since immutable, can be replicated (no duplicates in the database)
+%% each node has its own set of records which it contributes
+
+Files are an abstraction of the configuration record.
+
 \paragraph {Files.}
 The main concept in MFTP is that of a \emph{file}.
+
 A file is a sequence of bytes with a length and a user-defined type.
-Files are broken down into pieces of equal size called \emph{fragments}.
-Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's content.
+Files are divided into pieces of equal size called \emph{fragments}.
+Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's contents.
+
 % TODO:  We need to move this.
@@
 
-\paragraph{Messages.}
-File transfer takes place via \emph{messages}.
-There are three types of messages in the protocol.
-The first message type is a \emph{request}.
-A request contains a file ID and a set of fragment indices.
-We implemented requests as the file ID and a set of contiguous regions.  
-However, one can envision the request to contain anything that defines a set.
-The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
-Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.
-
 \paragraph {Tasks.}
+The two main tasks in MFTP are acquiring a file and establishing a relationship between two files.
+The process of acquiring a file involves sending a request for the file's fragments and receiving fragments sent in response.
 MFTP is a peer-to-peer protocol because each participant can act as both a client and a server.
@@
 However, if $\iota$($fid_1$,$fid_2$) is true, $\mu$($f_1$,$f_2$) does \emph{not} have to be true.
+
+\paragraph{Messages.}
+File transfer takes place via \emph{messages}.
+There are three types of messages in the protocol.
+The first message type is a \emph{request}.
+A request contains a file ID and a set of fragment indices.
+We implemented requests as the file ID and a set of contiguous regions.  
+However, one can envision the request to contain anything that defines a set.
+The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
+Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.
 

### 2011-07-18 include/mftp/mftp_automaton.hpp
fixed bug suddenly and randomly thought of

@@ -266,14 +266,14 @@ namespace mftp {
       case FRAGMENT:
 	{
 	  // If we are looking for our own file, it must be a fragment from our file and the offset must be correct.
-	  if (!m_file.complete () &&
-	      m->frag.fid == m_fileid &&
+	  if (m->frag.fid == m_fileid &&
 	      m->frag.offset < m_mfileid.get_final_length () &&
 	      m->frag.offset % FRAGMENT_SIZE == 0) {
 	    
 	    // Save the fragment.
-	    m_file.write_chunk (m->frag.offset, m->frag.data);
-	    
+	    if (!m_file.complete ()){
+	      m_file.write_chunk (m->frag.offset, m->frag.data);
+	    }
 	    uint32_t idx = (m->frag.offset / FRAGMENT_SIZE);
 	    if (m_requests[idx]) {
 	      // Somebody else wanted it and now has just seen it.

### 2011-07-18 papers/paper1/paper.tex
Workin' on the paper.

@@
 
-\title{A protocol for implementing read-only databases in dynamic distributed environments}
+\title{A gossip-based protocol for exchanging immutable data in dynamic distributed environments}
 \author{Adam Drescher \and Magdalena Cassel}
@@
 
-We want to establish relationships between configuration records, which means we need to be able to download configuration records.
+The two main tasks in configuration record processing are acquiring records from the network and establishing a relationship between two records.
+We adopt the view that configuration records can be serialized and stored in a file.
+The task of acquiring a record is then equivalent to downloading a file from the network.
+Once a configuration record has been downloaded, its content can then be compared to configuration records that are stored locally.
+
+A \emph{file} in MFTP is a finite sequence of bytes with an associated type.
+Each file is uniquely identified by its \emph{file ID} which contains the file's type, length, and a hash of the file's contents.
+The type is defined by the user and is used to limit the potential relationships among files.
+We assume the existence of a user-defined predicate $\gamma(fid_1, fid_2)$ that returns true if there is potentially a relationship between the two files.
+The predicate $\gamma$ is symmetric, i.e., $\gamma(fid_1, fid_2) = \gamma(fid_2, fid_1)$ and non-reflexive, i.e., $\gamma(fid_1, fid_1)$ is false.
+Note that $\gamma$ is only defined using file IDs and can therefore only check file types and lengths.
+We also assume the existence of a user-defined predicate $\mu(file_1, file_2)$ that returns true if a relationship between $file_1$ and $file_2$ exists.
+Like $\gamma$, the predicate $\mu$ is symmetric and non-reflexive.
+The predicate $\mu$ implies $\gamma$, i.e., $\mu(file_1, file_2) \to \gamma(fid(file_1), fid(file_2))$.
+
+MFTP assumes that all communication takes place by sending fixed-size messages over a broadcast or multicast channel.
+This allows MFTP to take advantage of gossip since file contents and matching relationships can be communicated efficiently to multiple interested parties.
+To enable the sending of file contents using fixed-size messaging, we divide a file into pieces of equal size called \emph{fragments}.
+
+The MFTP protocol contains three types of messages: fragments, requests, and matches.
+Each message in MFTP is prefixed with a header indicating its type.
+A \emph{fragment message} consists of a file ID, a fragment index, and the data comprising the indicated fragment.
+A \emph{request message} consists of a file ID and a set of fragment indices.
+For example, our implementation uses a set of spans where each span requests a contiguous set of fragments.
+However, one can envision the request to contain more advanced structures, e.g., functions, that can define arbitrary sets.
+A \emph{match message} consist of a file ID and a set of matching file IDs for which $\gamma$ is true.
+
+Two files with the same content and type will have the same file ID.
+With a strong enough hashing mechanism, we can assume two files are equal if their file IDs are the same.
+Since the identity of a file is completely determined by its content, files are immutable and can be freely replicated.
+Peers in the protocol are free to cache any fragment and respond to any requests for which they have a fragment.
+The process of acquiring a file consists of repeatedly sending requests and receiving fragments from any peers that chose to respond until all of the fragments have been acquired.
+
+Learning relationships can be accomplished in one of two ways.
+First, peers in MFTP periodically send a fragment to announce the presence of the file in the network.
+Assume that a peer serving $file_1$ receives an announcement for $file_2$.
+The peer checks $\gamma(fid(file_1), fid(file_2))$ to determine whether a relationship is possible.
+If a relationship is possible, the peer acquires the file and checks $\mu(file_1, file_2)$ to determine whether the relationship exists.
+Peers periodically send match messages to indicate which files have a relationship with the file they serve.
+The second way relationships can be learned is by eavesdropping on the match messages sent by other peers.
+
+%% \paragraph {Files.}
+%% The main concept in MFTP is that of a \emph{file}.
+%% A file is a sequence of bytes with a length and a user-defined type.
+%% Files are divided into pieces of equal size called \emph{fragments}.
+%% Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's contents.
+
+%% \paragraph{Messages.}
+%% File transfer takes place via \emph{messages}.
+%% There are three types of messages in the protocol.
+%% The first message type is a \emph{request}.
+%% A request contains a file ID and a set of fragment indices.
+%% We implemented requests as the file ID and a set of contiguous regions.  
+%% However, one can envision the request to contain anything that defines a set.
+%% The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
+%% Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.
+
+%% \paragraph {Tasks.}
+%% The two main tasks in MFTP are acquiring a file and establishing a relationship between two files.
+%% The process of acquiring a file involves sending a request for the file's fragments and receiving fragments sent in response.
+%% MFTP is a peer-to-peer protocol because each participant can act as both a client and a server.
+%% A peer can send out any fragments for which it gets a request.
+%% Fragments are periodically broadcast, even in the absence of requests, to indicate what files exist in the network.
+%% A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
+%% A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.
+
+%% Matching is a two-step process.
+%% The first step applies a matching function $\iota$ to two file IDs to determine whether a match is possible between the two files.  
+%% The second step applies a matching function $\mu$ to the two files to determine whether there is a match.
+%% If $\mu$($f_1$,$f_2$) is true, $\iota$($fid_1$,$fid_2$) must be true.
+%% However, if $\iota$($fid_1$,$fid_2$) is true, $\mu$($f_1$,$f_2$) does \emph{not} have to be true.
 
@@
 
-Files are an abstraction of the configuration record.
-
-\paragraph {Files.}
-The main concept in MFTP is that of a \emph{file}.
-
-A file is a sequence of bytes with a length and a user-defined type.
-Files are divided into pieces of equal size called \emph{fragments}.
-Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's contents.
-
-% TODO:  We need to move this.
-% Given two files $f_1$ and $f_2$, one can define a matching predicate $\mu$.
-
-\paragraph {Tasks.}
-The two main tasks in MFTP are acquiring a file and establishing a relationship between two files.
-The process of acquiring a file involves sending a request for the file's fragments and receiving fragments sent in response.
-MFTP is a peer-to-peer protocol because each participant can act as both a client and a server.
-A peer can send out any fragments for which it gets a request.
-Fragments are periodically broadcast, even in the absence of requests, to indicate what files exist in the network.
-A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
-A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.
-
-Matching is a two-step process.
-The first step applies a matching function $\iota$ to two file IDs to determine whether a match is possible between the two files.  
-The second step applies a matching function $\mu$ to the two files to determine whether there is a match.
-If $\mu$($f_1$,$f_2$) is true, $\iota$($fid_1$,$fid_2$) must be true.
-However, if $\iota$($fid_1$,$fid_2$) is true, $\mu$($f_1$,$f_2$) does \emph{not} have to be true.
-
-\paragraph{Messages.}
-File transfer takes place via \emph{messages}.
-There are three types of messages in the protocol.
-The first message type is a \emph{request}.
-A request contains a file ID and a set of fragment indices.
-We implemented requests as the file ID and a set of contiguous regions.  
-However, one can envision the request to contain anything that defines a set.
-The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
-Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.
+\section{Implementation}
 
 %%TODO: Talk about validation
-
-\section{Implementation}
 

### 2011-07-18 include/mftp/mftp_sender_automaton.hpp

@@ -41,7 +41,7 @@ namespace mftp {
       add_observable (&send);
       add_observable (&send_complete);
 
-      ioa::automaton_manager<ioa::udp_sender_automaton>* sender = new ioa::automaton_manager<ioa::udp_sender_automaton> (this, ioa::make_generator<ioa::udp_sender_automaton> ());
+      ioa::automaton_manager<ioa::udp_sender_automaton>* sender = new ioa::automaton_manager<ioa::udp_sender_automaton> (this, ioa::make_generator<ioa::udp_sender_automaton> (sizeof (message)));
 
       ioa::make_binding_manager (this,
 				 &m_self, &mftp_sender_automaton::send_out,

### 2011-07-18 src/get.cpp

@@ -62,7 +62,7 @@ namespace jam {
       memcpy (&fid, f->get_data_ptr (), sizeof (mftp::fileid));
       fid.convert_to_host();
       
-      ioa::automaton_manager<mftp::mftp_automaton>* file_home = new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (mftp::file (fid), sender->get_handle(), receiver->get_handle(), true));  //TODO: CHANGE THIS BACK
+      ioa::automaton_manager<mftp::mftp_automaton>* file_home = new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (mftp::file (fid), sender->get_handle(), receiver->get_handle(), false));
       
       ioa::make_binding_manager (this,
 				 file_home, &mftp::mftp_automaton::download_complete,

### 2011-07-19 include/mftp/mftp_automaton.hpp
Modifications to make the protocol more responsive. The automaton now sends an announcement or request at the start. A match is sent out immediately when generated. First matches are responsive but subsequent matches still require the match timeout.

@@ -35,7 +35,6 @@ namespace mftp {
     public ioa::automaton
   {
   private:
-    static const ioa::time FRAGMENT_INTERVAL;
     static const ioa::time REQUEST_INTERVAL;
     static const ioa::time INIT_ANNOUNCEMENT_INTERVAL;
     static const ioa::time MAX_ANNOUNCEMENT_INTERVAL;
@@ -59,7 +58,9 @@ namespace mftp {
     uint32_t m_requests_count; // Number of true elements in m_requests.
     std::queue<ioa::const_shared_ptr<message_buffer> > m_sendq;
     send_state_t m_send_state;
-    timer_state_t m_fragment_timer_state;
+    bool m_send_request;
+    bool m_send_announcement;
+    bool m_send_match;
     timer_state_t m_request_timer_state;
     timer_state_t m_announcement_timer_state;
     timer_state_t m_matching_timer_state;
@@ -96,19 +97,6 @@ namespace mftp {
 				 &m_receiver, &mftp_receiver_automaton::receive,
 				 &m_self, &mftp_automaton::receive);
 
-      ioa::automaton_manager<ioa::alarm_automaton>* fragment_clock = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
-      ioa::make_binding_manager (this,
-			     &m_self,
-			     &mftp_automaton::set_fragment_timer,
-			     fragment_clock,
-			     &ioa::alarm_automaton::set);
-      ioa::make_binding_manager (this,
-			     fragment_clock,
-			     &ioa::alarm_automaton::alarm,
-			     &m_self,
-			     &mftp_automaton::fragment_timer_interrupt);
-
-    
       ioa::automaton_manager<ioa::alarm_automaton>* request_clock = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
       ioa::make_binding_manager (this,
 			     &m_self,
@@ -163,7 +151,9 @@ namespace mftp {
       m_requests (m_mfileid.get_fragment_count ()),
       m_requests_count (0),
       m_send_state (SEND_READY),
-      m_fragment_timer_state (SET_READY),
+      m_send_request (!m_file.complete ()),
+      m_send_announcement (!m_file.empty ()),
+      m_send_match (false),
       m_request_timer_state (SET_READY),
       m_announcement_timer_state (SET_READY),
       m_matching_timer_state (SET_READY),
[…]

### 2011-07-19 include/mftp/mftp.hpp
Changed order of fields when converting fileid to string.

@@ -61,12 +61,16 @@ namespace mftp {
     }
 
     std::string to_string () const {
-      char temp[2 * sizeof (fileid) + 1];
+      char buffer[2 * sizeof (fileid) + 1];
+      char* ptr = buffer;
+      const char* end = ptr + sizeof (buffer);
+
+      ptr += snprintf (ptr, end - ptr, "%08x%08x", type, length);
+      
       for (size_t idx = 0; idx < HASH_SIZE; ++idx) {
-      	sprintf (&temp[2 * idx], "%02x", hash[idx]);
+      	ptr += snprintf (ptr, end - ptr, "%02x", hash[idx]);
       }
-      sprintf (temp + 2 * HASH_SIZE, "%08x%08x", type, length);
-      return std::string (temp, 2 * sizeof (fileid));
+      return std::string (buffer, ptr - buffer);
     }
   };
 

### 2011-07-19 lib/file.cpp
Created sha2_256.cpp.

@@ -1,5 +1,5 @@
 #include <mftp/file.hpp>
-#include <mftp/sha2_256.hpp>
+#include "sha2_256.hpp"
 
 #include <math.h>
 #include <cstdlib>

### 2011-07-19 lib/sha2_256.cpp

@@ -1,0 +1,173 @@
+#include "sha2_256.hpp"
+
+const unsigned int sha2_256::h_init[8] = {
+  0x6a09e667, 0xbb67ae85, 0x3c6ef372, 0xa54ff53a, 0x510e527f, 0x9b05688c, 0x1f83d9ab, 0x5be0cd19
+};
+
+const unsigned int sha2_256::k[64] = {
+  0x428a2f98, 0x71374491, 0xb5c0fbcf, 0xe9b5dba5, 0x3956c25b, 0x59f111f1, 0x923f82a4, 0xab1c5ed5,
+  0xd807aa98, 0x12835b01, 0x243185be, 0x550c7dc3, 0x72be5d74, 0x80deb1fe, 0x9bdc06a7, 0xc19bf174,
+  0xe49b69c1, 0xefbe4786, 0x0fc19dc6, 0x240ca1cc, 0x2de92c6f, 0x4a7484aa, 0x5cb0a9dc, 0x76f988da,
+  0x983e5152, 0xa831c66d, 0xb00327c8, 0xbf597fc7, 0xc6e00bf3, 0xd5a79147, 0x06ca6351, 0x14292967,
+  0x27b70a85, 0x2e1b2138, 0x4d2c6dfc, 0x53380d13, 0x650a7354, 0x766a0abb, 0x81c2c92e, 0x92722c85,
+  0xa2bfe8a1, 0xa81a664b, 0xc24b8b70, 0xc76c51a3, 0xd192e819, 0xd6990624, 0xf40e3585, 0x106aa070,
+  0x19a4c116, 0x1e376c08, 0x2748774c, 0x34b0bcb5, 0x391c0cb3, 0x4ed8aa4a, 0x5b9cca4f, 0x682e6ff3,
+  0x748f82ee, 0x78a5636f, 0x84c87814, 0x8cc70208, 0x90befffa, 0xa4506ceb, 0xbef9a3f7, 0xc67178f2
+};
+
+unsigned int sha2_256::right_rotate (unsigned int x,
+				     unsigned int s) {
+  return (x >> s) | (x << (32 - s));
+}
+
+void sha2_256::process () {
+  assert (buf_length == 64);
+
+  // Break chunk into 16 4-byte variables.
+  unsigned int w[64];
+  for (unsigned int idx = 0; idx < 16; ++idx) {
+    memcpy (&w[idx], &buf[idx * 4], 4);
+    w[idx] = ntohl (w[idx]);
+  }
+
[…]

### 2011-07-19 lib/sha2_256.hpp

@@ -1,0 +1,37 @@
+#ifndef __sha2_256_hpp__
+#define __sha2_256_hpp__
+
+/*
+  Implementation of the SHA2-256 digest algorithm.
+  Based on pseuedo-code found at http://en.wikipedia.org/wiki/SHA-2
+ */
+
+#include <algorithm>
+#include <arpa/inet.h>
+#include <cassert>
+#include <cstring>
+
+class sha2_256 {
+private:
+  static const unsigned int h_init[8];
+  static const unsigned int k[64];
+  unsigned int h[8];
+  unsigned int total_length;
+  unsigned int buf_length;
+  unsigned char buf[64];
+
+  static unsigned int right_rotate (unsigned int x,
+				    unsigned int s);
+  void process ();
+
+public:
+  sha2_256 ();
+  sha2_256 (const unsigned int length,
+	    const unsigned char* hash);
+  void finalize ();
+  void get (unsigned char* hash) const;
+  void update (const unsigned char* data,
+	       size_t length);
+};
+
+#endif

### 2011-07-19 include/mftp/file.hpp
Refactored mftp.hpp.

@@ -1,9 +1,9 @@
 #ifndef __file_hpp__
 #define __file_hpp__
 
-#include <mftp/mftp.hpp>
-
 #include <vector>
+#include <mftp/fileid.hpp>
+#include <mftp/span.hpp>
 
 namespace mftp {
   
@@ -19,8 +19,10 @@ namespace mftp {
 
     uint32_t offset_to_hash (const uint32_t) const;
     uint32_t hash_to_offset (const uint32_t) const;
-    bool get_previous_hash (const uint32_t, unsigned char*) const;
-    bool get_hash (const uint32_t, unsigned char*) const;
+    bool get_previous_hash (const uint32_t,
+			    unsigned char*) const;
+    bool get_hash (const uint32_t,
+		   unsigned char*) const;
     void validate (const uint32_t);
 
   public:
@@ -28,7 +30,9 @@ namespace mftp {
 	  const uint32_t);
     file (const fileid& f);
     file (const file& other);
-    file (const void* ptr, uint32_t size, uint32_t type);
+    file (const void* ptr,
+	  uint32_t size,
+	  uint32_t type);
     ~file ();
     
     const mfileid& get_mfileid () const;

### 2011-07-19 include/mftp/fileid.hpp

@@ -1,0 +1,165 @@
+#ifndef __fileid_hpp__
+#define __fileid_hpp__
+
+#include <stdint.h>
+#include <cstddef>
+#include <string>
+#include <cmath>
+
+namespace mftp {
+  const size_t HASH_SIZE = 32;
+  const size_t FRAGMENT_SIZE = 512;
+
+  struct fileid
+  {
+    uint32_t type;
+    uint32_t length;
+    uint8_t hash[HASH_SIZE];
+
+    bool operator== (const fileid& other) const {
+      return type == other.type &&
+	length == other.length &&
+	memcmp (hash, other.hash, HASH_SIZE) == 0;
+    }
+
+    bool operator!= (const fileid& other) const {
+      return type != other.type &&
+	length != other.length &&
+	memcmp (hash, other.hash, HASH_SIZE) != 0;
+    }
+
+    bool operator< (const fileid& other) const {
+      if (type != other.type) {
+	return type < other.type;
+      }
+      if (length != other.length) {
+	return length < other.length;
+      }
+      return memcmp (hash, other.hash, HASH_SIZE) < 0;
+    }
+
+    void convert_to_network () {
+      type = htonl (type);
+      length = htonl (length);
+    }
+
+    void convert_to_host () {
+      type = ntohl (type);
+      length = ntohl (length);
+    }
+
+    std::string to_string () const {
+      char buffer[2 * sizeof (fileid) + 1];
+      char* ptr = buffer;
+      const char* end = ptr + sizeof (buffer);
+
+      ptr += snprintf (ptr, end - ptr, "%08x%08x", type, length);
+      
+      for (size_t idx = 0; idx < HASH_SIZE; ++idx) {
+      	ptr += snprintf (ptr, end - ptr, "%02x", hash[idx]);
+      }
+      return std::string (buffer, ptr - buffer);
+    }
+  };
+
+  // Memoized fileid.
+  class mfileid
+  {
+  private:
+    uint32_t m_fragment_count;
+    uint32_t m_padded_length;
+    uint32_t m_final_length;
+    fileid m_fileid;
+
+    void calculate_lengths () {
+      // LET'S DO MATH!!
+
+      // Calculate the size after we pad with 0s to fall on a HASH_SIZE boundary.
+      uint32_t padded_length = m_fileid.length + HASH_SIZE - (m_fileid.length % HASH_SIZE);
+      assert ((padded_length % HASH_SIZE) == 0);
+
+      /*
[…]

### 2011-07-19 include/mftp/message.hpp

@@ -1,0 +1,244 @@
+#ifndef __message_hpp__
+#define __message_hpp__
+
+#include <mftp/fileid.hpp>
+#include <mftp/span.hpp>
+#include <ioa/buffer.hpp>
+
+namespace mftp {
+  const uint32_t REQUEST = 0;
+  const uint32_t FRAGMENT = 1;
+  const uint32_t MATCH = 2;
+
+  const uint32_t SPANS_SIZE = 64;
+  const uint32_t MATCHES_SIZE = 12;
+
+  struct fragment
+  {
+    fileid fid;
+    uint32_t offset;
+    uint8_t data[FRAGMENT_SIZE];
+
+    void convert_to_network () {
+      fid.convert_to_network ();
+      offset = htonl (offset);
+    }
+
+    bool convert_to_host () {
+      fid.convert_to_host ();
+      offset = ntohl (offset);
+
+      mfileid mid (fid);
+      return ((offset % FRAGMENT_SIZE) == 0) && (offset < mid.get_final_length ());
+    }
+  };
+    
+  struct request
+  {
+    fileid fid;
+    uint32_t span_count;
+    span_t spans[SPANS_SIZE];
+    
+    void convert_to_network () {
+      fid.convert_to_network ();
+      for (uint32_t i = 0; i < span_count; ++i) {
+	spans[i].convert_to_network ();
+      }
+      span_count = htonl (span_count);
+    }
+    
+    bool convert_to_host () {
+      fid.convert_to_host ();
+      span_count = ntohl (span_count);
+      
+      if (span_count == 0 || span_count > SPANS_SIZE) {
+	return false;
+      }
+      
+      mfileid mid (fid);
+      for (uint32_t i = 0; i < span_count; ++i) {
+	spans[i].convert_to_host ();
+	
+	if (!((spans[i].start < spans[i].stop) &&
+	      (spans[i].stop <= mid.get_final_length () ) &&
+	      (spans[i].start % FRAGMENT_SIZE == 0) &&
+	      (spans[i].stop % FRAGMENT_SIZE == 0) )) {
+	  return false;
+	}
+      }
+      return true;
+    }
+  };
+
+  struct match
+  {
+    fileid fid;
+    uint32_t match_count;
+    fileid matches[MATCHES_SIZE];
+
+    void convert_to_network () {
+      fid.convert_to_network ();
+      for (uint32_t i = 0; i < match_count; ++i) {
+	matches[i].convert_to_network ();
+      }
[…]

### 2011-07-19 include/mftp/mftp.hpp

@@ -1,419 +1,6 @@
 #ifndef __mftp_hpp__
 #define __mftp_hpp__
 
-#include <stdint.h>
-#include <cstring>
-#include <cassert>
-#include <math.h>
-#include <utility>
-#include <arpa/inet.h>
-#include <ioa/buffer.hpp>
-#include <iostream>
-#include <stdio.h>
-
-namespace mftp {
-  const size_t HASH_SIZE = 32;
-  const size_t FRAGMENT_SIZE = 512;
-
-  const uint32_t REQUEST = 0;
-  const uint32_t FRAGMENT = 1;
-  const uint32_t MATCH = 2;
-
-  const uint32_t SPANS_SIZE = 64;
-  const uint32_t MATCHES_SIZE = 12;
-
-  struct fileid
-  {
-    uint32_t type;
-    uint32_t length;
-    uint8_t hash[HASH_SIZE];
-
-    bool operator== (const fileid& other) const {
-      return type == other.type &&
-	length == other.length &&
-	memcmp (hash, other.hash, HASH_SIZE) == 0;
-    }
-
-    bool operator!= (const fileid& other) const {
-      return type != other.type &&
-	length != other.length &&
-	memcmp (hash, other.hash, HASH_SIZE) != 0;
-    }
-
-    bool operator< (const fileid& other) const {
-      if (type != other.type) {
-	return type < other.type;
-      }
-      if (length != other.length) {
-	return length < other.length;
-      }
-      return memcmp (hash, other.hash, HASH_SIZE) < 0;
-    }
-
-    void convert_to_network () {
-      type = htonl (type);
-      length = htonl (length);
-    }
-
-    void convert_to_host () {
-      type = ntohl (type);
-      length = ntohl (length);
-    }
-
-    std::string to_string () const {
-      char buffer[2 * sizeof (fileid) + 1];
-      char* ptr = buffer;
-      const char* end = ptr + sizeof (buffer);
-
-      ptr += snprintf (ptr, end - ptr, "%08x%08x", type, length);
-      
-      for (size_t idx = 0; idx < HASH_SIZE; ++idx) {
-      	ptr += snprintf (ptr, end - ptr, "%02x", hash[idx]);
-      }
-      return std::string (buffer, ptr - buffer);
-    }
-  };
-
-  // Memoized fileid.
-  class mfileid
-  {
-  private:
-    uint32_t m_fragment_count;
-    uint32_t m_padded_length;
[…]

### 2011-07-19 include/mftp/mftp_automaton.hpp

@@ -438,6 +438,7 @@ namespace mftp {
     }
 
     void send_request_effect () {
+      
       if (!m_file.complete ()) {
 	span_t spans[SPANS_SIZE];
 	spans[0] = m_file.get_next_range();

### 2011-07-19 include/mftp/mftp_sender_automaton.hpp

@@ -2,8 +2,10 @@
 #define __mftp_sender_automaton_hpp__
 
 #include <ioa/udp_sender_automaton.hpp>
+#include <mftp/message.hpp>
 
 #include <queue>
+#include <iostream>
 
 namespace mftp {
 

### 2011-07-19 include/mftp/span.hpp

@@ -1,0 +1,25 @@
+#ifndef __span_hpp__
+#define __span_hpp__
+
+namespace mftp {
+
+  struct span_t
+  {
+    uint32_t start;
+    uint32_t stop;
+
+    void convert_to_network () {
+      start = htonl (start);
+      stop = htonl (stop);
+    }
+    
+    void convert_to_host () {
+      start = ntohl (start);
+      stop = ntohl (stop);
+    }
+
+  };
+
+}
+
+#endif

### 2011-07-19 lib/file.cpp

@@ -224,12 +224,12 @@ namespace mftp {
       ;;
       
     // Request range [m_start_idx, end_idx).
-    span_t retspan;
-    retspan.start = m_start_idx * FRAGMENT_SIZE;
-    retspan.stop = end_idx * FRAGMENT_SIZE;
+    span_t retval;
+    retval.start = m_start_idx * FRAGMENT_SIZE;
+    retval.stop = end_idx * FRAGMENT_SIZE;
     m_start_idx = end_idx;
 
-    return retspan;
+    return retval;
   }
 
   uint32_t file::get_random_index () const {

### 2011-07-19 src/jam.hpp

@@ -1,7 +1,7 @@
 #ifndef __jam_hpp__
 #define __jam_hpp__
 
-#include <mftp/mftp_automaton.hpp>
+#include <mftp/mftp.hpp>
 
 namespace jam {
   

### 2011-07-19 include/mftp/fileid.hpp
Fixed mac magic include errors

@@ -1,10 +1,14 @@
 #ifndef __fileid_hpp__
 #define __fileid_hpp__
 
-#include <stdint.h>
-#include <cstddef>
-#include <string>
+#include <arpa/inet.h>
+#include <cassert>
 #include <cmath>
+#include <cstddef>
+#include <cstring>
+#include <stdint.h>
+#include <stdio.h>
+#include <string>
 
 namespace mftp {
   const size_t HASH_SIZE = 32;

### 2011-07-19 include/mftp/file.hpp
Send queue can always contain one fragment.

@@ -38,6 +38,7 @@ namespace mftp {
     const mfileid& get_mfileid () const;
     void* get_data_ptr ();
     const void* get_data_ptr () const;
+    uint32_t have_count () const;
     bool complete () const;
     bool empty () const;
     bool have (const uint32_t offset) const;

### 2011-07-19 include/mftp/mftp_automaton.hpp

@@ -13,7 +13,6 @@
 #include <set>
 #include <vector>
 #include <cstring>
-#include <math.h>
 
 // TODO:  We accumulate matches without ever forgetting them.  We should purge matches that we haven't seen in a while.
 
@@ -39,6 +38,7 @@ namespace mftp {
     static const ioa::time INIT_ANNOUNCEMENT_INTERVAL;
     static const ioa::time MAX_ANNOUNCEMENT_INTERVAL;
     static const ioa::time MATCHING_INTERVAL;
+    static const size_t MAX_FRAGMENT_COUNT;
 
     enum send_state_t {
       SEND_READY,
@@ -58,6 +58,7 @@ namespace mftp {
     uint32_t m_requests_count; // Number of true elements in m_requests.
     std::queue<ioa::const_shared_ptr<message_buffer> > m_sendq;
     send_state_t m_send_state;
+    size_t m_fragment_count;
     bool m_send_request;
     bool m_send_announcement;
     bool m_send_match;
@@ -151,6 +152,7 @@ namespace mftp {
       m_requests (m_mfileid.get_fragment_count ()),
       m_requests_count (0),
       m_send_state (SEND_READY),
+      m_fragment_count (0),
       m_send_request (!m_file.complete ()),
       m_send_announcement (!m_file.empty ()),
       m_send_match (false),
@@ -183,6 +185,7 @@ namespace mftp {
       m_requests (m_mfileid.get_fragment_count ()),
       m_requests_count (0),
       m_send_state (SEND_READY),
+      m_fragment_count (0),
       m_send_request (!m_file.complete ()),
       m_send_announcement (!m_file.empty ()),
       m_send_match (false),
@@ -246,6 +249,10 @@ namespace mftp {
     ioa::const_shared_ptr<message_buffer> send_effect () {
       ioa::const_shared_ptr<message_buffer> m = m_sendq.front ();
       m_sendq.pop ();
+      if (m->msg.header.message_type == htonl (FRAGMENT)) {
+	--m_fragment_count;
+      }
+
       m_send_state = SEND_COMPLETE_WAIT;
       return m;
     }
@@ -415,7 +422,7 @@ namespace mftp {
 
   private:
     bool send_fragment_precondition () const {
-      return m_requests_count != 0 && m_sendq.empty ();
+      return m_requests_count != 0 && m_fragment_count < MAX_FRAGMENT_COUNT;
[…]

### 2011-07-19 lib/file.cpp

@@ -174,6 +174,10 @@ namespace mftp {
     return m_data;
   }
 
+  uint32_t file::have_count () const {
+    return m_have_count;
+  }
+
   bool file::complete () const {
     return m_valid_count == m_mfileid.get_fragment_count ();
   }

### 2011-07-19 include/mftp/file.hpp
adding svgs for get and share diagrams

@@ -45,6 +45,7 @@ namespace mftp {
 		      const void* data);
     span_t get_next_range ();
     uint32_t get_random_index () const;
+    uint32_t get_progress () const;
   };
 }
 

### 2011-07-19 include/mftp/mftp_automaton.hpp

@@ -39,6 +39,7 @@ namespace mftp {
     static const ioa::time INIT_ANNOUNCEMENT_INTERVAL;
     static const ioa::time MAX_ANNOUNCEMENT_INTERVAL;
     static const ioa::time MATCHING_INTERVAL;
+    static const ioa::time PROGRESS_INTERVAL;
 
     enum send_state_t {
       SEND_READY,
@@ -64,8 +65,10 @@ namespace mftp {
     timer_state_t m_request_timer_state;
     timer_state_t m_announcement_timer_state;
     timer_state_t m_matching_timer_state;
+    timer_state_t m_progress_timer_state;
     ioa::time m_announcement_interval;
     bool m_reported; // True when we have reported a complete download.
+    bool m_progress;
 
     ioa::handle_manager<mftp_sender_automaton> m_sender;
     ioa::handle_manager<mftp_receiver_automaton> m_receiver;
@@ -76,6 +79,7 @@ namespace mftp {
     const bool m_get_matching_files; // Always get matching files.
 
     bool m_suicide_flag;  //Self-destruct when job is done.
+
 
     std::set<fileid> m_pending_matches;
     std::set<fileid> m_matches;
@@ -134,6 +138,18 @@ namespace mftp {
 				   &m_self,
 				   &mftp_automaton::matching_timer_interrupt);
       }
+
+      ioa::automaton_manager<ioa::alarm_automaton>* progress_clock = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
+      ioa::make_binding_manager (this,
+				 &m_self,
+				 &mftp_automaton::set_progress_timer,
+				 progress_clock,
+				 &ioa::alarm_automaton::set);
+      ioa::make_binding_manager (this,
+				 progress_clock,
+				 &ioa::alarm_automaton::alarm,
+				 &m_self,
+				 &mftp_automaton::progress_timer_interrupt);
       
       schedule ();
     }
@@ -159,6 +175,7 @@ namespace mftp {
       m_matching_timer_state (SET_READY),
       m_announcement_interval (INIT_ANNOUNCEMENT_INTERVAL),
       m_reported (m_file.complete ()),
+      m_progress (false),
       m_sender (sender),
       m_receiver (receiver),
       m_matching (false),
@@ -191,6 +208,7 @@ namespace mftp {
       m_matching_timer_state (SET_READY),
[…]

### 2011-07-19 lib/file.cpp

@@ -239,6 +239,10 @@ namespace mftp {
     return rf;  
   }
 
+  uint32_t file::get_progress () const {
+    return m_have_count;
+  }
+
   uint32_t file::offset_to_hash (const uint32_t offset) const {
     return m_mfileid.get_padded_length () + HASH_SIZE * offset / FRAGMENT_SIZE;
   }

### 2011-07-19 src/get.cpp

@@ -15,9 +15,12 @@ namespace jam {
   private:
     ioa::handle_manager<mftp_client_automaton> m_self;
     std::set<mftp::fileid> meta_files;
+    std::map<mftp::fileid, ioa::automaton_manager<mftp::mftp_automaton>* > data_files;
     ioa::automaton_manager<mftp::mftp_sender_automaton>* sender;
     ioa::automaton_manager<mftp::mftp_receiver_automaton>* receiver;
     std::string m_filename;
+
+    static const ioa::time PROGRESS_INTERVAL;
 
   public:
     mftp_client_automaton (std::string fname) :
@@ -63,10 +66,16 @@ namespace jam {
       fid.convert_to_host();
       
       ioa::automaton_manager<mftp::mftp_automaton>* file_home = new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (mftp::file (fid), sender->get_handle(), receiver->get_handle(), false));
+
+      data_files.insert(std::make_pair(fid, file_home));
       
       ioa::make_binding_manager (this,
 				 file_home, &mftp::mftp_automaton::download_complete,
 				 &m_self, &mftp_client_automaton::file_complete);
+
+      /*ioa::make_binding_manager (this,
+				 file_home, &mftp::mftp_automaton::report_progress,
+				 &m_self, &mftp_client_automaton::update_progress);*/
     }
   
   public:
@@ -85,8 +94,36 @@ namespace jam {
   public:
     V_AP_INPUT (mftp_client_automaton, file_complete, mftp::file);
   
-  };
 
+  private:
+    void progress_update_effect (const uint32_t& have, ioa::aid_t id) {
+      //Move the iterator to the right fileid.
+      std::map<mftp::fileid, ioa::automaton_manager<mftp::mftp_automaton>*>::iterator it;
+      for(it = data_files.begin ();
+	  it->second->get_handle() != id;
+	  ++it) { ;  }
+
+      if (it == data_files.end ()){
+	std::cerr << "OH NOES" << std::endl;
+      }
+
+      mftp::mfileid mid (it->first);
+      std::cout << "Received " << have << " of " << mid.get_fragment_count () << " fragments" << std::endl;
+    }
+    
+  public:
+    V_AP_INPUT (mftp_client_automaton, progress_update, uint32_t);
+    
+  private:
+    bool report_progress_precondition () const {
+      return false; 
+    }
+    
[…]

### 2011-07-19 include/mftp/mftp_automaton.hpp
added progress reporting

@@ -255,6 +255,9 @@ namespace mftp {
       if (suicide_precondition ()) {
 	ioa::schedule (&mftp_automaton::suicide);
       }
+      if (set_progress_timer_precondition ()){
+	ioa::schedule (&mftp_automaton::set_progress_timer);
+      }
       if (report_progress_precondition ()) {
 	ioa::schedule (&mftp_automaton::report_progress);
       }
@@ -284,6 +287,7 @@ namespace mftp {
 
   private:
     void receive_effect (const ioa::const_shared_ptr<message>& m) {
+      std::cout << __func__ << std::endl;
       switch (m->header.message_type) {
       case FRAGMENT:
 	{
@@ -594,6 +598,8 @@ namespace mftp {
 
     bool set_progress_timer_precondition () const {
       return m_progress_timer_state == SET_READY && m_progress == false && ioa::binding_count (&mftp_automaton::set_progress_timer) != 0;
+      //std::cout << __func__ << " : " << b << std::endl;
+      //return b;
     }
 
     ioa::time set_progress_timer_effect () {
@@ -615,6 +621,7 @@ namespace mftp {
 
   private:
     bool report_progress_precondition () const {
+      //std::cout << __func__ << " : " << m_progress << std::endl;
       return m_progress;
     }
 
@@ -711,7 +718,7 @@ namespace mftp {
   const ioa::time mftp_automaton::INIT_ANNOUNCEMENT_INTERVAL (1, 0); //1 second
   const ioa::time mftp_automaton::MAX_ANNOUNCEMENT_INTERVAL (64, 0); //slightly over 1 minute
   const ioa::time mftp_automaton::MATCHING_INTERVAL (4, 0); //4 seconds
-  const ioa::time mftp_automaton::PROGRESS_INTERVAL (5, 0);
+  const ioa::time mftp_automaton::PROGRESS_INTERVAL (2, 0);
 
 }
 

### 2011-07-19 src/get.cpp

@@ -73,9 +73,9 @@ namespace jam {
 				 file_home, &mftp::mftp_automaton::download_complete,
 				 &m_self, &mftp_client_automaton::file_complete);
 
-      /*ioa::make_binding_manager (this,
+      ioa::make_binding_manager (this,
 				 file_home, &mftp::mftp_automaton::report_progress,
-				 &m_self, &mftp_client_automaton::update_progress);*/
+				 &m_self, &mftp_client_automaton::update_progress);
     }
   
   public:
@@ -96,7 +96,7 @@ namespace jam {
   
 
   private:
-    void progress_update_effect (const uint32_t& have, ioa::aid_t id) {
+    void update_progress_effect (const uint32_t& have, ioa::aid_t id) {
       //Move the iterator to the right fileid.
       std::map<mftp::fileid, ioa::automaton_manager<mftp::mftp_automaton>*>::iterator it;
       for(it = data_files.begin ();
@@ -112,7 +112,7 @@ namespace jam {
     }
     
   public:
-    V_AP_INPUT (mftp_client_automaton, progress_update, uint32_t);
+    V_AP_INPUT (mftp_client_automaton, update_progress, uint32_t);
     
   private:
     bool report_progress_precondition () const {

### 2011-07-19 include/mftp/mftp_automaton.hpp
added progress reporting; weird bug

@@ -185,6 +185,7 @@ namespace mftp {
       m_suicide_flag (suicide)
     {
       create_bindings ();
+      std::cout << "Created " << this << std::endl;
     }
 
     // Matching.
@@ -221,6 +222,7 @@ namespace mftp {
       m_suicide_flag (suicide)
     {
       create_bindings ();
+      std::cout << "Created " << this << std::endl;
     }
 
   private:
@@ -294,7 +296,6 @@ namespace mftp {
 
   private:
     void receive_effect (const ioa::const_shared_ptr<message>& m) {
-      std::cout << __func__ << std::endl;
       switch (m->header.message_type) {
       case FRAGMENT:
 	{
@@ -609,7 +610,7 @@ namespace mftp {
     UV_UP_INPUT (mftp_automaton, matching_timer_interrupt);
 
     bool set_progress_timer_precondition () const {
-      return m_progress_timer_state == SET_READY && m_progress == false && ioa::binding_count (&mftp_automaton::set_progress_timer) != 0;
+      return m_progress_timer_state == SET_READY && ioa::binding_count (&mftp_automaton::set_progress_timer) != 0;
       //std::cout << __func__ << " : " << b << std::endl;
       //return b;
     }
@@ -624,6 +625,7 @@ namespace mftp {
 
   private:
     void progress_timer_interrupt_effect () {
+      std::cout << __func__ << std::endl;
       m_progress = true;
       m_progress_timer_state = SET_READY;
     }
@@ -638,6 +640,7 @@ namespace mftp {
     }
 
     uint32_t report_progress_effect () {
+      std::cout << __func__ << std::endl;
       m_progress = false;
       return m_file.get_progress ();
     }
@@ -724,13 +727,18 @@ namespace mftp {
       return new message_buffer (fragment_type (), m_fileid, offset, static_cast<const char*> (m_file.get_data_ptr ()) + offset);
     }
 
+
+    ~mftp_automaton (){
+      std::cout << "destroyed" << this << std::endl;
+    }
+
   };
 
   const ioa::time mftp_automaton::REQUEST_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::INIT_ANNOUNCEMENT_INTERVAL (1, 0); //1 second
   const ioa::time mftp_automaton::MAX_ANNOUNCEMENT_INTERVAL (64, 0); //slightly over 1 minute
[…]

### 2011-07-19 lib/file.cpp

@@ -95,7 +95,12 @@ namespace mftp {
     m_valid (m_mfileid.get_fragment_count ()),
     m_valid_count (0),
     m_start_idx (0)
-  { }
+  {
+    for (uint32_t idx = 0; idx < m_mfileid.get_fragment_count (); ++idx) {
+      assert (m_have[idx] == false);
+      assert (m_valid[idx] == false);
+    }
+  }
 
   file::file (const file& other) :
     m_mfileid (other.m_mfileid),
@@ -214,11 +219,15 @@ namespace mftp {
     //bring m_start_indx into range
     m_start_idx = m_start_idx % m_mfileid.get_fragment_count();
  
+    size_t loop_counter = 0;
+
     // Move start until we don't have the fragment.
     for (;
-	 m_have[m_start_idx] == true;
-	 m_start_idx = (m_start_idx + 1) % m_mfileid.get_fragment_count ())
+	 m_have[m_start_idx] == true && loop_counter < m_mfileid.get_fragment_count ();
+	 m_start_idx = (m_start_idx + 1) % m_mfileid.get_fragment_count (), loop_counter++)
       ;;
+
+    assert (loop_counter < m_mfileid.get_fragment_count ());
       
     // Move end until we do have a fragment or reach the end.
     uint32_t end_idx;

### 2011-07-19 src/get.cpp

@@ -97,10 +97,11 @@ namespace jam {
 
   private:
     void update_progress_effect (const uint32_t& have, ioa::aid_t id) {
+      std::cout << __func__ << std::endl;
       //Move the iterator to the right fileid.
       std::map<mftp::fileid, ioa::automaton_manager<mftp::mftp_automaton>*>::iterator it;
       for(it = data_files.begin ();
-	  it->second->get_handle() != id;
+	  it->second->get_handle() != id && it != data_files.end ();
 	  ++it) { ;  }
 
       if (it == data_files.end ()){

### 2011-07-19 include/mftp/mftp_automaton.hpp
Combinding sender and receiver.

@@ -3,8 +3,7 @@
 
 #include <mftp/file.hpp>
 
-#include <mftp/mftp_sender_automaton.hpp>
-#include <mftp/mftp_receiver_automaton.hpp>
+#include <mftp/mftp_channel_automaton.hpp>
 #include <ioa/alarm_automaton.hpp>
 
 #include <config.hpp>
@@ -68,8 +67,7 @@ namespace mftp {
     ioa::time m_announcement_interval;
     bool m_reported; // True when we have reported a complete download.
 
-    ioa::handle_manager<mftp_sender_automaton> m_sender;
-    ioa::handle_manager<mftp_receiver_automaton> m_receiver;
+    ioa::handle_manager<mftp_channel_automaton> m_channel;
 
     const bool m_matching; // Try to find matches for this file.
     std::auto_ptr<match_candidate_predicate> m_match_candidate_predicate;
@@ -88,14 +86,14 @@ namespace mftp {
     void create_bindings () {
       ioa::make_binding_manager (this,
 				 &m_self, &mftp_automaton::send,
-				 &m_sender, &mftp_sender_automaton::send);
+				 &m_channel, &mftp_channel_automaton::send);
       
       ioa::make_binding_manager (this,
-				 &m_sender, &mftp_sender_automaton::send_complete,
+				 &m_channel, &mftp_channel_automaton::send_complete,
 				 &m_self, &mftp_automaton::send_complete);
       
       ioa::make_binding_manager (this,
-				 &m_receiver, &mftp_receiver_automaton::receive,
+				 &m_channel, &mftp_channel_automaton::receive,
 				 &m_self, &mftp_automaton::receive);
 
       ioa::automaton_manager<ioa::alarm_automaton>* request_clock = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
@@ -142,8 +140,7 @@ namespace mftp {
   public:
     // Not matching.
     mftp_automaton (const file& file,
-		    const ioa::automaton_handle<mftp_sender_automaton>& sender,
-		    const ioa::automaton_handle<mftp_receiver_automaton>& receiver,
+		    const ioa::automaton_handle<mftp_channel_automaton>& channel,
 		    const bool suicide) :
       m_self (ioa::get_aid ()),
       m_file (file),
@@ -161,8 +158,7 @@ namespace mftp {
       m_matching_timer_state (SET_READY),
[…]

### 2011-07-19 include/mftp/mftp_channel_automaton.hpp

@@ -1,7 +1,8 @@
-#ifndef __mftp_sender_automaton_hpp__
-#define __mftp_sender_automaton_hpp__
+#ifndef __mftp_channel_automaton_hpp__
+#define __mftp_channel_automaton_hpp__
 
 #include <ioa/udp_sender_automaton.hpp>
+#include <ioa/udp_receiver_automaton.hpp>
 #include <mftp/message.hpp>
 
 #include <queue>
@@ -9,18 +10,19 @@
 
 namespace mftp {
 
-  class mftp_sender_automaton :
+  class mftp_channel_automaton :
     public ioa::automaton,
     private ioa::observer
   {
   private:
-    ioa::handle_manager<mftp_sender_automaton> m_self;
+    ioa::handle_manager<mftp_channel_automaton> m_self;
     typedef std::pair<ioa::const_shared_ptr<message_buffer>, ioa::aid_t> message_aid;
     std::list<message_aid > m_outgoing_messages;
     std::set<ioa::aid_t> m_outgoing_set;
     ioa::aid_t m_pending_aid;
     std::set<ioa::aid_t> m_outgoing_completes;
     const ioa::inet_address m_send;
+    std::queue<ioa::const_shared_ptr<mftp::message> > m_incoming_messages;
 
     struct message_aid_equal {
       const ioa::aid_t m_aid;
@@ -35,7 +37,9 @@ namespace mftp {
     };
 
   public:
-    mftp_sender_automaton (ioa::inet_address send_address) :
+    mftp_channel_automaton (const ioa::inet_address& send_address,
+			    const ioa::inet_address& local_address,
+			    const bool multicast) :
       m_self (ioa::get_aid ()),
       m_pending_aid (-1),
       m_send (send_address)
@@ -46,25 +50,42 @@ namespace mftp {
       ioa::automaton_manager<ioa::udp_sender_automaton>* sender = new ioa::automaton_manager<ioa::udp_sender_automaton> (this, ioa::make_generator<ioa::udp_sender_automaton> (sizeof (message)));
 
       ioa::make_binding_manager (this,
-				 &m_self, &mftp_sender_automaton::send_out,
+				 &m_self, &mftp_channel_automaton::send_out,
 				 sender, &ioa::udp_sender_automaton::send);
 
       ioa::make_binding_manager (this,
 				 sender, &ioa::udp_sender_automaton::send_complete,
-				 &m_self, &mftp_sender_automaton::send_in_complete);
[…]

### 2011-07-19 src/get.cpp

@@ -15,8 +15,7 @@ namespace jam {
   private:
     ioa::handle_manager<mftp_client_automaton> m_self;
     std::set<mftp::fileid> meta_files;
-    ioa::automaton_manager<mftp::mftp_sender_automaton>* sender;
-    ioa::automaton_manager<mftp::mftp_receiver_automaton>* receiver;
+    ioa::automaton_manager<mftp::mftp_channel_automaton>* channel;
     std::string m_filename;
 
   public:
@@ -24,11 +23,9 @@ namespace jam {
       m_self (ioa::get_aid ()),
       m_filename (fname)
     {
-      sender = new ioa::automaton_manager<mftp::mftp_sender_automaton> (this, ioa::make_generator<mftp::mftp_sender_automaton> (jam::SEND_ADDR));
-      receiver = new ioa::automaton_manager<mftp::mftp_receiver_automaton> (this, ioa::make_generator<mftp::mftp_receiver_automaton> (jam::LOCAL_ADDR, jam::MULTICAST_ADDR));
+      channel = new ioa::automaton_manager<mftp::mftp_channel_automaton> (this, ioa::make_generator<mftp::mftp_channel_automaton> (jam::SEND_ADDR, jam::LOCAL_ADDR, true));
 
-      add_observable (sender);
-      add_observable (receiver);
+      add_observable (channel);
     }
   private:
     void schedule () const { }
@@ -36,19 +33,16 @@ namespace jam {
     void observe (ioa::observable* o) {
       //need to do -1 == sender->get_handle () instead of the normal syntax
       //the other way it gets confused whether to convert -1 to a handle or the get_handle () handle to an int
-      if (o == sender && -1 == sender->get_handle ()) {
-	sender = 0;
-      }
-      else if (o == receiver && -1 == receiver->get_handle()) {
-	receiver = 0;
+      if (o == channel && -1 == channel->get_handle ()) {
+	channel = 0;
       }
 
-      if (sender != 0 && receiver != 0) {
-	if (sender->get_handle () != -1 && receiver->get_handle () != -1) {
+      if (channel != 0) {
+	if (channel->get_handle () != -1) {
 	  mftp::file f (m_filename.c_str (), m_filename.size (), QUERY_TYPE);
 
 	  // Create the query server.
[…]

### 2011-07-20 include/mftp/file.hpp
Fixed progress bug where state was uninitialized. Fixed bugs in fileid where offsets were exaggerated. Validation works correctly but I plan to take this out. Answering requests with random fragments makes requesting hard because the span become smaller and smaller. The next step is to redo the request mechanism.

@@ -17,8 +17,8 @@ namespace mftp {
     uint32_t m_valid_count;
     uint32_t m_start_idx;
 
-    uint32_t offset_to_hash (const uint32_t) const;
-    uint32_t hash_to_offset (const uint32_t) const;
+    uint32_t idx_to_hash (const uint32_t) const;
+    uint32_t hash_to_idx (const uint32_t) const;
     bool get_previous_hash (const uint32_t,
 			    unsigned char*) const;
     bool get_hash (const uint32_t,

### 2011-07-20 include/mftp/mftp_automaton.hpp

@@ -148,7 +148,6 @@ namespace mftp {
 				 &ioa::alarm_automaton::alarm,
 				 &m_self,
 				 &mftp_automaton::progress_timer_interrupt);
-      
       schedule ();
     }
 
@@ -171,6 +170,7 @@ namespace mftp {
       m_request_timer_state (SET_READY),
       m_announcement_timer_state (SET_READY),
       m_matching_timer_state (SET_READY),
+      m_progress_timer_state (SET_READY),
       m_announcement_interval (INIT_ANNOUNCEMENT_INTERVAL),
       m_reported (m_file.complete ()),
       m_progress (false),
@@ -180,7 +180,6 @@ namespace mftp {
       m_suicide_flag (suicide)
     {
       create_bindings ();
-      std::cout << "Created " << this << std::endl;
     }
 
     // Matching.
@@ -204,6 +203,7 @@ namespace mftp {
       m_request_timer_state (SET_READY),
       m_announcement_timer_state (SET_READY),
       m_matching_timer_state (SET_READY),
+      m_progress_timer_state (SET_READY),
       m_announcement_interval (INIT_ANNOUNCEMENT_INTERVAL),
       m_reported (m_file.complete ()),
       m_progress (false),
@@ -215,7 +215,6 @@ namespace mftp {
       m_suicide_flag (suicide)
     {
       create_bindings ();
-      std::cout << "Created " << this << std::endl;
     }
 
   private:
@@ -604,8 +603,6 @@ namespace mftp {
 
     bool set_progress_timer_precondition () const {
       return m_progress_timer_state == SET_READY && ioa::binding_count (&mftp_automaton::set_progress_timer) != 0;
-      //std::cout << __func__ << " : " << b << std::endl;
-      //return b;
     }
 
     ioa::time set_progress_timer_effect () {
@@ -618,7 +615,7 @@ namespace mftp {
 
   private:
     void progress_timer_interrupt_effect () {
-      std::cout << __func__ << std::endl;
+      assert (m_progress_timer_state == INTERRUPT_WAIT);
       m_progress = true;
       m_progress_timer_state = SET_READY;
     }
@@ -628,12 +625,10 @@ namespace mftp {
 
   private:
     bool report_progress_precondition () const {
-      //std::cout << __func__ << " : " << m_progress << std::endl;
       return m_progress;
     }
 
[…]

### 2011-07-20 lib/file.cpp

@@ -9,6 +9,8 @@
 #include <unistd.h>
 #include <stdio.h>
 
+#include <iostream>
+
 namespace mftp {
 
   file::file (const std::string& name,
@@ -95,12 +97,7 @@ namespace mftp {
     m_valid (m_mfileid.get_fragment_count ()),
     m_valid_count (0),
     m_start_idx (0)
-  {
-    for (uint32_t idx = 0; idx < m_mfileid.get_fragment_count (); ++idx) {
-      assert (m_have[idx] == false);
-      assert (m_valid[idx] == false);
-    }
-  }
+  { }
 
   file::file (const file& other) :
     m_mfileid (other.m_mfileid),
@@ -160,7 +157,7 @@ namespace mftp {
 	digester.get (samp);
 	m_mfileid.set_hash (samp);
       }
-    }    
+    }
   }
 
   file::~file () {
@@ -209,7 +206,7 @@ namespace mftp {
       // Now we have it.
       m_have[idx] = true;
       ++m_have_count;
-      validate (offset);
+      validate (idx);
     }
   }
 
@@ -219,15 +216,11 @@ namespace mftp {
     //bring m_start_indx into range
     m_start_idx = m_start_idx % m_mfileid.get_fragment_count();
  
-    size_t loop_counter = 0;
-
     // Move start until we don't have the fragment.
     for (;
-	 m_have[m_start_idx] == true && loop_counter < m_mfileid.get_fragment_count ();
-	 m_start_idx = (m_start_idx + 1) % m_mfileid.get_fragment_count (), loop_counter++)
+	 m_have[m_start_idx] == true;
+	 m_start_idx = (m_start_idx + 1) % m_mfileid.get_fragment_count ())
       ;;
-
-    assert (loop_counter < m_mfileid.get_fragment_count ());
       
     // Move end until we do have a fragment or reach the end.
     uint32_t end_idx;
@@ -256,81 +249,93 @@ namespace mftp {
     return m_have_count;
   }
 
-  uint32_t file::offset_to_hash (const uint32_t offset) const {
-    return m_mfileid.get_padded_length () + HASH_SIZE * offset / FRAGMENT_SIZE;
+  uint32_t file::idx_to_hash (const uint32_t idx) const {
+    return m_mfileid.get_padded_length () + HASH_SIZE * idx;
   }
 
-  uint32_t file::hash_to_offset (const uint32_t hash) const {
-    return (hash - m_mfileid.get_padded_length ()) * FRAGMENT_SIZE / HASH_SIZE;
[…]

### 2011-07-20 src/get.cpp

@@ -91,7 +91,6 @@ namespace jam {
 
   private:
     void update_progress_effect (const uint32_t& have, ioa::aid_t id) {
-      std::cout << __func__ << std::endl;
       //Move the iterator to the right fileid.
       std::map<mftp::fileid, ioa::automaton_manager<mftp::mftp_automaton>*>::iterator it;
       for(it = data_files.begin ();
@@ -103,7 +102,9 @@ namespace jam {
       }
 
       mftp::mfileid mid (it->first);
-      std::cout << "Received " << have << " of " << mid.get_fragment_count () << " fragments" << std::endl;
+      if (have < mid.get_fragment_count ()) {
+	std::cout << "Received " << have << " of " << mid.get_fragment_count () << " fragments" << std::endl;
+      }
     }
     
   public:

### 2011-07-20 lib/file.cpp
Verified validation.

@@ -318,9 +318,9 @@ namespace mftp {
 
     	// Try to validate fragments that depend on this fragment.
     	// We move from the back to the front since the hash moved front to back.
-    	// Also, we go one past the end (FRAGMENT_SIZE instead of FRAGMENT_SIZE - 1)
+    	// Also, we go one past the end (FRAGMENT_SIZE instead of FRAGMENT_SIZE - HASH_SIZE)
     	// because that fragment might depend on this one.
-    	for (int32_t off = FRAGMENT_SIZE; off >= static_cast<int32_t>(- HASH_SIZE); off -= HASH_SIZE) {
+    	for (int32_t off = FRAGMENT_SIZE; off >= 0; off -= HASH_SIZE) {
 	  const uint32_t hash_location = offset + off;
 	  if (hash_location >= m_mfileid.get_padded_length () && hash_location < m_mfileid.get_final_length ()) {
 	    const uint32_t dep_idx = hash_to_idx (hash_location);

### 2011-07-20 papers/paper1/paper.tex
A little progress on implentation, changed diagram one

@@
 
+%%TODO: Talk about validation - possible coinage 'incremental digest'
+
+\paragraph{Matching.}
+Each MFTP automaton contains a flag that determines whether it will perform matching. 
+It has another flag that determines whether it will self-destruct when it has finished downloading its file.
+If the matching flag is set, the automaton applies the matching function $\iota$ to the file IDs of incoming fragments.
+If $\iota$ is true, the automaton recursively constructs another MFTP automaton to download the potential match.
+This child automaton does not perform matching.  
+Its self-destruct flag is set so that it will self-destruct once the potential match has been downloaded.
+When the download is complete, the parent automaton applies the matching function $\mu$ to the file to determine whether it is a match.
+
+\section{Design}
+%% General stuff -- The short argument is we're developing a new protocol
+%% the I/O Automata formal model is good for modeling protocols, therefore we modeled it as an mftp automaton
+%% Diagram with the 3 pieces
+%% Describe the diagram, this is how we're going to do it, this is how we're interfaced and kinda put things together
+%% then go into MFTP Automaton, state, etc
+%% then go on and talk about the actions (send receive etc)
+
+The I/O Automata formal model is a well-established technique for building protocols.
+%% Talk about why I/O Automata is a good method (even though we didn't have a choice)
+I/O Automata model stateful distributed system components that can interact and be composed with each other.
+They change state through transitions classified as input, output or internal actions.
+Input and output actions in different automata are bound; output actions in the one trigger input actions in the other.
+As a result, we present our design of the MFTP protocol as automata.
+
+\begin{figure}
+  \center
+  \includegraphics[scale=0.65]{diagramOne}
+  \caption{Basic implementational view of the protocol \label{DesignOne}}
+\end{figure}
+
+\ref{DesignOne} outlines the necessary pieces which constitute our protocol.
+We define a basic MFTP automaton and a broadcast/multicast channel automaton.
+We define a basic MFTP automaton, a sender, and a receiver.
+The responsibilities of the MFTP automaton are to acquire data and to help others acquire it.
+
+Each MFTP automaton begins with a file, encapsulated in a File object, which is either incomplete or complete.
+If the file is incomplete, the automaton generates requests for the missing fragments.
+If the file is complete, the automaton may serve the file, or it may serve the file while performing matching on it.
+
+MFTP automata receive messages via the receive input action.
+If the input message is a valid fragment of its own file, the automaton saves the fragment.
+If the input message is a request for fragments of its file, it flags those fragments as having been requested.
+If the input message is a match containing its file ID, it records the matching IDs for download.
+
+The automaton holds a queue of one message to send. 
+The internal actions which fill this queue are triggered periodically.
+The automaton generates messages containing one requested fragment if any fragments in the file have been requested.
+The fragment sent is selected at random to help avoid collisions with other MFTP automata serving the same file.
+The automaton generates requests if some fragments in the automaton's own file are missing.
+The automaton generates match messages if it has recorded any matches to its file.
+
+The broadcast/multicast channel automaton uses the UDP sender and receiver built into ioa.
+The channel may be connected to more than one MFTP automaton.
+
+The automaton reports that its file is complete using the download complete output action.
+%%  to multicast messages one at a time.
+%%The receiving automaton uses the ioa UDP receiver.
+
+\section{Implementation}
+%% We implimented our IO Automata design using ioa framework
+%% Interesting techniques - suicide, recursive automata
+
+\section{Evaluation and Discussion\label{evaluation}}
+\paragraph{Jam.} %%needs different header
+Our implementation defines three file types, \emph{data}, \emph{meta} and \emph{query}.  
+A data file contains data from any file in the network.
+A meta file contains as data the file ID of a file in the network and its name. 
+A query file contains only the name of the file which is being queried.
+
+We wrote two programs, \emph{Share} and \emph{Get}, which use MFTP, sender and receiver automata to disseminate a file or files. %%Can one 'disseminate' a file?
+Share creates two MFTP automata, \emph{file server} and \emph{meta server}.
+File server is constructed with a complete file \emph{f}, and meta server is constructed with a meta file.
+The meta file is generated from the fileID and name of f.
+At the start, Get creates one MFTP automaton, \emph{query server}, constructed with a query file.
+The query file is generated from the name of the file to download, supplied by a command line argument.
+
+We defined a functor interface so that Share and Get could use different $\iota$ and $\mu$ functions.
+In Share, meta server performs matching on the meta file and broadcasts matches to f.
+Its $\iota$ returns true if the incoming file is a query file.
+Its $\mu$ returns true if the contents of the downloaded query file match the name of f.
+
+\begin{figure}
+  \center
+  \includegraphics[scale=0.65]{share_diagram}
+  \caption{The share program \label{EvaluationOne}}
+\end{figure}
+
+In Get, query server performs matching on the query file.
+Its $\iota$ returns true if the incoming file is a meta file.
+Its $\mu$ returns true if the contents of the downloaded meta file match the name of the file being queried.
+Query server passes the meta files downloaded by its child automata to Get.
+Using the file IDs contained in the meta files, Get constructs further MFTP automata which download the files pertaining to those file IDs.
+
+\begin{figure}
+  \center
+  \includegraphics[scale=0.65]{get_diagram}
+  \caption{The get program \label{EvalutationTwo}}
+\end{figure}
+
+\section{Related Work}
+
+\section{Conclusion}
+
+\end{document}
+
 %% \paragraph {Files.}
@@
 %% each node has its own set of records which it contributes
-
-\section{Implementation}
-
-%%TODO: Talk about validation
-
-%%TODO: Talk about validation - possible coinage 'incremental digest'
-
-\paragraph{MFTP Automaton.}
-Our implementation of MFTP uses the ioa framework.
-I/O Automata model stateful distributed system components that can interact and be composed with each other.
-They change state through transitions classified as input, output or internal actions.
-Input and output actions in different automata are bound; output actions in the one trigger input actions in the other.
-
-%%Describe our implementation at the 50k ft view and work our way down until things are no longer interesting.
-We define a basic MFTP automaton, a sender, and a receiver.
-The responsibilities of the MFTP automaton are to acquire data and to help others acquire it.
-Each MFTP automaton begins with a file, encapsulated in a File object, which is either incomplete or complete.
-If the file is incomplete, the automaton generates requests for the missing fragments.
-If the file is complete, the automaton may serve the file, or it may serve the file while performing matching on it.
-
-%%Picture of mFTP automaton?
-
-MFTP automata receive messages via the receive input action.
-If the input message is a valid fragment of its own file, the automaton saves the fragment.
-If the input message is a request for fragments of its file, it flags those fragments as having been requested.
-If the input message is a match containing its file ID, it records the matching IDs for download.
-
-The automaton holds a queue of one message to send. 
-The internal actions which fill this queue are triggered periodically.
-The automaton generates messages containing one requested fragment if any fragments in the file have been requested.
-The fragment sent is selected at random to help avoid collisions with other MFTP automata serving the same file.
-The automaton generates requests if some fragments in the automaton's own file are missing.
-The automaton generates match messages if it has recorded any matches to its file.
-
-The automaton reports that its file is complete using the download complete output action.
-
-\paragraph{Matching.}
-Each MFTP automaton contains a flag that determines whether it will perform matching. 
-It has another flag that determines whether it will self-destruct when it has finished downloading its file.
-If the matching flag is set, the automaton applies the matching function $\iota$ to the file IDs of incoming fragments.
-If $\iota$ is true, the automaton recursively constructs another MFTP automaton to download the potential match.
-This child automaton does not perform matching.  
-Its self-destruct flag is set so that it will self-destruct once the potential match has been downloaded.
-When the download is complete, the parent automaton applies the matching function $\mu$ to the file to determine whether it is a match.
-
-\paragraph{Some useful stuff.} %% We know it's rough, but what else can we do?
-The sending automaton uses the UDP sender built into ioa to multicast messages one at a time.
-The receiving automaton uses the ioa UDP receiver.
-The sender and receiver automata may be connected to more than one MFTP automaton.
-
-\vspace{5mm}
-\includegraphics[scale=0.65]{diagramOne}
-
-\paragraph{Jam.} %%needs different header
-Our implementation defines three file types, \emph{data}, \emph{meta} and \emph{query}.  
-A data file contains data from any file in the network.
-A meta file contains as data the file ID of a file in the network and its name. 
-A query file contains only the name of the file which is being queried.
-
-We wrote two programs, \emph{Share} and \emph{Get}, which use MFTP, sender and receiver automata to disseminate a file or files. %%Can one 'disseminate' a file?
-Share creates two MFTP automata, \emph{file server} and \emph{meta server}.
-File server is constructed with a complete file \emph{f}, and meta server is constructed with a meta file.
-The meta file is generated from the fileID and name of f.
-At the start, Get creates one MFTP automaton, \emph{query server}, constructed with a query file.
-The query file is generated from the name of the file to download, supplied by a command line argument.
-
-We defined a functor interface so that Share and Get could use different $\iota$ and $\mu$ functions.
-In Share, meta server performs matching on the meta file and broadcasts matches to f.
-Its $\iota$ returns true if the incoming file is a query file.
-Its $\mu$ returns true if the contents of the downloaded query file match the name of f.
-
-\vspace{2 mm}
-
-\includegraphics[scale=0.55]{share_diagram}
-
-\vspace{2 mm}
-
-In Get, query server performs matching on the query file.
-Its $\iota$ returns true if the incoming file is a meta file.
-Its $\mu$ returns true if the contents of the downloaded meta file match the name of the file being queried.
-Query server passes the meta files downloaded by its child automata to Get.
-Using the file IDs contained in the meta files, Get constructs further MFTP automata which download the files pertaining to those file IDs.
-
-\vspace{2.5 mm}
-
-\includegraphics[scale=0.65]{get_diagram}
-
-\vspace{2 mm}
-
-%%\begin{itemize}
-%%  \item Problem
-%%  \item Design forces
-%%  \item Solution
-%%  \item Consequences
-%%\end{itemize}
-
-\section{Evaluation and Discussion\label{evaluation}}
-
-\begin{itemize}
-\item Translation of automaton in \emph{Distributed Algorithms} to C++.
-\item Simulate a protocol then replace with network components
-\item Compare a protocol using: single-threaded event-based (select loop), multi-threaded, I/O automata
-\item Show a buggy program and then apply an invariant to find the bug.
-\end{itemize}
-
-Bring up style mentioned in section \ref{representation}.
-
-\section{Related Work\label{related_work}}
-
-pthreads (?)
-Edward Lee - The Problem with Threads
-Herb Sutter - The Free Lunch is Over
-Early work on concurrency
-
-\section{Conclusion and Future Work\label{conclusion}}
-
-\begin{itemize}
-  \item We are going to use it to build the substrate.
-  \item Speculate on moving down into operating system (device drivers would be easy, IPC including filesystem replaced by automata)
-  \item Speculate on moving down into the hardware level (Local talent, Ivan Sutherland)
-  \item We can take advantage of multi-core in a very straight-forward way
-  \item New problems in scheduling (Pinning automata to processors to minimize actions that span two processors.  Maximum independent set.)
-  \item We are non-blocking all the way.  Combine this with a deterministic implementation of the scheduler and model and there are serious opportunities for real-time.
-\end{itemize}
-
-\end{document}

### 2011-07-20 include/mftp/mftp_automaton.hpp
Fixed a little strerror pound define misunderstanding

@@ -6,7 +6,6 @@
 #include <mftp/mftp_channel_automaton.hpp>
 #include <ioa/alarm_automaton.hpp>
 
-#include <config.hpp>
 #include <cstdlib>
 #include <queue>
 #include <set>

### 2011-07-20 include/mftp/mftp_automaton.hpp
solved Dr. Jaydub's riddle

@@ -56,6 +56,7 @@ namespace mftp {
     const fileid& m_fileid;
     std::vector<bool> m_requests; // Bit vector indicating fragments that have been requested.
     uint32_t m_requests_count; // Number of true elements in m_requests.
+    uint32_t m_last_sent_idx; //Index of last fragment sent
     std::queue<ioa::const_shared_ptr<message_buffer> > m_sendq;
     send_state_t m_send_state;
     size_t m_fragment_count;
@@ -162,6 +163,7 @@ namespace mftp {
       m_fileid (m_mfileid.get_fileid ()),
       m_requests (m_mfileid.get_fragment_count ()),
       m_requests_count (0),
+      m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
       m_send_state (SEND_READY),
       m_fragment_count (0),
       m_send_request (!m_file.complete ()),
@@ -195,6 +197,7 @@ namespace mftp {
       m_fileid (m_mfileid.get_fileid ()),
       m_requests (m_mfileid.get_fragment_count ()),
       m_requests_count (0),
+      m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
       m_send_state (SEND_READY),
       m_fragment_count (0),
       m_send_request (!m_file.complete ()),
@@ -445,13 +448,13 @@ namespace mftp {
 
     void send_fragment_effect () {
       // Purpose is to produce a randomly selected requested fragment.
-      // Get a random index.
-      uint32_t randy = get_random_request_index ();
-      m_requests[randy] = false;
+      advance_to_next_request_index ();
+      m_requests[m_last_sent_idx] = false;
       --m_requests_count;
       
+      
       // Get the fragment for that index.
-      message_buffer* m = get_fragment (randy);
+      message_buffer* m = get_fragment (m_last_sent_idx);
       m->convert_to_network ();
       m_sendq.push (ioa::const_shared_ptr<message_buffer> (m));
       ++m_fragment_count;
@@ -703,11 +706,9 @@ namespace mftp {
     V_UP_OUTPUT (mftp_automaton, match_complete, ioa::const_shared_ptr<file>);
 
   private:
-    uint32_t get_random_request_index () {
+    void advance_to_next_request_index () {
       assert (m_requests_count != 0);
-      uint32_t rf = rand () % m_requests.size();
-      for (; !m_requests[rf]; rf = (rf + 1) % m_requests.size ()) { }
[…]

### 2011-07-20 include/mftp/mftp_automaton.hpp
Added error checking to file_complete_effect.

@@ -2,7 +2,6 @@
 #define	__mftp_automaton_hpp__
 
 #include <mftp/file.hpp>
-
 #include <mftp/mftp_channel_automaton.hpp>
 #include <ioa/alarm_automaton.hpp>
 
@@ -11,8 +10,6 @@
 #include <set>
 #include <vector>
 #include <cstring>
-
-// TODO:  We accumulate matches without ever forgetting them.  We should purge matches that we haven't seen in a while.
 
 namespace mftp {
   
@@ -667,7 +664,6 @@ namespace mftp {
 					 ioa::aid_t aid) {
       //For use when the child has reported a download_complete.
       process_match_candidate (f);
-      // TODO:  Destroy the mftp_automaton that provided us with the file.
     }
 
     V_AP_INPUT (mftp_automaton, match_download_complete, file);

### 2011-07-20 src/get.cpp

@@ -80,14 +80,28 @@ namespace jam {
       std::string path (m_filename + "-" + f.get_mfileid ().get_fileid ().to_string ());
       // TODO:  Add error checking.
       FILE* fp = fopen (path.c_str (), "w");
-      fwrite (f.get_data_ptr (), 1, f.get_mfileid ().get_original_length (), fp);
-      fclose (fp);
+      if (fp == NULL) {
+	perror ("fopen");
+	exit (EXIT_FAILURE);
+      }
+
+      size_t er = fwrite (f.get_data_ptr (), 1, f.get_mfileid ().get_original_length (), fp);
+      if (er < f.get_mfileid ().get_original_length ()) {
+	std::cerr << "fwrite: couldn't write to " << m_filename << std::endl;
+	exit (EXIT_FAILURE);
+      }
+
+      int err = fclose (fp);
+      if (err != 0) {
+	perror ("fclose");
+	exit (EXIT_FAILURE);
+      }
+      
       std::cout << "Created " << path << std::endl;
     }
   
   public:
     V_AP_INPUT (mftp_client_automaton, file_complete, mftp::file);
-  
 
   private:
     void update_progress_effect (const uint32_t& have, ioa::aid_t id) {
@@ -109,17 +123,8 @@ namespace jam {
     
   public:
     V_AP_INPUT (mftp_client_automaton, update_progress, uint32_t);
-    
-  private:
-    bool report_progress_precondition () const {
-      return false; 
-    }
-    
-    void report_progress_effect () {
-      //print progress
-    }
-  };
-  
+
+  };  
 }
 
 int main (int argc, char* argv[]) {

### 2011-07-20 include/mftp/file.hpp
We can request more frequently as we receive fragments.

@@ -42,7 +42,7 @@ namespace mftp {
     bool complete () const;
     bool empty () const;
     bool have (const uint32_t offset) const;
-    void write_chunk (const uint32_t offset,
+    bool write_chunk (const uint32_t offset,
 		      const void* data);
     span_t get_next_range ();
     uint32_t get_random_index () const;

### 2011-07-20 include/mftp/mftp_automaton.hpp

@@ -46,6 +46,9 @@ namespace mftp {
       INTERRUPT_WAIT,
     };
 
+    #define REQUEST_NUMERATOR 4
+    #define REQUEST_DENOMINATOR 5
+    
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;
     const mfileid& m_mfileid;
@@ -66,6 +69,8 @@ namespace mftp {
     ioa::time m_announcement_interval;
     bool m_reported; // True when we have reported a complete download.
     bool m_progress;
+    uint32_t m_pending_requests;
+    uint32_t m_since_received;
 
     ioa::handle_manager<mftp_channel_automaton> m_channel;
 
@@ -172,6 +177,8 @@ namespace mftp {
       m_announcement_interval (INIT_ANNOUNCEMENT_INTERVAL),
       m_reported (m_file.complete ()),
       m_progress (false),
+      m_pending_requests (0),
+      m_since_received (0),
       m_channel (channel),
       m_matching (false),
       m_get_matching_files (false),
@@ -206,6 +213,8 @@ namespace mftp {
       m_announcement_interval (INIT_ANNOUNCEMENT_INTERVAL),
       m_reported (m_file.complete ()),
       m_progress (false),
+      m_pending_requests (0),
+      m_since_received (0),
       m_channel (channel),
       m_matching (true),
       m_match_candidate_predicate (match_candidate_pred.clone ()),
@@ -297,7 +306,13 @@ namespace mftp {
 	    
 	    // Save the fragment.
 	    if (!m_file.complete ()){
-	      m_file.write_chunk (m->frag.offset, m->frag.data);
+	      if (m_file.write_chunk (m->frag.offset, m->frag.data)) {
+		++m_since_received;
+		//If we have received more than 80% of the fragments we requested...
+		if (REQUEST_DENOMINATOR * m_since_received > REQUEST_NUMERATOR * m_pending_requests){
+		  m_send_request = true;
+		}
+	      }
 	    }
 	    uint32_t idx = (m->frag.offset / FRAGMENT_SIZE);
 	    if (m_requests[idx]) {
@@ -319,6 +334,7 @@ namespace mftp {
 
 	    file f (m->frag.fid);
 	    f.write_chunk (m->frag.offset, m->frag.data);
+
 	    if (f.complete()) {
 	      // We received the whole file.
 	      process_match_candidate (f);
[…]

### 2011-07-20 lib/file.cpp

@@ -194,7 +194,7 @@ namespace mftp {
     return m_have[offset / FRAGMENT_SIZE];
   }
 
-  void file::write_chunk (const uint32_t offset,
+  bool file::write_chunk (const uint32_t offset,
 			  const void* data) {
     assert (offset % FRAGMENT_SIZE == 0);
     assert (offset < m_mfileid.get_final_length ());
@@ -207,7 +207,9 @@ namespace mftp {
       m_have[idx] = true;
       ++m_have_count;
       validate (idx);
+      return m_have[idx];
     }
+    return false;
   }
 
   span_t file::get_next_range () {

### 2011-08-08 lib/mftp_automaton.cpp
Corrected send_match reset.

@@ -192,14 +192,14 @@ namespace mftp {
   void mftp_automaton::send_match (bool reset) {
     if (!m_matches.empty ()) {
 
-      ioa::time now = ioa::time::now ();
       // Reset if required.
       if (reset) {
-	m_match_time = now;
+	m_match_time = ioa::time ();
 	m_match_interval = INIT_INTERVAL;
       }
 
       // Send matches if ready.
+      ioa::time now = ioa::time::now ();
       if (m_match_time + m_match_interval <= now) {
 	// TODO:  We assume that all matches can be send in m_match_interval.
 	std::set<fileid>::const_iterator pos = m_matches.begin ();

### 2011-08-08 lib/mftp_automaton.cpp
Still working on send_request.

@@ -146,22 +146,18 @@ namespace mftp {
     if (!m_file.complete ()) {
 
       ioa::time now = ioa::time::now ();
-      // Reset if necessary.
-      if (reset) {
-	m_request_time = now;
-	m_request_interval = INIT_INTERVAL;
+      bool time_flag = (m_request_time + m_request_interval <= now);
+      bool frag_flag = (REQUEST_DENOMINATOR * m_fragments_since_request > REQUEST_NUMERATOR * m_last_request_size);
+
+      if (time_flag && !frag_flag) {
+	// Increase the interval.
+	m_request_interval += m_request_interval;
+	m_request_interval = std::min (m_request_interval, MAX_INTERVAL);
       }
 
       // If we have received more than 80% of the fragments we requested
       // Or enough time has elapsed
-      if ((REQUEST_DENOMINATOR * m_fragments_since_request > REQUEST_NUMERATOR * m_last_request_size) ||
-	  (m_request_time + m_request_interval <= now)) {
-	// Increase the interval.
-	if (m_request_time + m_request_interval <= now) {
-	  m_request_interval += m_request_interval;
-	  m_request_interval = std::min (m_request_interval, MAX_INTERVAL);
-	}
-
+      if (time_flag || frag_flag) {
 	span_t spans[SPANS_SIZE];
 	spans[0] = m_file.get_next_range ();
 	uint32_t sp_count = 1;

### 2011-08-08 lib/mftp_automaton.cpp
Finding interesting problems in interval set.

@@ -174,8 +174,14 @@ namespace mftp {
 	    ++sp_count;
 	    m_last_request_size += (spans[sp_count].stop - spans[sp_count].start)/FRAGMENT_SIZE;
 	  }
+
+	  std::cout << "Requesting ";
+	  for (uint32_t i = 0; i < sp_count; ++i) {
+	    std::cout << "[" << spans[i].start / FRAGMENT_SIZE << "," << spans[i].stop / FRAGMENT_SIZE << ") ";
+	  }
+	  std::cout << std::endl;
 	}
-	
+
 	message_buffer* m = new message_buffer (request_type (), m_fileid, sp_count, spans);
 	m->convert_to_network ();
 	m_sendq.push (ioa::const_shared_ptr<message_buffer> (m));
@@ -256,17 +262,26 @@ namespace mftp {
 	  // Record the time.
 	  m_fragment_time = ioa::time::now ();
 
+	  // Remove fragment from requests.
+	  uint32_t idx = (m->frag.offset / FRAGMENT_SIZE);
+	  m_requests.erase (std::make_pair (idx, idx + 1));
+
 	  // Save the fragment.
 	  if (!m_file.complete ()) {
-	    if (m_file.write_chunk (m->frag.offset, m->frag.data)) {
+	    std::cout << "Writing " << m->frag.offset / FRAGMENT_SIZE << std::endl;
+	    if (m_file.write_chunk (m->frag.offset, m->frag.data)) {	      
 	      ++m_fragments_since_request;
-	      // Remove fragment from requests.
-	      uint32_t idx = (m->frag.offset / FRAGMENT_SIZE);
-	      m_requests.erase (std::make_pair (idx, idx + 1));
 	      // Reset the request interval.
 	      send_request (true);
 	      ++m_fragments_since_report;
 	    }
+
+	    for (interval_set<uint32_t>::const_iterator pos = m_file.m_dont_have.begin ();
+		 pos != m_file.m_dont_have.end ();
+		 ++pos) {
+	      std::cout << "[" << pos->first << "," << pos->second << ") ";
+	    }
+	    std::cout << std::endl;
 	  }
 	}
 
@@ -311,7 +326,6 @@ namespace mftp {
 		m->req.spans[sp].stop % FRAGMENT_SIZE == 0 &&
 		m->req.spans[sp].start < m->req.spans[sp].stop &&
[…]

### 2011-08-08 lib/mftp_automaton.cpp
Stashing local changes.

@@ -427,6 +427,13 @@ namespace mftp {
     m->convert_to_network ();
     m_sendq.push (ioa::const_shared_ptr<message_buffer> (m));
     ++m_fragment_count;
+
+    for (interval_set<uint32_t>::const_iterator pos = m_requests.begin ();
+	 pos != m_requests.end ();
+	 ++pos) {
+      std::cout << "[" << pos->first << "," << pos->second << ") ";
+    }
+    std::cout << std::endl;
   }  
 
   bool mftp_automaton::set_timer_precondition () const {

### 2011-08-09 include/mftp/file.hpp
Fixed interval tree bugs.

@@ -14,7 +14,6 @@ namespace mftp {
     mfileid m_mfileid;
     uint8_t* m_data;
     uint32_t m_have_count;
-    uint32_t m_start_idx;
 
   public:
     file (const std::string&,
@@ -32,10 +31,9 @@ namespace mftp {
     uint32_t have_count () const;
     bool complete () const;
     bool empty () const;
-    bool write_chunk (const uint32_t offset,
+    bool write_chunk (const uint32_t idx,
 		      const void* data);
-    std::pair<uint32_t, uint32_t> get_next_range ();
-    uint32_t get_random_index () const;
+    uint32_t get_first_fragment_index () const;
   };
 }
 

### 2011-08-09 include/mftp/interval_set.hpp

@@ -94,6 +94,11 @@ public:
 public:
   static bool intersect (const interval_type& x,
 			 const interval_type& y) {
+    return std::max (x.first, y.first) < std::min (x.second, y.second);
+  } 
+
+  static bool touch (const interval_type& x,
+			 const interval_type& y) {
     return std::max (x.first, y.first) <= std::min (x.second, y.second);
   } 
 
@@ -105,26 +110,15 @@ public:
 	return m_set.insert (x);
       }
 
-      iterator pos = m_set.lower_bound (x);
-      if (pos == m_set.end ()) {
-	--pos;
-      }
-      if (!intersect (*pos, x)) {
+      iterator pos = find_first_touch (x);
+      if (pos == end ()) {
 	//nothing intersects x
 	return m_set.insert (x);
       }
-      if (pos != m_set.begin ()) {
-	iterator it = pos;
-	--it;
 
-	if (intersect (*it, x)) {
-	  pos = it;
-	}
-      }
-
-      const Key alpha = std::min (x.first, pos->first); 
+      const Key alpha = std::min (x.first, pos->first);
       Key beta = pos->second;
-      while (pos != m_set.end () && intersect (*pos, x)) {
+      while (pos != m_set.end () && touch (*pos, x)) {
 	beta = pos->second;
 	m_set.erase (pos++);
       }
@@ -152,18 +146,9 @@ public:
       if (empty ()) {
 	return 0;
       }
-      iterator pos = m_set.lower_bound (k);
-      if (pos != m_set.begin ()) {
-	iterator it = pos;
-	--it;
-	
-	if (intersect (*it, k)) {
-	  pos = it;
-	}
-      }
 
-      if (pos == m_set.end () || !intersect (*pos, k)) {
-	//nothing intersects k
+      iterator pos = find_first_intersect (k);
+      if (pos == end ()) {
 	return 0;
       }
 
@@ -209,7 +194,6 @@ public:
   equal_range (const key_type& k) const {
     return m_set.equal_range (k);
   }
-
   
   bool operator== (const interval_set& y) const {
     return m_set == y.m_set;
@@ -218,6 +202,42 @@ public:
   
   bool operator< (const interval_set& y) const {
     return m_set < y.m_set;
+  }
+
+  iterator find_first_intersect (const key_type& k) const {
+    const_iterator after = lower_bound (k);
+    const_iterator before = after;
+    if (before != begin ()) {
+      --before;
+    }
+
+    if (before != end () && intersect (*before, k)) {
+      return before;
+    }
[…]

### 2011-08-09 include/mftp/message.hpp

@@ -15,20 +15,20 @@ namespace mftp {
   struct fragment
   {
     fileid fid;
-    uint32_t offset;
+    uint32_t idx;
     uint8_t data[FRAGMENT_SIZE];
 
     void convert_to_network () {
       fid.convert_to_network ();
-      offset = htonl (offset);
+      idx = htonl (idx);
     }
 
     bool convert_to_host () {
       fid.convert_to_host ();
-      offset = ntohl (offset);
+      idx = ntohl (idx);
 
       mfileid mid (fid);
-      return ((offset % FRAGMENT_SIZE) == 0) && (offset < mid.get_final_length ());
+      return idx < mid.get_fragment_count ();
     }
   };
 
@@ -82,9 +82,7 @@ namespace mftp {
 	spans[i].convert_to_host ();
 	
 	if (!((spans[i].start < spans[i].stop) &&
-	      (spans[i].stop <= mid.get_final_length () ) &&
-	      (spans[i].start % FRAGMENT_SIZE == 0) &&
-	      (spans[i].stop % FRAGMENT_SIZE == 0) )) {
+	      (spans[i].stop <= mid.get_fragment_count ()))) {
 	  return false;
 	}
       }
@@ -153,12 +151,12 @@ namespace mftp {
 
     message (fragment_type /* */,
 	     const fileid& fileid,
-	     uint32_t offset,
+	     uint32_t idx,
 	     const void* data)
     {
       header.message_type = FRAGMENT;
       frag.fid = fileid;
-      frag.offset = offset;
+      frag.idx = idx;
       memcpy (frag.data, data, FRAGMENT_SIZE);
     }
 
@@ -226,9 +224,9 @@ namespace mftp {
 
     message_buffer (fragment_type type,
 		    const fileid& fileid,
-		    uint32_t offset,
+		    uint32_t idx,
 		    const void* data) :
-      msg (type, fileid, offset, data)
+      msg (type, fileid, idx, data)
     { }
 
     message_buffer (request_type type,

### 2011-08-09 include/mftp/mftp_automaton.hpp

@@ -15,11 +15,6 @@ namespace mftp {
     public ioa::automaton
   {
   private:
-    static const ioa::time TIMER_INTERVAL;
-    static const ioa::time INIT_INTERVAL;
-    static const ioa::time MAX_INTERVAL;
-    static const size_t MAX_FRAGMENT_COUNT;
-
     enum send_state_t {
       SEND_READY,
       SEND_COMPLETE_WAIT
@@ -30,46 +25,64 @@ namespace mftp {
       INTERRUPT_WAIT,
     };
 
-    #define REQUEST_NUMERATOR 4
-    #define REQUEST_DENOMINATOR 5
+    static const ioa::time TIMER_INTERVAL;
+    static const ioa::time INIT_INTERVAL;
+    static const ioa::time MAX_INTERVAL;
+    static const uint32_t MAX_FRAGMENT_COUNT;
+
+    static const uint32_t REQUEST_NUMERATOR;
+    static const uint32_t REQUEST_DENOMINATOR;
     
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;
     const mfileid& m_mfileid;
     const fileid& m_fileid;
-    interval_set<uint32_t> m_requests; // Set of intervals indicating fragments that have been requested.
-    uint32_t m_last_sent_idx; // Index of last fragment sent.
+
+    ioa::handle_manager<mftp_channel_automaton> m_channel; // The channel for sending/receiving.
+
+    // Sending.
     std::queue<ioa::const_shared_ptr<message_buffer> > m_sendq; // Send queue.
     send_state_t m_send_state; // State of send state machine.
-    size_t m_fragment_count; // Number of fragments in the send queue.
-    ioa::time m_fragment_time; // Time when this automaton last received a fragment (of this file).
-    ioa::time m_request_time; // Time when this automaton last sent a request.
+    uint32_t m_num_frag_in_sendq; // Number of fragments in the send queue.
+    uint32_t m_num_req_in_sendq; // Number of requests in the send queue.
+    uint32_t m_num_match_in_sendq; // Number of matches in the send queue.
+
+    // Answering requests.
+    interval_set<uint32_t> m_requests; // Set of intervals indicating fragments that have been requested.
+    uint32_t m_last_sent_idx; // Index of last fragment sent.
+
+    // Timestamps for certain events.
+    ioa::time m_frag_recv_time; // Time when this automaton last received a fragment (of this file).
[…]

### 2011-08-09 lib/file.cpp

@@ -14,8 +14,7 @@
 namespace mftp {
 
   file::file (const std::string& name,
-	      const uint32_t type) :
-    m_start_idx (0)
+	      const uint32_t type)
   {
     int fd = open (name.c_str (), O_RDONLY);
     assert (fd != -1);
@@ -66,8 +65,7 @@ namespace mftp {
   file::file (const fileid& f) :
     m_mfileid (f),
     m_data (new unsigned char[m_mfileid.get_final_length ()]),
-    m_have_count (0),
-    m_start_idx (0)
+    m_have_count (0)
   { 
     m_dont_have.insert (std::make_pair (0, m_mfileid.get_fragment_count ()));
   }
@@ -75,16 +73,14 @@ namespace mftp {
   file::file (const file& other) :
     m_dont_have (other.m_dont_have),
     m_mfileid (other.m_mfileid),
-    m_have_count (other.m_have_count),
-    m_start_idx (0)
+    m_have_count (other.m_have_count)
   {
     const uint32_t final_length = m_mfileid.get_final_length ();
     m_data = new unsigned char[final_length];
     memcpy (m_data, other.m_data, final_length);
   }
 
-  file::file (const void* ptr, uint32_t size, uint32_t type) :
-    m_start_idx (0)
+  file::file (const void* ptr, uint32_t size, uint32_t type)
   {
     m_mfileid.set_length (size);
     m_mfileid.set_type (type);
@@ -134,37 +130,16 @@ namespace mftp {
     return m_have_count == 0;
   }
 
-  // bool file::have (const uint32_t offset) const {
-  //   assert (offset % FRAGMENT_SIZE == 0);
-  //   assert (offset < m_mfileid.get_final_length ());
-  //   const uint32_t idx = offset / FRAGMENT_SIZE;
-  //   interval_set<uint32_t>::const_iterator pos = m_dont_have.lower_bound (std::make_pair (idx, idx + 1));
-  //   if (pos != m_dont_have.end () && pos->first <= offset) {
-  //     return false;
-  //   }
-  //   else {
-  //     return true;
-  //   }
-  // }
-
-  bool file::write_chunk (const uint32_t offset,
+  bool file::write_chunk (const uint32_t idx,
 			  const void* data) {
-    assert (offset % FRAGMENT_SIZE == 0);
-    assert (offset < m_mfileid.get_final_length ());
+    assert (idx < m_mfileid.get_fragment_count ());
 
[…]

### 2011-08-09 lib/mftp_automaton.cpp

@@ -4,7 +4,10 @@ namespace mftp {
   const ioa::time mftp_automaton::TIMER_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
-  const size_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
+  const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
+  const uint32_t mftp_automaton::REQUEST_NUMERATOR (4); // Send request when we have receive 80% of the previous request.
+  const uint32_t mftp_automaton::REQUEST_DENOMINATOR (5);
+
 
   // Not matching.
   mftp_automaton::mftp_automaton (const file& file,
@@ -15,22 +18,24 @@ namespace mftp {
     m_file (file),
     m_mfileid (file.get_mfileid ()),
     m_fileid (m_mfileid.get_fileid ()),
-    m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
+    m_channel (channel),
     m_send_state (SEND_READY),
-    m_fragment_count (0),
+    m_num_frag_in_sendq (0),
+    m_num_req_in_sendq (0),
+    m_num_match_in_sendq (0),
+    m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
     m_timer_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_request_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
-    m_reported (m_file.complete ()),
-    m_last_request_size (0),
-    m_fragments_since_request (0),
-    m_channel (channel),
+    m_num_frags_in_last_req (0),
+    m_num_new_frags_since_req (0),
+    m_fragments_since_report (0),
+    m_progress_threshold (progress_threshold),
     m_matching (false),
     m_get_matching_files (false),
     m_suicide_flag (suicide),
-    m_fragments_since_report (0),
-    m_progress_threshold (progress_threshold)
+    m_reported (m_file.complete ())
   {
     create_bindings ();
   }
@@ -47,24 +52,26 @@ namespace mftp {
     m_file (file),
     m_mfileid (file.get_mfileid ()),
[…]

### 2011-08-09 lib/mftp_channel_automaton.cpp

@@ -84,6 +84,7 @@ namespace mftp {
 
   void mftp_channel_automaton::send_effect (const ioa::const_shared_ptr<message_buffer>& message,
 					    ioa::aid_t aid) {
+    
     if (m_outgoing_set.count (aid) == 0 &&
 	m_pending_aid != aid &&
 	m_outgoing_completes.count (aid) == 0) {

### 2011-08-09 test/interval_set.cpp

@@ -4,6 +4,7 @@
 #include <list>
 #include <algorithm>
 #include <iostream>
+#include <vector>
 
 static const char* default_ctor () {
   std::cout << __func__ << std::endl;
@@ -417,6 +418,69 @@ static const char* erase_triple_intersect12 () {
   return 0;  
 }
 
+static const char* erase () {
+  std::cout << __func__ << std::endl;
+  
+  interval_set<int> tr;
+  std::vector<bool> q (100);
+
+  for (int round = 0; round < 1000; ++round) {
+    int low = rand () % 100;
+    int high = rand () % 100;
+
+    if (high < low) {
+      int tmp = low;
+      low = high;
+      high = tmp;
+    }
+
+    interval_set<int>::interval_type interval (low, high);
+
+    if (rand () % 2 == 0) {
+      std::cout << "insert [" << interval.first << "," << interval.second << ")" << std::endl;
+      tr.insert (interval);
+      for (int i = low; i < high; ++i) {
+	q[i] = true;
+      }
+    }
+    else {
+      std::cout << "erase [" << interval.first << "," << interval.second << ")" << std::endl;
+      tr.erase (interval);
+      for (int i = low; i < high; ++i) {
+	q[i] = false;
+      }
+    }
+
+    for (interval_set<int>::const_iterator pos = tr.begin ();
+	 pos != tr.end ();
+	 ++pos) {
+      interval_set<int>::const_iterator nxt = pos;
+      ++nxt;
+      if (nxt != tr.end ()) {
+	mu_assert (!interval_set<int>::touch (*pos, *nxt));
+      }
+
+      std::cout << "[" << pos->first << "," << pos->second << ") ";
+      for (int i = pos->first; i < pos->second; ++i) {
+	mu_assert (q[i]);
+      }
+    }
+    std::cout << std::endl;
+
+    for (int i = 0; i < 100; ++i) {
+      if (q[i]) {
+	mu_assert (tr.find_first_intersect (std::make_pair (i, i + 1)) != tr.end ());
+      }
+      else {
+	mu_assert (tr.find_first_intersect (std::make_pair (i, i + 1)) == tr.end ());
+      }
+    }
+
+  }
+
+  return 0;  
+}
+
 const char* all_tests () {
   mu_run_test (default_ctor);
   mu_run_test (insert_single);
@@ -445,6 +509,8 @@ const char* all_tests () {
   mu_run_test (erase_triple_intersect);
[…]

### 2011-08-09 include/mftp/message.hpp
Added feedback for sending.

@@ -9,7 +9,7 @@ namespace mftp {
   const uint32_t REQUEST = 1;
   const uint32_t MATCH = 2;
 
-  const uint32_t SPANS_SIZE = 64;
+  const uint32_t SPANS_SIZE = 63;
   const uint32_t MATCHES_SIZE = 12;
 
   struct fragment
@@ -58,11 +58,13 @@ namespace mftp {
   struct request
   {
     fileid fid;
+    uint32_t frags_per_second;
     uint32_t span_count;
     span_t spans[SPANS_SIZE];
     
     void convert_to_network () {
       fid.convert_to_network ();
+      frags_per_second = htonl (frags_per_second);
       for (uint32_t i = 0; i < span_count; ++i) {
 	spans[i].convert_to_network ();
       }
@@ -71,6 +73,7 @@ namespace mftp {
     
     bool convert_to_host () {
       fid.convert_to_host ();
+      frags_per_second = ntohl (frags_per_second);
       span_count = ntohl (span_count);
       
       if (span_count == 0 || span_count > SPANS_SIZE) {
@@ -162,11 +165,13 @@ namespace mftp {
 
     message (request_type /* */,
 	     const fileid& fileid,
+	     uint32_t frags_per_second,
 	     uint32_t span_count,
 	     const span_t * spans)
     {
       header.message_type = REQUEST;
       req.fid = fileid;
+      req.frags_per_second = frags_per_second;
       req.span_count = span_count;
       for (uint32_t i = 0; i<span_count; i++){
 	req.spans[i] = spans[i];
@@ -231,9 +236,10 @@ namespace mftp {
 
     message_buffer (request_type type,
 		    const fileid& fileid,
+		    uint32_t frags_per_second,
 		    uint32_t span_count,
 		    const span_t * spans) :
-      msg (type, fileid, span_count, spans)
+      msg (type, fileid, frags_per_second, span_count, spans)
     { }
 
     message_buffer (match_type type,

### 2011-08-09 include/mftp/mftp_automaton.hpp

@@ -20,18 +20,22 @@ namespace mftp {
       SEND_COMPLETE_WAIT
     };
 
-    enum timer_state_t {
+    enum alarm_state_t {
       SET_READY,
       INTERRUPT_WAIT,
     };
 
-    static const ioa::time TIMER_INTERVAL;
+    static const ioa::time ALARM_INTERVAL;
     static const ioa::time INIT_INTERVAL;
     static const ioa::time MAX_INTERVAL;
     static const uint32_t MAX_FRAGMENT_COUNT;
+    static const ioa::time REQUEST_INTERVAL;
 
-    static const uint32_t REQUEST_NUMERATOR;
-    static const uint32_t REQUEST_DENOMINATOR;
+    static const uint32_t REREQUEST_NUMERATOR;
+    static const uint32_t REREQUEST_DENOMINATOR;
+
+    static const uint32_t RATE_INCREASE_NUMERATOR;
+    static const uint32_t RATE_INCREASE_DENOMINATOR;
     
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;
@@ -46,6 +50,7 @@ namespace mftp {
     uint32_t m_num_frag_in_sendq; // Number of fragments in the send queue.
     uint32_t m_num_req_in_sendq; // Number of requests in the send queue.
     uint32_t m_num_match_in_sendq; // Number of matches in the send queue.
+    uint32_t m_target_send_frag_per_second;
 
     // Answering requests.
     interval_set<uint32_t> m_requests; // Set of intervals indicating fragments that have been requested.
@@ -57,12 +62,15 @@ namespace mftp {
     ioa::time m_match_time; // Time when this automaton last received a match (of this file).
 
     // Periodic acitivities.
-    timer_state_t m_timer_state; // State of time state machine.
+    alarm_state_t m_alarm_state; // State of alarm state machine.
+    alarm_state_t m_fragment_alarm_state; // State of fragment alarm state machine.
     ioa::time m_announcement_interval; // Must wait this amount of time after m_fragment_time to send an announcement.
-    ioa::time m_request_interval; // Must wait this amount of time after m_request_time to send a request.
     ioa::time m_match_interval; // Must wait this amount of time after m_match_time to send a match.
 
     // Sending requests.
+    // TODO:  Message should be fragment.
+    uint32_t m_num_msg_in_last_interval; // Number of messages received in last interval.
[…]

### 2011-08-09 lib/mftp_automaton.cpp

@@ -1,13 +1,16 @@
 #include <mftp/mftp_automaton.hpp>
 
 namespace mftp {
-  const ioa::time mftp_automaton::TIMER_INTERVAL (1, 0); // 1 second
+  const ioa::time mftp_automaton::ALARM_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
   const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
-  const uint32_t mftp_automaton::REQUEST_NUMERATOR (4); // Send request when we have receive 80% of the previous request.
-  const uint32_t mftp_automaton::REQUEST_DENOMINATOR (5);
+  const uint32_t mftp_automaton::REREQUEST_NUMERATOR (4); // Send request when we have receive 80% of the previous request.
+  const uint32_t mftp_automaton::REREQUEST_DENOMINATOR (5);
+  const ioa::time mftp_automaton::REQUEST_INTERVAL (INIT_INTERVAL);
 
+  const uint32_t mftp_automaton::RATE_INCREASE_NUMERATOR (1); // Increase the request rate this must percent per second.
+  const uint32_t mftp_automaton::RATE_INCREASE_DENOMINATOR (100);
 
   // Not matching.
   mftp_automaton::mftp_automaton (const file& file,
@@ -23,11 +26,14 @@ namespace mftp {
     m_num_frag_in_sendq (0),
     m_num_req_in_sendq (0),
     m_num_match_in_sendq (0),
+    m_target_send_frag_per_second (-1),
     m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
-    m_timer_state (SET_READY),
+    m_alarm_state (SET_READY),
+    m_fragment_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
-    m_request_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
+    m_num_msg_in_last_interval (0),
+    m_target_msg_per_second (0),
     m_num_frags_in_last_req (0),
     m_num_new_frags_since_req (0),
     m_fragments_since_report (0),
@@ -57,11 +63,14 @@ namespace mftp {
     m_num_frag_in_sendq (0),
[…]

### 2011-08-09 papers/paper1/paper.tex

@@
 \end{abstract}
+
+\section{Adam's Section}
+A couple of things were hard to digest, the first of which being indirect triggers.  
+It's hard to swallow the idea that you can queue an action but it's time of execution is far from guaranteed.  
+The second hardest thing was simply the intimidation factor of realizing your entire summer would be working with an unknown, rather large library that nobody but the creator had used before--which was modelled after a complicated formalism I had never heard of that no one had bothered to fully develop since its inception.
+That's quite the pill to swallow.
+
+An thing that I learned was easy was to model the automata very closely to Lynch's design.  
+It was much easier for me to learn the relationship between the formalism and the code by looking at well-modelled automata, so the same would extend onto someone looking at the formalism and our code.
+Maintaining this allows for a less steep learning curve, and hopefully contributes to the automata maintaining their structural integrity as more people use the library.
+
+It's hard to describe what is ``harder'' to implement using I/O Automata.  
+Though I/O Automata present a \emph{different} problem set than using threads, the resulting problem set is respectively easier and more intuitive to handle.
+
+However, what is easer to implement is much more plentiful.
+The design and debugging of the program is much more intuitive.
+It allowed for a high degree of abstraction, which is key for easily-understandable code.
+It allowed \emph{extremely} easy refactoring of any class when compared to the alleged dregs of refactoring a thread-intensive class.
+
+In regards to building a protocol, I learned that it is extremely helpful to add and then test small bits of functionality at a time, in order to get a small working product.
+Once this is complete, repeat by adding or extending the functionality.
+Though this may seem trivial, calculated baby-steps made developing a protocol seem much easier than trying it all at once.
 

### 2011-08-09 include/mftp/mftp_automaton.hpp
Added exponentiall weighted moving average.

@@ -36,6 +36,9 @@ namespace mftp {
 
     static const uint32_t RATE_INCREASE_NUMERATOR;
     static const uint32_t RATE_INCREASE_DENOMINATOR;
+
+    static const uint32_t RATE_ALPHA;
+    static const uint32_t RATE_BETA;
     
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;

### 2011-08-09 lib/mftp_automaton.cpp

@@ -5,12 +5,16 @@ namespace mftp {
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
   const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
-  const uint32_t mftp_automaton::REREQUEST_NUMERATOR (4); // Send request when we have receive 80% of the previous request.
-  const uint32_t mftp_automaton::REREQUEST_DENOMINATOR (5);
-  const ioa::time mftp_automaton::REQUEST_INTERVAL (INIT_INTERVAL);
+  const uint32_t mftp_automaton::REREQUEST_NUMERATOR (9); // Send request when we have received this percentage of the previous request.  Reflects our assumption about the network.
+  const uint32_t mftp_automaton::REREQUEST_DENOMINATOR (10);
+  const ioa::time mftp_automaton::REQUEST_INTERVAL (INIT_INTERVAL); // Send a request using this frequency.
 
-  const uint32_t mftp_automaton::RATE_INCREASE_NUMERATOR (1); // Increase the request rate this must percent per second.
+  const uint32_t mftp_automaton::RATE_INCREASE_NUMERATOR (1); // Inflate the request rate by this percent.
   const uint32_t mftp_automaton::RATE_INCREASE_DENOMINATOR (100);
+
+  // Constants for exponentially weighted moving average.
+  const uint32_t mftp_automaton::RATE_ALPHA (4);
+  const uint32_t mftp_automaton::RATE_BETA (4);
 
   // Not matching.
   mftp_automaton::mftp_automaton (const file& file,
@@ -360,8 +364,7 @@ namespace mftp {
 	  if (!m_requests.empty () &&
 	      m->req.frags_per_second != 0) {
 	    if (m_target_send_frag_per_second != static_cast<uint32_t> (-1)) {
-	      // TODO:  Get rid of constants.
-	      m_target_send_frag_per_second = (m_target_send_frag_per_second * 4 + m->req.frags_per_second * 1) / 5;
+	      m_target_send_frag_per_second = (m_target_send_frag_per_second * RATE_ALPHA + m->req.frags_per_second * RATE_BETA) / (RATE_ALPHA + RATE_BETA);
 	    }
 	    else {
[…]

### 2011-08-10 include/mftp/message.hpp
Intermediate commit for testing.

@@ -58,13 +58,11 @@ namespace mftp {
   struct request
   {
     fileid fid;
-    uint32_t frags_per_second;
     uint32_t span_count;
     span_t spans[SPANS_SIZE];
     
     void convert_to_network () {
       fid.convert_to_network ();
-      frags_per_second = htonl (frags_per_second);
       for (uint32_t i = 0; i < span_count; ++i) {
 	spans[i].convert_to_network ();
       }
@@ -73,7 +71,6 @@ namespace mftp {
     
     bool convert_to_host () {
       fid.convert_to_host ();
-      frags_per_second = ntohl (frags_per_second);
       span_count = ntohl (span_count);
       
       if (span_count == 0 || span_count > SPANS_SIZE) {
@@ -165,13 +162,11 @@ namespace mftp {
 
     message (request_type /* */,
 	     const fileid& fileid,
-	     uint32_t frags_per_second,
 	     uint32_t span_count,
 	     const span_t * spans)
     {
       header.message_type = REQUEST;
       req.fid = fileid;
-      req.frags_per_second = frags_per_second;
       req.span_count = span_count;
       for (uint32_t i = 0; i<span_count; i++){
 	req.spans[i] = spans[i];
@@ -236,10 +231,9 @@ namespace mftp {
 
     message_buffer (request_type type,
 		    const fileid& fileid,
-		    uint32_t frags_per_second,
 		    uint32_t span_count,
 		    const span_t * spans) :
-      msg (type, fileid, frags_per_second, span_count, spans)
+      msg (type, fileid, span_count, spans)
     { }
 
     message_buffer (match_type type,

### 2011-08-10 include/mftp/mftp_automaton.hpp

@@ -37,9 +37,6 @@ namespace mftp {
     static const uint32_t RATE_INCREASE_NUMERATOR;
     static const uint32_t RATE_INCREASE_DENOMINATOR;
 
-    static const uint32_t RATE_ALPHA;
-    static const uint32_t RATE_BETA;
-    
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;
     const mfileid& m_mfileid;
@@ -53,7 +50,6 @@ namespace mftp {
     uint32_t m_num_frag_in_sendq; // Number of fragments in the send queue.
     uint32_t m_num_req_in_sendq; // Number of requests in the send queue.
     uint32_t m_num_match_in_sendq; // Number of matches in the send queue.
-    uint32_t m_target_send_frag_per_second;
 
     // Answering requests.
     interval_set<uint32_t> m_requests; // Set of intervals indicating fragments that have been requested.
@@ -71,9 +67,6 @@ namespace mftp {
     ioa::time m_match_interval; // Must wait this amount of time after m_match_time to send a match.
 
     // Sending requests.
-    // TODO:  Message should be fragment.
-    uint32_t m_num_msg_in_last_interval; // Number of messages received in last interval.
-    uint32_t m_target_msg_per_second;  // Target number of messages per second.
     uint32_t m_num_frags_in_last_req; // Number of fragments in most recent request.
     uint32_t m_num_new_frags_since_req; // New fragments received since most recent request.
 

### 2011-08-10 lib/mftp_automaton.cpp

@@ -12,10 +12,6 @@ namespace mftp {
   const uint32_t mftp_automaton::RATE_INCREASE_NUMERATOR (1); // Inflate the request rate by this percent.
   const uint32_t mftp_automaton::RATE_INCREASE_DENOMINATOR (100);
 
-  // Constants for exponentially weighted moving average.
-  const uint32_t mftp_automaton::RATE_ALPHA (4);
-  const uint32_t mftp_automaton::RATE_BETA (4);
-
   // Not matching.
   mftp_automaton::mftp_automaton (const file& file,
 				  const ioa::automaton_handle<mftp_channel_automaton>& channel,
@@ -30,14 +26,11 @@ namespace mftp {
     m_num_frag_in_sendq (0),
     m_num_req_in_sendq (0),
     m_num_match_in_sendq (0),
-    m_target_send_frag_per_second (-1),
     m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
     m_alarm_state (SET_READY),
     m_fragment_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
-    m_num_msg_in_last_interval (0),
-    m_target_msg_per_second (0),
     m_num_frags_in_last_req (0),
     m_num_new_frags_since_req (0),
     m_fragments_since_report (0),
@@ -67,14 +60,11 @@ namespace mftp {
     m_num_frag_in_sendq (0),
     m_num_req_in_sendq (0),
     m_num_match_in_sendq (0),
-    m_target_send_frag_per_second (-1),
     m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
     m_alarm_state (SET_READY),
     m_fragment_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
-    m_num_msg_in_last_interval (0),
-    m_target_msg_per_second (0),
     m_num_frags_in_last_req (0),
     m_num_new_frags_since_req (0),
     m_fragments_since_report (0),
@@ -203,7 +193,7 @@ namespace mftp {
 	  ++pos;
 	}
 
-	message_buffer* m = new message_buffer (request_type (), m_fileid, m_target_msg_per_second, span_count, spans);
+	message_buffer* m = new message_buffer (request_type (), m_fileid, span_count, spans);
[…]

### 2011-08-10 include/mftp/mftp_automaton.hpp
Intermediate commit.

@@ -62,7 +62,6 @@ namespace mftp {
 
     // Periodic acitivities.
     alarm_state_t m_alarm_state; // State of alarm state machine.
-    alarm_state_t m_fragment_alarm_state; // State of fragment alarm state machine.
     ioa::time m_announcement_interval; // Must wait this amount of time after m_fragment_time to send an announcement.
     ioa::time m_match_interval; // Must wait this amount of time after m_match_time to send a match.
 
@@ -87,6 +86,8 @@ namespace mftp {
     // Termination.
     bool m_suicide_flag;  // Self-destruct when job is done.
     bool m_reported; // True when we have reported a complete download.
+
+    uint32_t m_frags_sent;
 
   public:
     // Not matching.
@@ -131,12 +132,9 @@ namespace mftp {
     void alarm_interrupt_effect ();
     UV_UP_INPUT (mftp_automaton, alarm_interrupt);
 
-    bool set_fragment_alarm_precondition () const;
-    ioa::time set_fragment_alarm_effect ();
-    V_UP_OUTPUT (mftp_automaton, set_fragment_alarm, ioa::time);
-
-    void fragment_alarm_interrupt_effect ();
-    UV_UP_INPUT (mftp_automaton, fragment_alarm_interrupt);
+    bool send_fragment_precondition () const;
+    void send_fragment_effect ();
+    UP_INTERNAL (mftp_automaton, send_fragment);
 
     bool suicide_precondition () const;
     void suicide_effect ();

### 2011-08-10 lib/mftp_automaton.cpp

@@ -28,7 +28,6 @@ namespace mftp {
     m_num_match_in_sendq (0),
     m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
     m_alarm_state (SET_READY),
-    m_fragment_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
     m_num_frags_in_last_req (0),
@@ -38,7 +37,8 @@ namespace mftp {
     m_matching (false),
     m_get_matching_files (false),
     m_suicide_flag (suicide),
-    m_reported (m_file.complete ())
+    m_reported (m_file.complete ()),
+    m_frags_sent (0)
   {
     create_bindings ();
   }
@@ -62,7 +62,6 @@ namespace mftp {
     m_num_match_in_sendq (0),
     m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
     m_alarm_state (SET_READY),
-    m_fragment_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
     m_num_frags_in_last_req (0),
@@ -74,7 +73,8 @@ namespace mftp {
     m_match_predicate (match_pred.clone ()),
     m_get_matching_files (get_matching_files),
     m_suicide_flag (suicide),
-    m_reported (m_file.complete ())
+    m_reported (m_file.complete ()),
+    m_frags_sent (0)
   {
     create_bindings ();
   }
@@ -105,18 +105,6 @@ namespace mftp {
 			       &m_self,
 			       &mftp_automaton::alarm_interrupt);
 
-    ioa::automaton_manager<ioa::alarm_automaton>* fragment_alarm = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
-    ioa::make_binding_manager (this,
-			       &m_self,
-			       &mftp_automaton::set_fragment_alarm,
-			       fragment_alarm,
-			       &ioa::alarm_automaton::set);
-    ioa::make_binding_manager (this,
-			       fragment_alarm,
-			       &ioa::alarm_automaton::alarm,
-			       &m_self,
-			       &mftp_automaton::fragment_alarm_interrupt);
-
     send_announcement ();
     send_request ();
     schedule ();
@@ -129,8 +117,8 @@ namespace mftp {
     if (set_alarm_precondition ()) {
[…]

### 2011-08-10 include/mftp/mftp_automaton.hpp
Requesting one fragment at a time.

@@ -29,13 +29,6 @@ namespace mftp {
     static const ioa::time INIT_INTERVAL;
     static const ioa::time MAX_INTERVAL;
     static const uint32_t MAX_FRAGMENT_COUNT;
-    static const ioa::time REQUEST_INTERVAL;
-
-    static const uint32_t REREQUEST_NUMERATOR;
-    static const uint32_t REREQUEST_DENOMINATOR;
-
-    static const uint32_t RATE_INCREASE_NUMERATOR;
-    static const uint32_t RATE_INCREASE_DENOMINATOR;
 
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;
@@ -57,7 +50,6 @@ namespace mftp {
 
     // Timestamps for certain events.
     ioa::time m_frag_recv_time; // Time when this automaton last received a fragment (of this file).
-    ioa::time m_request_time; // Time when this automaton last sent a request or received a new fragment.
     ioa::time m_match_time; // Time when this automaton last received a match (of this file).
 
     // Periodic acitivities.

### 2011-08-10 lib/mftp_automaton.cpp

@@ -5,12 +5,6 @@ namespace mftp {
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
   const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
-  const uint32_t mftp_automaton::REREQUEST_NUMERATOR (9); // Send request when we have received this percentage of the previous request.  Reflects our assumption about the network.
-  const uint32_t mftp_automaton::REREQUEST_DENOMINATOR (10);
-  const ioa::time mftp_automaton::REQUEST_INTERVAL (INIT_INTERVAL); // Send a request using this frequency.
-
-  const uint32_t mftp_automaton::RATE_INCREASE_NUMERATOR (1); // Inflate the request rate by this percent.
-  const uint32_t mftp_automaton::RATE_INCREASE_DENOMINATOR (100);
 
   // Not matching.
   mftp_automaton::mftp_automaton (const file& file,
@@ -154,38 +148,46 @@ namespace mftp {
   
   void mftp_automaton::send_request () {
 
-    const ioa::time now = ioa::time::now ();
-
     // We have requests to send.
     // There are no requests in the sendq.
     // Enough time has elapsed since we sent a request or received a new fragment.
     // We have received some fraction of the fragments that we last requested.
     if (!m_file.complete () &&
 	m_num_req_in_sendq == 0 &&
-	((m_request_time + REQUEST_INTERVAL <= now) ||
-	 (REREQUEST_DENOMINATOR * m_num_new_frags_since_req > REREQUEST_NUMERATOR * m_num_frags_in_last_req))) {
+	m_num_new_frags_since_req == m_num_frags_in_last_req) {
+
       // Reset.
-      m_request_time = now;
-      m_num_frags_in_last_req = 0;
       m_num_new_frags_since_req = 0;
+      m_num_frags_in_last_req = 0;
 
-      // Turn the fragments we don't have into requests.
-      interval_set<uint32_t>::const_iterator pos = m_file.m_dont_have.begin ();
-      while (pos != m_file.m_dont_have.end ()) {
-	span_t spans[SPANS_SIZE];
-	uint32_t span_count = 0;
+      span_t spans[SPANS_SIZE];
[…]

### 2011-08-10 include/mftp/mftp_automaton.hpp
Increasing request rate.

@@ -28,6 +28,7 @@ namespace mftp {
     static const ioa::time ALARM_INTERVAL;
     static const ioa::time INIT_INTERVAL;
     static const ioa::time MAX_INTERVAL;
+    static const ioa::time REQUEST_INTERVAL;
     static const uint32_t MAX_FRAGMENT_COUNT;
 
     ioa::handle_manager<mftp_automaton> m_self;
@@ -50,6 +51,7 @@ namespace mftp {
 
     // Timestamps for certain events.
     ioa::time m_frag_recv_time; // Time when this automaton last received a fragment (of this file).
+    ioa::time m_request_time; // Time when this automaton last sent a request.
     ioa::time m_match_time; // Time when this automaton last received a match (of this file).
 
     // Periodic acitivities.
@@ -58,6 +60,7 @@ namespace mftp {
     ioa::time m_match_interval; // Must wait this amount of time after m_match_time to send a match.
 
     // Sending requests.
+    uint32_t m_request_count; // Number of fragments to request.
     uint32_t m_num_frags_in_last_req; // Number of fragments in most recent request.
     uint32_t m_num_new_frags_since_req; // New fragments received since most recent request.
 

### 2011-08-10 lib/mftp_automaton.cpp

@@ -4,6 +4,7 @@ namespace mftp {
   const ioa::time mftp_automaton::ALARM_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
+  const ioa::time mftp_automaton::REQUEST_INTERVAL (INIT_INTERVAL);
   const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
 
   // Not matching.
@@ -24,6 +25,7 @@ namespace mftp {
     m_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
+    m_request_count (1),
     m_num_frags_in_last_req (0),
     m_num_new_frags_since_req (0),
     m_fragments_since_report (0),
@@ -58,6 +60,7 @@ namespace mftp {
     m_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
+    m_request_count (1),
     m_num_frags_in_last_req (0),
     m_num_new_frags_since_req (0),
     m_fragments_since_report (0),
@@ -153,41 +156,56 @@ namespace mftp {
     // Enough time has elapsed since we sent a request or received a new fragment.
     // We have received some fraction of the fragments that we last requested.
     if (!m_file.complete () &&
-	m_num_req_in_sendq == 0 &&
-	m_num_new_frags_since_req == m_num_frags_in_last_req) {
+	m_num_req_in_sendq == 0) {
 
-      // Reset.
-      m_num_new_frags_since_req = 0;
-      m_num_frags_in_last_req = 0;
+      ioa::time now = ioa::time::now ();
 
-      span_t spans[SPANS_SIZE];
-      spans[0].start = m_file.m_dont_have.begin ()->first;
-      spans[0].stop = spans[0].start + 1;
-      std::cout << "Requesting [" << spans[0].start << "," << spans[0].stop << ")" << std::endl;
-      m_num_frags_in_last_req = 1;
+      const bool frag_flag = 2 * m_num_new_frags_since_req >= 1 * m_num_frags_in_last_req;
+      const bool time_flag = m_request_time + REQUEST_INTERVAL <= now;
 
[…]

### 2011-08-10 lib/mftp_automaton.cpp
Test programs.

@@ -4,7 +4,7 @@ namespace mftp {
   const ioa::time mftp_automaton::ALARM_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
-  const ioa::time mftp_automaton::REQUEST_INTERVAL (INIT_INTERVAL);
+  const ioa::time mftp_automaton::REQUEST_INTERVAL (1, 0);
   const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
 
   // Not matching.
@@ -163,18 +163,19 @@ namespace mftp {
       const bool frag_flag = 2 * m_num_new_frags_since_req >= 1 * m_num_frags_in_last_req;
       const bool time_flag = m_request_time + REQUEST_INTERVAL <= now;
 
-      if (frag_flag) {
-	std::cout << "Increase" << std::endl;
-	++m_request_count;
-      }
-      else if (time_flag) {
-	std::cout << "Reduce" << std::endl;
-	m_request_count /= 2;
-      }
-
-      m_request_count = std::max (m_request_count, 1u);
-
       if (frag_flag || time_flag) {
+
+	if (frag_flag) {
+	  std::cout << "Increase" << std::endl;
+	  ++m_request_count;
+	}
+	else if (time_flag) {
+	  std::cout << "Reduce" << std::endl;
+	  m_request_count /= 2;
+	}
+	
+	m_request_count = std::max (m_request_count, 1u);
+	
 	// Reset.
 	m_num_new_frags_since_req = 0;
 	m_num_frags_in_last_req = 0;

### 2011-08-10 src/udp_receiver.cpp

@@ -1,0 +1,66 @@
+#include <ioa/ioa.hpp>
+#include <ioa/udp_receiver_automaton.hpp>
+#include <ioa/global_fifo_scheduler.hpp>
+
+#include <config.hpp>
+
+#include <iostream>
+
+class udp_receiver :
+  public ioa::automaton
+{
+private:
+  ioa::handle_manager<udp_receiver> m_self;
+  ioa::inet_address m_group;
+  ioa::inet_address m_address;
+
+public:
+  udp_receiver (const ioa::inet_address& group,
+		const ioa::inet_address& address) :
+    m_self (ioa::get_aid ()),
+    m_group (group),
+    m_address (address)
+  {
+    ioa::automaton_manager<ioa::udp_receiver_automaton>* receiver = new ioa::automaton_manager<ioa::udp_receiver_automaton> (this, ioa::make_generator<ioa::udp_receiver_automaton> (m_group, m_address));
+
+    ioa::make_binding_manager (this, receiver, &ioa::udp_receiver_automaton::receive, &m_self, &udp_receiver::receive);
+  }
+
+private:
+  void schedule () const { }
+
+  void receive_effect (const ioa::udp_receiver_automaton::receive_val& v) {
+    if (v.err != 0) {
+      char buf[256];
+#ifdef STRERROR_R_CHAR_P
+      std::cerr << "Couldn't receive udp_receiver_automaton: " << strerror_r (v.err, buf, 256) << std::endl;
+#else
+      strerror_r (v.err, buf, 256);
+      std::cerr << "Couldn't receive udp_receiver_automaton: " << buf << std::endl;
+#endif
+      self_destruct ();
+    }
+    else {
+      assert (v.buffer.get () != 0);
+      uint32_t idx;
+      memcpy (&idx, v.buffer->data (), sizeof (idx));
+      idx = ntohl (idx);
+      std::cout << idx << std::endl;
+    }
+  }
+  
+  V_UP_INPUT (udp_receiver, receive, ioa::udp_receiver_automaton::receive_val);
+
+};
+
+int
+main (int argc, char* argv[]) {
+  ioa::inet_address group ("224.0.0.137");
+  ioa::inet_address address ("0.0.0.0", 54321);
+
+  ioa::global_fifo_scheduler ss;
+  ioa::run (ss, ioa::make_generator<udp_receiver> (group, address));
+
+  return 0; 
+}
+

### 2011-08-10 src/udp_sender.cpp

@@ -1,0 +1,148 @@
+#include <iostream>
+#include <ioa/ioa.hpp>
+#include <ioa/global_fifo_scheduler.hpp>
+#include <ioa/alarm_automaton.hpp>
+#include <ioa/udp_sender_automaton.hpp>
+#include <queue>
+#include <config.hpp>
+
+class udp_sender :
+  public ioa::automaton
+{
+private:
+  enum alarm_state_t {
+    ALARM_SET,
+    ALARM_WAIT
+  };
+
+  enum send_state_t {
+    SEND_SEND,
+    SEND_WAIT
+  };
+
+  ioa::handle_manager<udp_sender> m_self;
+  alarm_state_t m_alarm_state;
+  send_state_t m_send_state;
+  const ioa::inet_address m_address;
+  const ioa::time m_interval;
+  const uint32_t m_message_size;
+  uint32_t m_message_count;
+  std::queue<ioa::const_shared_ptr<ioa::buffer> > m_sendq;
+
+public:
+  udp_sender (const ioa::inet_address& address,
+	      const uint32_t message_rate,
+	      const uint32_t message_size,
+	      const uint32_t message_count) :
+    m_self (ioa::get_aid ()),
+    m_alarm_state (ALARM_SET),
+    m_send_state (SEND_SEND),
+    m_address (address),
+    m_interval (0, 1000000 / message_rate),
+    m_message_size (message_size),
+    m_message_count (message_count)
+  {
+    ioa::automaton_manager<ioa::alarm_automaton>* alarm = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
+    ioa::make_binding_manager (this,
+			       &m_self,
+			       &udp_sender::set_alarm,
+			       alarm,
+			       &ioa::alarm_automaton::set);
+    ioa::make_binding_manager (this,
+			       alarm,
+			       &ioa::alarm_automaton::alarm,
+			       &m_self,
+			       &udp_sender::alarm_interrupt);
+    
+    ioa::automaton_manager<ioa::udp_sender_automaton>* sender = new ioa::automaton_manager<ioa::udp_sender_automaton> (this, ioa::make_generator<ioa::udp_sender_automaton> (message_size));
+    ioa::make_binding_manager (this,
+			       &m_self, &udp_sender::send,
+			       sender, &ioa::udp_sender_automaton::send);
+    
[…]

### 2011-08-10 papers/paper1/paper.tex
Added a few examples to paper.  Not sure how much more I can expand...

@@
 \end{abstract}
-
-\section{Adam's Section}
-A couple of things were hard to digest, the first of which being indirect triggers.  
-It's hard to swallow the idea that you can queue an action but it's time of execution is far from guaranteed.  
-The second hardest thing was simply the intimidation factor of realizing your entire summer would be working with an unknown, rather large library that nobody but the creator had used before--which was modelled after a complicated formalism I had never heard of that no one had bothered to fully develop since its inception.
-That's quite the pill to swallow.
-
-An thing that I learned was easy was to model the automata very closely to Lynch's design.  
-It was much easier for me to learn the relationship between the formalism and the code by looking at well-modelled automata, so the same would extend onto someone looking at the formalism and our code.
-Maintaining this allows for a less steep learning curve, and hopefully contributes to the automata maintaining their structural integrity as more people use the library.
-
-It's hard to describe what is ``harder'' to implement using I/O Automata.  
-Though I/O Automata present a \emph{different} problem set than using threads, the resulting problem set is respectively easier and more intuitive to handle.
-
-However, what is easer to implement is much more plentiful.
-The design and debugging of the program is much more intuitive.
-It allowed for a high degree of abstraction, which is key for easily-understandable code.
-It allowed \emph{extremely} easy refactoring of any class when compared to the alleged dregs of refactoring a thread-intensive class.
-
-In regards to building a protocol, I learned that it is extremely helpful to add and then test small bits of functionality at a time, in order to get a small working product.
-Once this is complete, repeat by adding or extending the functionality.
-Though this may seem trivial, calculated baby-steps made developing a protocol seem much easier than trying it all at once.
 
@@
 
+\section{Adam's Section}
+%% Go into more detail / provide more examples
+A couple of things were hard to digest, the first of which being indirect triggers.  
+It's hard to swallow the idea that you can queue an action but it's time of execution is far from guaranteed.  
+The second hardest thing was simply the intimidation factor of realizing your entire summer would be working with an unknown, rather large library that nobody but the creator had used before--which was modelled after a complicated formalism I had never heard of that no one had bothered to fully develop since its inception.
+That's quite the pill to swallow.
+
+One thing that I learned was easy was to model the automata very closely to Lynch's design.  
+It was much easier for me to learn the relationship between the formalism and the code by looking at well-modelled automata, so the same would extend onto someone looking at the formalism and our code.
+Maintaining this allows for a less steep learning curve, and hopefully contributes to the automata maintaining their structural integrity as more people use the library.
+An example of this would be correctly classifying input, output, and internal actions.
+
+It's hard to describe what is ``harder'' to implement using I/O Automata.  
+Though I/O Automata present a \emph{different} problem set than using threads, the resulting problem set is respectively easier and more intuitive to handle.
+
+However, what is easer to implement is much more plentiful.
+The design and debugging of the program is much more intuitive.
+It allowed for a high degree of abstraction, which is key for easily-understandable code.
+It allowed \emph{extremely} easy refactoring of any class when compared to the alleged dregs of refactoring a thread-intensive class.
+
+In regards to building a protocol, I learned that it is extremely helpful to add and then test small bits of functionality at a time, in order to get a small working product.
+Once this is complete, repeat by adding or extending the functionality.
+Though this may seem trivial, calculated baby-steps made developing a protocol seem much easier than trying it all at once.
+For example, developing a simple transfer channel to correctly model our mftp automaton.  After this was well tested, we then moved it to transfer over multicast/broadcast.
+
 \section{Related Work}

### 2011-08-10 include/mftp/message.hpp
Changed request to 129 at a time.

@@ -9,7 +9,7 @@ namespace mftp {
   const uint32_t REQUEST = 1;
   const uint32_t MATCH = 2;
 
-  const uint32_t SPANS_SIZE = 63;
+  const uint32_t REQUEST_SIZE = 129;
   const uint32_t MATCHES_SIZE = 12;
 
   struct fragment
@@ -32,57 +32,26 @@ namespace mftp {
     }
   };
 
-  struct span_t
-  {
-    uint32_t start;
-    uint32_t stop;
-
-    span_t& operator= (const std::pair<uint32_t, uint32_t>& p) {
-      start = p.first;
-      stop = p.second;
-      return *this;
-    }
-
-    void convert_to_network () {
-      start = htonl (start);
-      stop = htonl (stop);
-    }
-    
-    void convert_to_host () {
-      start = ntohl (start);
-      stop = ntohl (stop);
-    }
-
-  };
-    
   struct request
   {
     fileid fid;
-    uint32_t span_count;
-    span_t spans[SPANS_SIZE];
+    uint32_t fragments[REQUEST_SIZE];
     
     void convert_to_network () {
       fid.convert_to_network ();
-      for (uint32_t i = 0; i < span_count; ++i) {
-	spans[i].convert_to_network ();
+      for (uint32_t i = 0; i < REQUEST_SIZE; ++i) {
+	fragments[i] = htonl (fragments[i]);
       }
-      span_count = htonl (span_count);
     }
     
     bool convert_to_host () {
       fid.convert_to_host ();
-      span_count = ntohl (span_count);
-      
-      if (span_count == 0 || span_count > SPANS_SIZE) {
-	return false;
-      }
       
       mfileid mid (fid);
-      for (uint32_t i = 0; i < span_count; ++i) {
-	spans[i].convert_to_host ();
+      for (uint32_t i = 0; i < REQUEST_SIZE; ++i) {
+	fragments[i] = ntohl (fragments[i]);
 	
-	if (!((spans[i].start < spans[i].stop) &&
-	      (spans[i].stop <= mid.get_fragment_count ()))) {
+	if (fragments[i] >= mid.get_fragment_count ()) {
 	  return false;
 	}
       }
@@ -162,14 +131,12 @@ namespace mftp {
 
     message (request_type /* */,
 	     const fileid& fileid,
-	     uint32_t span_count,
-	     const span_t * spans)
+	     const uint32_t* fragments)
     {
       header.message_type = REQUEST;
[…]

### 2011-08-10 include/mftp/mftp_automaton.hpp

@@ -4,7 +4,6 @@
 #include <mftp/match.hpp>
 #include <mftp/mftp_channel_automaton.hpp>
 #include <ioa/alarm_automaton.hpp>
-#include <mftp/interval_set.hpp>
 
 #include <queue>
 #include <set>
@@ -28,8 +27,10 @@ namespace mftp {
     static const ioa::time ALARM_INTERVAL;
     static const ioa::time INIT_INTERVAL;
     static const ioa::time MAX_INTERVAL;
-    static const ioa::time REQUEST_INTERVAL;
     static const uint32_t MAX_FRAGMENT_COUNT;
+    static const uint32_t SHUFFLE_LIMIT;
+    static const uint32_t REREQUEST_NUMERATOR;
+    static const uint32_t REREQUEST_DENOMINATOR;
 
     ioa::handle_manager<mftp_automaton> m_self;
     file m_file;
@@ -46,23 +47,25 @@ namespace mftp {
     uint32_t m_num_match_in_sendq; // Number of matches in the send queue.
 
     // Answering requests.
-    interval_set<uint32_t> m_requests; // Set of intervals indicating fragments that have been requested.
-    uint32_t m_last_sent_idx; // Index of last fragment sent.
+    std::set<uint32_t> m_requests_set; // Set of fragments that have been requested.
+    std::deque<uint32_t> m_requests_deque; // Superset of set organized as deque.
+    uint32_t m_inserts_since_shuffle; // Number of inserts since shuffle.
+
+    // Making requests.
+    uint32_t m_request_idx; // Index for requests.
+    uint32_t m_last_request_size; // Number of fragments in last request (<= REQUEST_SIZE).
+    uint32_t m_fragments_since_request; // Number of fragments received since request.
 
     // Timestamps for certain events.
     ioa::time m_frag_recv_time; // Time when this automaton last received a fragment (of this file).
-    ioa::time m_request_time; // Time when this automaton last sent a request.
+    ioa::time m_request_timeout_start; // Time when this automaton last sent a request or received a fragment (of this file).
     ioa::time m_match_time; // Time when this automaton last received a match (of this file).
 
     // Periodic acitivities.
     alarm_state_t m_alarm_state; // State of alarm state machine.
     ioa::time m_announcement_interval; // Must wait this amount of time after m_fragment_time to send an announcement.
+    ioa::time m_request_interval; // Must wait this amount of time after m_request_time to send a request for timeout.
[…]

### 2011-08-10 lib/mftp_automaton.cpp

@@ -4,8 +4,10 @@ namespace mftp {
   const ioa::time mftp_automaton::ALARM_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::INIT_INTERVAL (1, 0); // 1 second
   const ioa::time mftp_automaton::MAX_INTERVAL (64, 0); // slightly over 1 minute
-  const ioa::time mftp_automaton::REQUEST_INTERVAL (1, 0);
   const uint32_t mftp_automaton::MAX_FRAGMENT_COUNT (1); // Number of fragments allowed in sendq.
+  const uint32_t mftp_automaton::SHUFFLE_LIMIT (REQUEST_SIZE);
+  const uint32_t mftp_automaton::REREQUEST_NUMERATOR (9);
+  const uint32_t mftp_automaton::REREQUEST_DENOMINATOR (10);
 
   // Not matching.
   mftp_automaton::mftp_automaton (const file& file,
@@ -21,20 +23,20 @@ namespace mftp {
     m_num_frag_in_sendq (0),
     m_num_req_in_sendq (0),
     m_num_match_in_sendq (0),
-    m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
+    m_inserts_since_shuffle (0),
+    m_request_idx (0),
+    m_last_request_size (0),
+    m_fragments_since_request (0),
     m_alarm_state (SET_READY),
     m_announcement_interval (INIT_INTERVAL),
+    m_request_interval (INIT_INTERVAL),
     m_match_interval (INIT_INTERVAL),
-    m_request_count (1),
-    m_num_frags_in_last_req (0),
-    m_num_new_frags_since_req (0),
     m_fragments_since_report (0),
     m_progress_threshold (progress_threshold),
     m_matching (false),
     m_get_matching_files (false),
     m_suicide_flag (suicide),
-    m_reported (m_file.complete ()),
-    m_frags_sent (0)
+    m_reported (m_file.complete ())
   {
     create_bindings ();
   }
@@ -56,13 +58,14 @@ namespace mftp {
     m_num_frag_in_sendq (0),
     m_num_req_in_sendq (0),
     m_num_match_in_sendq (0),
-    m_last_sent_idx (rand () % m_mfileid.get_fragment_count ()),
+    m_inserts_since_shuffle (0),
+    m_request_idx (0),
+    m_last_request_size (0),
+    m_fragments_since_request (0),
[…]

### 2011-08-10 lib/mftp_automaton.cpp
Removed printf.

@@ -167,7 +167,6 @@ namespace mftp {
 	// Increase the interval.
 	m_request_interval += m_request_interval;
 	m_request_interval = std::min (m_request_interval, MAX_INTERVAL);
-	std::cout << "Timeout" << std::endl;
       }
 
       if (timeout || percent) {

### 2011-08-10 include/mftp/message.hpp
Repaired from removing ioa::buffer.

@@ -2,7 +2,6 @@
 #define __message_hpp__
 
 #include <mftp/mfileid.hpp>
-#include <ioa/buffer.hpp>
 
 namespace mftp {
   const uint32_t FRAGMENT = 0;
@@ -130,28 +129,18 @@ namespace mftp {
     }
 
     message (request_type /* */,
-	     const fileid& fileid,
-	     const uint32_t* fragments)
+	     const fileid& fileid)
     {
       header.message_type = REQUEST;
       req.fid = fileid;
-      for (uint32_t i = 0; i < REQUEST_SIZE; i++){
-	req.fragments[i] = fragments[i];
-      }
     }
 
     message (match_type /* */,
-	     const fileid& fid,
-	     uint32_t match_count,
-	     const fileid * matches)
+	     const fileid& fid)
     {
       header.message_type = MATCH;
       mat.fid = fid;
-      mat.match_count = match_count;
-      for (uint32_t i = 0; i < match_count; i++) {
-	mat.matches[i] = matches[i];
-      }
-
+      mat.match_count = 0;
     }
 
     void convert_to_network () {
@@ -184,47 +173,6 @@ namespace mftp {
     }
   };
 
-  struct message_buffer :
-    public ioa::buffer_interface
-  {
-    message msg;
-
-    message_buffer (fragment_type type,
-		    const fileid& fileid,
-		    uint32_t idx,
-		    const void* data) :
-      msg (type, fileid, idx, data)
-    { }
-
-    message_buffer (request_type type,
-		    const fileid& fileid,
-		    const uint32_t* fragments) :
-      msg (type, fileid, fragments)
-    { }
-
-    message_buffer (match_type type,
-		    const fileid& fid,
-		    uint32_t match_count,
-		    const fileid * matches) :
-      msg (type, fid, match_count, matches)
-    { }
-
-    const void* data () const {
-      return &msg;
-    }
-
-    size_t size () const {
-      return sizeof (message);
-    }
-
-    void convert_to_network () {
-      msg.convert_to_network ();
-    }
-
-    void convert_to_host () {
-      msg.convert_to_host ();
-    }
-  };
 }
 
 #endif

### 2011-08-10 include/mftp/mftp_automaton.hpp

@@ -40,7 +40,7 @@ namespace mftp {
     ioa::handle_manager<mftp_channel_automaton> m_channel; // The channel for sending/receiving.
 
     // Sending.
-    std::queue<ioa::const_shared_ptr<message_buffer> > m_sendq; // Send queue.
+    std::queue<ioa::const_shared_ptr<std::string> > m_sendq; // Send queue.
     send_state_t m_send_state; // State of send state machine.
     uint32_t m_num_frag_in_sendq; // Number of fragments in the send queue.
     uint32_t m_num_req_in_sendq; // Number of requests in the send queue.
@@ -105,15 +105,15 @@ namespace mftp {
     void create_bindings ();
     void schedule () const;
     void process_match_candidate (const file& f);
-    message_buffer* get_fragment (uint32_t idx);
+    std::string* get_fragment (uint32_t idx);
     void send_announcement ();
     void send_request ();
     void send_match (bool reset);
     void add_match (const fileid& fid);
 
     bool send_precondition () const;
-    ioa::const_shared_ptr<message_buffer> send_effect ();
-    V_UP_OUTPUT (mftp_automaton, send, ioa::const_shared_ptr<message_buffer>);
+    ioa::const_shared_ptr<std::string> send_effect ();
+    V_UP_OUTPUT (mftp_automaton, send, ioa::const_shared_ptr<std::string>);
 
     void send_complete_effect ();
     UV_UP_INPUT (mftp_automaton, send_complete);

### 2011-08-10 include/mftp/mftp_channel_automaton.hpp

@@ -15,7 +15,7 @@ namespace mftp {
   {
   private:
     ioa::handle_manager<mftp_channel_automaton> m_self;
-    typedef std::pair<ioa::const_shared_ptr<message_buffer>, ioa::aid_t> message_aid;
+    typedef std::pair<ioa::const_shared_ptr<std::string>, ioa::aid_t> message_aid;
     std::list<message_aid > m_outgoing_messages;
     std::set<ioa::aid_t> m_outgoing_set;
     ioa::aid_t m_pending_aid;
@@ -44,10 +44,10 @@ namespace mftp {
     void observe (ioa::observable* o);
     void purge (const ioa::aid_t aid);
 
-    void send_effect (const ioa::const_shared_ptr<message_buffer>& message,
+    void send_effect (const ioa::const_shared_ptr<std::string>& message,
 		      ioa::aid_t aid);
   public:
-    V_AP_INPUT (mftp_channel_automaton, send, ioa::const_shared_ptr<message_buffer>);
+    V_AP_INPUT (mftp_channel_automaton, send, ioa::const_shared_ptr<std::string>);
 
   private:
     bool send_out_precondition () const;

### 2011-08-10 lib/mftp_automaton.cpp

@@ -141,9 +141,7 @@ namespace mftp {
       const ioa::time now = ioa::time::now ();
       if (m_frag_recv_time + m_announcement_interval <= now) {
 	// Send a fragment.
-	message_buffer* m = get_fragment (m_file.get_first_fragment_index ());
-	m->convert_to_network ();
-	m_sendq.push (ioa::const_shared_ptr<message_buffer> (m));
+	m_sendq.push (ioa::const_shared_ptr<std::string> (get_fragment (m_file.get_first_fragment_index ())));
 	++m_num_frag_in_sendq;
 	m_announcement_interval += m_announcement_interval;
 	m_announcement_interval = std::min (m_announcement_interval, MAX_INTERVAL);
@@ -176,13 +174,13 @@ namespace mftp {
 	}
 
 	std::set<uint32_t> frags;
-	uint32_t fragments[REQUEST_SIZE];
+	message m (request_type (), m_fileid);
 
 	uint32_t idx = 0;
 	while (idx < REQUEST_SIZE) {
 	  if (interval_set<uint32_t>::intersect (*pos, std::make_pair (m_request_idx, m_request_idx + 1))) {
 	    frags.insert (m_request_idx);
-	    fragments[idx++] = m_request_idx++;
+	    m.req.fragments[idx++] = m_request_idx++;
 	  }
 	  else {
 	    ++pos;
@@ -195,9 +193,8 @@ namespace mftp {
 
 	m_last_request_size = frags.size ();
 
-	message_buffer* m = new message_buffer (request_type (), m_fileid, fragments);
-	m->convert_to_network ();
-	m_sendq.push (ioa::const_shared_ptr<message_buffer> (m));
+	m.convert_to_network ();
+	m_sendq.push (ioa::const_shared_ptr<std::string> (new std::string (reinterpret_cast<char *> (&m), sizeof (m))));
 	++m_num_req_in_sendq;
 
 	// Reset.
@@ -230,17 +227,14 @@ namespace mftp {
 	// Create match messages.
 	std::set<fileid>::const_iterator pos = m_matches.begin ();
 	while (pos != m_matches.end ()) {
-	  uint32_t count = 0;
-	  fileid matches[MATCHES_SIZE];
-	  while (count < MATCHES_SIZE && pos != m_matches.end ()) {
-	    matches[count] = *pos;
+	  message m (match_type (), m_fileid);
+	  while (m.mat.match_count < MATCHES_SIZE && pos != m_matches.end ()) {
[…]

### 2011-08-10 lib/mftp_channel_automaton.cpp

@@ -82,7 +82,7 @@ namespace mftp {
     m_outgoing_completes.erase (aid);
   }
 
-  void mftp_channel_automaton::send_effect (const ioa::const_shared_ptr<message_buffer>& message,
+  void mftp_channel_automaton::send_effect (const ioa::const_shared_ptr<std::string>& message,
 					    ioa::aid_t aid) {
     
     if (m_outgoing_set.count (aid) == 0 &&

### 2011-08-10 src/get.cpp

@@ -48,8 +48,8 @@ namespace jam {
 	  //Create the instance server.
 	  uuid_t id;
 	  uuid_generate (id);
-	  ioa::buffer buff;
-	  buff.append (id, sizeof (uuid_t));
+	  std::string buff;
+	  buff.append (reinterpret_cast<char*> (id), sizeof (uuid_t));
 	  buff.append (m_filename.c_str (), m_filename.size ());
 
 	  mftp::file f (buff.data (), buff.size (), QUERY_TYPE);

### 2011-08-10 src/share.cpp

@@ -46,14 +46,14 @@ namespace jam {
 	  mftp::fileid copy = file.get_mfileid ().get_fileid ();
 	  copy.convert_to_network ();
 
-	  ioa::buffer buff;
-	  buff.append (&copy, sizeof (mftp::fileid));
+	  std::string buff;
+	  buff.append (reinterpret_cast<char *> (&copy), sizeof (mftp::fileid));
 	  buff.append (m_sharename.c_str (), m_sharename.size ());
 
 	  mftp::file meta (buff.data (), buff.size (), META_TYPE);
 
-	  ioa::buffer buff2;
-	  buff2.append (m_sharename.c_str (), m_sharename.size ());
+	  std::string buff2;
+	  buff2.append (m_sharename);
 
 	  mftp::file query (buff.data (), buff.size (), QUERY_TYPE);
 	

### 2011-08-10 src/udp_sender.cpp

@@ -27,7 +27,7 @@ private:
   const ioa::time m_interval;
   const uint32_t m_message_size;
   uint32_t m_message_count;
-  std::queue<ioa::const_shared_ptr<ioa::buffer> > m_sendq;
+  std::queue<ioa::const_shared_ptr<std::string> > m_sendq;
 
 public:
   udp_sender (const ioa::inet_address& address,
@@ -87,11 +87,12 @@ private:
   V_UP_OUTPUT (udp_sender, set_alarm, ioa::time);
 
   void alarm_interrupt_effect () {
-    ioa::buffer* buf = new ioa::buffer (m_message_size);
-    uint32_t t = htonl (m_message_count--);
-    memcpy (buf->data (), &t, sizeof (t));
     m_alarm_state = ALARM_SET;
-    m_sendq.push (ioa::const_shared_ptr<ioa::buffer> (buf));
+
+    uint32_t t = htonl (m_message_count--);
+    std::string* buf = new std::string (reinterpret_cast<char *> (&t), sizeof (t));
+    buf->append ('\0', m_message_size - sizeof (t));
+    m_sendq.push (ioa::const_shared_ptr<std::string> (buf));
     if (m_sendq.size () > 1) {
       std::cout << "Queuing: " << m_sendq.size () << std::endl;
     }
@@ -105,7 +106,7 @@ private:
 
   ioa::udp_sender_automaton::send_arg send_effect () {
     m_send_state = SEND_WAIT;
-    ioa::const_shared_ptr<ioa::buffer> m = m_sendq.front ();
+    ioa::const_shared_ptr<std::string> m = m_sendq.front ();
     m_sendq.pop ();
     return ioa::udp_sender_automaton::send_arg (m_address, m);
   }

### 2011-08-11 include/mftp/file.hpp
Repaced many unsigned chars with chars and strings.

@@ -12,27 +12,23 @@ namespace mftp {
 
   private:
     mfileid m_mfileid;
-    uint8_t* m_data;
+    std::string m_data;
     uint32_t m_have_count;
 
   public:
-    file (const std::string&,
-	  const uint32_t);
     file (const fileid& f);
     file (const file& other);
-    file (const void* ptr,
+    file (const char* ptr,
 	  uint32_t size,
 	  uint32_t type);
-    ~file ();
     
     const mfileid& get_mfileid () const;
-    void* get_data_ptr ();
-    const void* get_data_ptr () const;
+    const std::string& get_data () const;
     uint32_t have_count () const;
     bool complete () const;
     bool empty () const;
     bool write_chunk (const uint32_t idx,
-		      const void* data);
+		      const char* data);
     uint32_t get_first_fragment_index () const;
   };
 }

### 2011-08-11 include/mftp/mfileid.hpp

@@ -40,7 +40,7 @@ namespace mftp {
       calculate_lengths ();
     }
 
-    void set_hash (const uint8_t* hash) {
+    void set_hash (const char* hash) {
       memcpy (m_fileid.hash, hash, HASH_SIZE);
     }
 

### 2011-08-11 lib/file.cpp

@@ -1,93 +1,31 @@
 #include <mftp/file.hpp>
 #include "sha2_256.hpp"
 
-#include <math.h>
-#include <cstdlib>
-#include <fcntl.h>
-#include <sys/stat.h>
-#include <cassert>
-#include <unistd.h>
-#include <stdio.h>
-
-#include <iostream>
-
 namespace mftp {
-
-  file::file (const std::string& name,
-	      const uint32_t type)
-  {
-    int fd = open (name.c_str (), O_RDONLY);
-    assert (fd != -1);
-
-    struct stat stats;
-
-    if (fstat (fd, &stats) >= 0) {
-      m_mfileid.set_length (stats.st_size);
-    }
-    else {
-      perror ("fstat");
-      exit (EXIT_FAILURE);
-    }
-
-    m_mfileid.set_type (type);
-
-    // We have all of the file.
-    m_have_count = m_mfileid.get_fragment_count ();
-
-    /*
-      Layout of the file will be
-      +-------------------------+-------------+
-      |  DATA (original_length) | 0 (padding) |
-      +-------------------------+-------------+
-    */
-
-    // Allocate storage for the data and read it.
-    m_data = new unsigned char[m_mfileid.get_final_length ()];
-    if (read (fd, m_data, m_mfileid.get_original_length ()) != static_cast<ssize_t> (m_mfileid.get_original_length ())) {
-      perror ("read");
-      exit (EXIT_FAILURE);
-    }
-    close (fd);
-
-    // Clear the padding.
-    for (uint32_t idx = m_mfileid.get_original_length (); idx < m_mfileid.get_padded_length (); ++idx) {
-      m_data[idx] = 0;
-    }
-
-    sha2_256 digester;
-    digester.update (m_data, m_mfileid.get_final_length ());
-    digester.finalize ();
-    unsigned char samp[HASH_SIZE];
-    digester.get (samp);
-    m_mfileid.set_hash (samp);
-  }
 
   file::file (const fileid& f) :
     m_mfileid (f),
-    m_data (new unsigned char[m_mfileid.get_final_length ()]),
     m_have_count (0)
-  { 
+  {
+    m_data.resize (m_mfileid.get_final_length ());
     m_dont_have.insert (std::make_pair (0, m_mfileid.get_fragment_count ()));
   }
 
   file::file (const file& other) :
     m_dont_have (other.m_dont_have),
     m_mfileid (other.m_mfileid),
+    m_data (other.m_data),
[…]

### 2011-08-11 lib/mftp_automaton.cpp

@@ -526,7 +526,7 @@ namespace mftp {
   }
 
   std::string* mftp_automaton::get_fragment (uint32_t idx) {
-    message m (fragment_type (), m_fileid, idx, static_cast<const char*> (m_file.get_data_ptr ()) + idx * FRAGMENT_SIZE);
+    message m (fragment_type (), m_fileid, idx, m_file.get_data ().data () + idx * FRAGMENT_SIZE);
     m.convert_to_network ();
     return new std::string (reinterpret_cast<char*> (&m), sizeof (m));
   }

### 2011-08-11 lib/sha2_256.cpp

@@ -89,7 +89,7 @@ sha2_256::sha2_256 () :
 }
 
 sha2_256::sha2_256 (const unsigned int length,
-		    const unsigned char* hash) :
+		    const char* hash) :
   total_length (length),
   buf_length (0)
 {
@@ -146,7 +146,7 @@ void sha2_256::finalize () {
   process ();
 }
 
-void sha2_256::get (unsigned char* hash) const {
+void sha2_256::get (char* hash) const {
   assert (buf_length == 0);
   for (unsigned int idx = 0; idx < 8; ++idx) {
     unsigned int x = htonl (h[idx]);
@@ -154,7 +154,7 @@ void sha2_256::get (unsigned char* hash) const {
   }
 }
 
-void sha2_256::update (const unsigned char* data,
+void sha2_256::update (const char* data,
 		       size_t length) {
   size_t read_idx = 0;
   while (read_idx < length) {

### 2011-08-11 lib/sha2_256.hpp

@@ -18,7 +18,7 @@ private:
   unsigned int h[8];
   unsigned int total_length;
   unsigned int buf_length;
-  unsigned char buf[64];
+  char buf[64];
 
   static unsigned int right_rotate (unsigned int x,
 				    unsigned int s);
@@ -27,10 +27,10 @@ private:
 public:
   sha2_256 ();
   sha2_256 (const unsigned int length,
-	    const unsigned char* hash);
+	    const char* hash);
   void finalize ();
-  void get (unsigned char* hash) const;
-  void update (const unsigned char* data,
+  void get (char* hash) const;
+  void update (const char* data,
 	       size_t length);
 };
 

### 2011-08-11 src/dd.hpp

@@ -1,90 +1,0 @@
-#ifndef __dd_hpp__
-#define __dd_hpp__
-
-#include <mftp/mftp.hpp>
-
-#include <string.h>
-#include <stdio.h>
-
-namespace dd {
-  
-#define DESCRIPTION 0
-#define SPECIFICATION 1
-  
-  const ioa::inet_address SEND_ADDR ("224.0.0.137", 54321);
-  const ioa::inet_address LOCAL_ADDR ("0.0.0.0", 54321);
-  
-  struct description_predicate :
-    public mftp::match_candidate_predicate
-  {
-    bool operator() (const mftp::fileid& fid) const {
-      return fid.type == DESCRIPTION;
-    }
-
-    description_predicate* clone () const {
-      return new description_predicate (*this);
-    }
-  };
-    
-  struct full_description_predicate :
-    public mftp::match_predicate
-  {
-    std::string specification;
-    
-    full_description_predicate (const std::string& spec) :
-      specification (spec)
-    { }
-
-    bool operator() (const mftp::file& f) const {
-      assert (f.get_mfileid ().get_fileid ().type == DESCRIPTION);
-      std::string description (static_cast<const char*> (f.get_data_ptr ()), f.get_mfileid ().get_original_length ());
-        if (description.find (specification) != std::string::npos) {
-	return true;
-      }
-      return false;
-    }
-
-    full_description_predicate* clone () const {
-      return new full_description_predicate (*this);
-    }
-
-  };
-  
-  struct specification_predicate :
-    public mftp::match_candidate_predicate
-  {
-    bool operator() (const mftp::fileid& fid) const {
-      return fid.type == SPECIFICATION;
-    }
-
-    specification_predicate* clone () const {
-      return new specification_predicate (*this);
-    }
-  };
-  
-  struct full_specification_predicate :
-    public mftp::match_predicate
-  {
-    std::string description;
-
-    full_specification_predicate (const std::string& desc) :
-      description (desc)
-    { }
-
-    bool operator() (const mftp::file& f) const {
-      assert (f.get_mfileid ().get_fileid ().type == SPECIFICATION);
-      std::string specification (static_cast<const char*> (f.get_data_ptr ()), f.get_mfileid ().get_original_length ());
-        if (description.find (specification) != std::string::npos) {
-	return true;
-      }
-      return false;
-    }
-
-    full_specification_predicate* clone () const {
-      return new full_specification_predicate (*this);
-    }
-  };
-
-}
-
-#endif

### 2011-08-11 src/describe.cpp

@@ -1,64 +1,0 @@
-#include "dd.hpp"
-#include <ioa/global_fifo_scheduler.hpp>
-#include <ioa/ioa.hpp>
-
-#include <iostream>
-#include <stdio.h>
-#include <string>
-
-namespace dd {
-
-  class desc_automaton :
-    public ioa::automaton,
-    public ioa::observer
-
-  {
-  private:
-    ioa::automaton_manager<mftp::mftp_channel_automaton> * channel;
-
-    const std::string m_filename;
-
-  public:
-    desc_automaton (const std::string& fname) :
-      m_filename (fname)
-    {
-      mftp::file file (m_filename, DESCRIPTION);
-      std::cout << "Sharing " << m_filename << std::endl;
-
-      channel = new ioa::automaton_manager<mftp::mftp_channel_automaton> (this, ioa::make_generator<mftp::mftp_channel_automaton> (dd::SEND_ADDR, dd::LOCAL_ADDR, true));
-
-      add_observable (channel);
-    }
-
-    void observe (ioa::observable* o) {
-      if (o == channel && -1 == channel->get_handle ()) {
-	channel = 0;
-      }
-
-      if (channel != 0) {
-	if (channel->get_handle () != -1) {
-	  mftp::file file (m_filename, DESCRIPTION);
-
-	  std::string desc_string (static_cast<const char*> (file.get_data_ptr ()), file.get_mfileid ().get_original_length ());
-	  std::cout << "description: \n" << desc_string << std::endl;
-	  //Create the description server.
-	  new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (file, channel->get_handle (), specification_predicate (), full_specification_predicate (desc_string), false, false, 0));
-	}
-      }
-    }
-
-  };
-}
-
-int main (int argc, char* argv[]) {
-  if (argc != 2) {
-    std::cerr << "Usage: " << argv[0] << " FILE " << std::endl;
-    exit(EXIT_FAILURE);
-  }
-
-  const char* path = argv[1];
-  ioa::global_fifo_scheduler sched;
-  ioa::run (sched, ioa::make_generator<dd::desc_automaton> (path));
-
-  return 0;
-}

### 2011-08-11 src/get.cpp

@@ -69,7 +69,7 @@ namespace jam {
       new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (*meta_file, channel->get_handle (), query_predicate (m_filename), query_filename_predicate (m_filename), false, false, 0));
       
       mftp::fileid fid;
-      memcpy (&fid, meta_file->get_data_ptr (), sizeof (mftp::fileid));
+      meta_file->get_data ().copy (reinterpret_cast<char *> (&fid), sizeof (mftp::fileid));
       fid.convert_to_host();
       
       ioa::automaton_manager<mftp::mftp_automaton>* file_home = new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (mftp::file (fid), channel->get_handle(), false, FRAG_COUNT));
@@ -100,7 +100,8 @@ namespace jam {
 	exit (EXIT_FAILURE);
       }
 
-      size_t er = fwrite (f.get_data_ptr (), 1, f.get_mfileid ().get_original_length (), fp);
+      // TODO:  Use streams.
+      size_t er = fwrite (f.get_data ().data (), 1, f.get_mfileid ().get_original_length (), fp);
       if (er < f.get_mfileid ().get_original_length ()) {
 	std::cerr << "fwrite: couldn't write to " << m_filename << std::endl;
 	exit (EXIT_FAILURE);

### 2011-08-11 src/jam.hpp

@@ -41,7 +41,7 @@ namespace jam {
     { }
 
     bool operator() (const mftp::file& f) const {
-      return filename == std::string (static_cast<const char*> (f.get_data_ptr ()) + sizeof (mftp::fileid), filename.size ());
+      return filename == f.get_data ().substr (sizeof (mftp::fileid), filename.size ());
     }
 
     meta_filename_predicate* clone () const {
@@ -78,7 +78,7 @@ namespace jam {
     { }
 
     bool operator() (const mftp::file& f) const {
-      return filename == std::string (static_cast<const char*> (f.get_data_ptr ()) + sizeof (uuid_t), filename.size ());
+      return filename == f.get_data ().substr (sizeof (uuid_t), filename.size ());
     }
 
     query_filename_predicate* clone () const {

### 2011-08-11 src/share.cpp

@@ -5,6 +5,7 @@
 #include <iostream>
 #include <stdio.h>
 #include <string>
+#include <sys/stat.h>
 
 namespace jam {
   
@@ -24,10 +25,6 @@ namespace jam {
       m_filename (fname),
       m_sharename (sname)
     {
-      mftp::file file (m_filename, FILE_TYPE);
-      mftp::fileid copy = file.get_mfileid ().get_fileid ();
-      std::cout << "Sharing " << m_filename << " as " << (m_sharename + "-" + copy.to_string ()) << std::endl;
-
       channel = new ioa::automaton_manager<mftp::mftp_channel_automaton> (this, ioa::make_generator<mftp::mftp_channel_automaton> (jam::SEND_ADDR, jam::LOCAL_ADDR, true));
       
       add_observable (channel);
@@ -42,8 +39,32 @@ namespace jam {
 
       if (channel != 0) {
 	if (channel->get_handle () != -1) {
-	  mftp::file file (m_filename, FILE_TYPE);
+	  int fd = open (m_filename.c_str (), O_RDONLY);
+	  if (fd == -1) {
+	    perror ("open");
+	    exit (EXIT_FAILURE);
+	  }
+
+	  struct stat stats;
+	  if (fstat (fd, &stats) == -1) {
+	    perror ("fstat");
+	    exit (EXIT_FAILURE);
+	  }
+
+	  // TODO:  Use streams.
+	  char* buf = new char[stats.st_size];
+	  if (read (fd, buf, stats.st_size) != stats.st_size) {
+	    perror ("read");
+	    exit (EXIT_FAILURE);
+	  }
+
+	  close (fd);
+
+	  mftp::file file (buf, stats.st_size, FILE_TYPE);
+	  delete[] buf;
+
 	  mftp::fileid copy = file.get_mfileid ().get_fileid ();
+	  std::cout << "Sharing " << m_filename << " as " << (m_sharename + "-" + copy.to_string ()) << std::endl;
 	  copy.convert_to_network ();
 
 	  std::string buff;

### 2011-08-11 src/specify.cpp

@@ -1,104 +1,0 @@
-#include "dd.hpp"
-#include <ioa/global_fifo_scheduler.hpp>
-#include <ioa/ioa.hpp>
-
-#include <iostream>
-#include <queue>
-#include <set>
-
-namespace dd {
-
-  class spec_automaton :
-    public ioa::automaton,
-    private ioa::observer {
-
-  private:
-    ioa::handle_manager<spec_automaton> m_self;
-    ioa::automaton_manager<mftp::mftp_channel_automaton>* channel;
-    //std::map<mftp::fileid, ioa::automaton_manager<mftp::mftp_automaton>* > descs;
-    std::string m_filename;
-
-    //ioa::time m_transfer;  //necessary?
-
-    static const ioa::time PROGRESS_INTERVAL;
-
-  public:
-
-    spec_automaton (const std::string& sp) :
-      m_self (ioa::get_aid ()),
-      m_filename (sp)
-    {
-      channel = new ioa::automaton_manager<mftp::mftp_channel_automaton> (this, ioa::make_generator<mftp::mftp_channel_automaton> (dd::SEND_ADDR, dd::LOCAL_ADDR, true));
-
-      add_observable (channel);
-    }
-
-  private:
-    void schedule () const {  }
-
-    void observe (ioa::observable* o) {
-      if (o == channel && -1 == channel->get_handle ()) {
-	channel = 0;
-      }
-
-      if (channel != 0) {
-	if (channel->get_handle () != -1) {
-	  mftp::file spec (m_filename, SPECIFICATION);
-
-	  std::string spec_string (static_cast<const char*> (spec.get_data_ptr ()), spec.get_mfileid ().get_original_length ());
-	  std::cout << "specification: " << spec_string << std::endl;
-	  //Create the spec server.
-	  ioa::automaton_manager<mftp::mftp_automaton>* spec_server = new ioa::automaton_manager<mftp::mftp_automaton> (this, ioa::make_generator<mftp::mftp_automaton> (spec, channel->get_handle(), description_predicate (), full_description_predicate (spec_string), true, false, 0));
-
-	  ioa::make_binding_manager (this,
-				     spec_server, &mftp::mftp_automaton::match_complete,
-				     &m_self, &spec_automaton::match_complete);
-	}
-      }
-    }
-
-    void match_complete_effect (const ioa::const_shared_ptr<mftp::file>& f) {
[…]

### 2011-08-11 src/udp_receiver.cpp

@@ -1,66 +1,0 @@
-#include <ioa/ioa.hpp>
-#include <ioa/udp_receiver_automaton.hpp>
-#include <ioa/global_fifo_scheduler.hpp>
-
-#include <config.hpp>
-
-#include <iostream>
-
-class udp_receiver :
-  public ioa::automaton
-{
-private:
-  ioa::handle_manager<udp_receiver> m_self;
-  ioa::inet_address m_group;
-  ioa::inet_address m_address;
-
-public:
-  udp_receiver (const ioa::inet_address& group,
-		const ioa::inet_address& address) :
-    m_self (ioa::get_aid ()),
-    m_group (group),
-    m_address (address)
-  {
-    ioa::automaton_manager<ioa::udp_receiver_automaton>* receiver = new ioa::automaton_manager<ioa::udp_receiver_automaton> (this, ioa::make_generator<ioa::udp_receiver_automaton> (m_group, m_address));
-
-    ioa::make_binding_manager (this, receiver, &ioa::udp_receiver_automaton::receive, &m_self, &udp_receiver::receive);
-  }
-
-private:
-  void schedule () const { }
-
-  void receive_effect (const ioa::udp_receiver_automaton::receive_val& v) {
-    if (v.err != 0) {
-      char buf[256];
-#ifdef STRERROR_R_CHAR_P
-      std::cerr << "Couldn't receive udp_receiver_automaton: " << strerror_r (v.err, buf, 256) << std::endl;
-#else
-      strerror_r (v.err, buf, 256);
-      std::cerr << "Couldn't receive udp_receiver_automaton: " << buf << std::endl;
-#endif
-      self_destruct ();
-    }
-    else {
-      assert (v.buffer.get () != 0);
-      uint32_t idx;
-      memcpy (&idx, v.buffer->data (), sizeof (idx));
-      idx = ntohl (idx);
-      std::cout << idx << std::endl;
-    }
-  }
-  
-  V_UP_INPUT (udp_receiver, receive, ioa::udp_receiver_automaton::receive_val);
-
-};
-
-int
-main (int argc, char* argv[]) {
-  ioa::inet_address group ("224.0.0.137");
-  ioa::inet_address address ("0.0.0.0", 54321);
-
-  ioa::global_fifo_scheduler ss;
-  ioa::run (ss, ioa::make_generator<udp_receiver> (group, address));
-
-  return 0; 
-}
-

### 2011-08-11 src/udp_sender.cpp

@@ -1,149 +1,0 @@
-#include <iostream>
-#include <ioa/ioa.hpp>
-#include <ioa/global_fifo_scheduler.hpp>
-#include <ioa/alarm_automaton.hpp>
-#include <ioa/udp_sender_automaton.hpp>
-#include <queue>
-#include <config.hpp>
-
-class udp_sender :
-  public ioa::automaton
-{
-private:
-  enum alarm_state_t {
-    ALARM_SET,
-    ALARM_WAIT
-  };
-
-  enum send_state_t {
-    SEND_SEND,
-    SEND_WAIT
-  };
-
-  ioa::handle_manager<udp_sender> m_self;
-  alarm_state_t m_alarm_state;
-  send_state_t m_send_state;
-  const ioa::inet_address m_address;
-  const ioa::time m_interval;
-  const uint32_t m_message_size;
-  uint32_t m_message_count;
-  std::queue<ioa::const_shared_ptr<std::string> > m_sendq;
-
-public:
-  udp_sender (const ioa::inet_address& address,
-	      const uint32_t message_rate,
-	      const uint32_t message_size,
-	      const uint32_t message_count) :
-    m_self (ioa::get_aid ()),
-    m_alarm_state (ALARM_SET),
-    m_send_state (SEND_SEND),
-    m_address (address),
-    m_interval (0, 1000000 / message_rate),
-    m_message_size (message_size),
-    m_message_count (message_count)
-  {
-    ioa::automaton_manager<ioa::alarm_automaton>* alarm = new ioa::automaton_manager<ioa::alarm_automaton> (this, ioa::make_generator<ioa::alarm_automaton> ());
-    ioa::make_binding_manager (this,
-			       &m_self,
-			       &udp_sender::set_alarm,
-			       alarm,
-			       &ioa::alarm_automaton::set);
-    ioa::make_binding_manager (this,
-			       alarm,
-			       &ioa::alarm_automaton::alarm,
-			       &m_self,
-			       &udp_sender::alarm_interrupt);
-    
-    ioa::automaton_manager<ioa::udp_sender_automaton>* sender = new ioa::automaton_manager<ioa::udp_sender_automaton> (this, ioa::make_generator<ioa::udp_sender_automaton> (message_size));
-    ioa::make_binding_manager (this,
-			       &m_self, &udp_sender::send,
-			       sender, &ioa::udp_sender_automaton::send);
-    
[…]
